\chapter{Contributors to \madx}
\label{chap:contributors}

%\begin{center} 
%\textbf{\madx Module Keepers} 
%\end{center}
%
%\begin{table}[H]
%  \begin{center}
%    \vspace*{1mm}
%    \footnotesize
%    \begin{tabular}{|l|l|l|}
%      \hline
%      \textbf{Module}    &\textbf{Description}        &\textbf{Keeper} \\\hline
%      MAD-X Custodian    &Maintenance \& Debug        &L.~Deniau \\\hline
%      APERTURE           &Modeling LHC Aperture       &G.~Roy \\\hline
%      C6T                &SixTrack Converter          &G.~Roy \\\hline
%      CORORBIT           &Orbit Correction            &G.~Roy \\\hline
%      DYNAP              &Tracking Postprocessing     &F.~Zimmermann 
%\\\hline 
%      EMIT               &Emittance, Radiation        &R ~Tom\'as \\\hline  
%      ERROR              &Error Assignment            &G.~Roy \\\hline
%      IBS                &Intra-Beam Scattering       &F.~Zimmermann \\\hline
%      MAKETHIN           &Thin-lens Converter         &H.~Burkhardt \\\hline
%      MATCH              &Matching Procedures         &O.~Br\"uning \\\hline
%      PLOT               &Plotting                    &R.~de~Maria \\\hline
%      PTC                &PTC proper                  &E.~Forest - KEK \\\hline
%      madx\_ptc\_module    &MAD-X PTC Interface       &F.~Schmidt 
%\\\hline
%      PTC\_NORMAL        &NormalForm Coefficients     &tba \\\hline
%      PTC\_TRACK         &Thick-lens Tracking         &V.~Kapin - ITEP (RU) 
%\\\hline
%      PTC\_TWISS         &Ripken Optics Parameters    &tba \\\hline
%      SODD               &Resonance Driving Terms     &F.~Schmidt \\\hline
%      SURVEY             &Machine Survey              &F.~Tecker \\\hline
%      SXF                &Standard eXchange Format    &N.~Malitsky - BNL  
%\\\hline
%      THREADER           &Threading Closed-Orbit      &G.~Roy \\\hline
%      TOUSCHEK           &Touschek Effect             &C.~Milardi - IFNL/LNF 
%\\\hline
%      TWISS              &Classical Optics Parameters &F.~Schmidt \\\hline
%      TRACK              &Thin-lens Tracking          &A.~Koschik \\\hline      
%    \end{tabular}
%    \label{Module_Keepers}
%  \end{center}
%\end{table}



Lists are provided in alphabetical order.

Disclaimer: any omissions in these lists are accidental. 

Feel free to contact mad support
\href{mailto:mad.support@cern.ch}{(mad@cern.ch)} 
if you have been left out or some of your contributions are not listed. 

\section{MAD team}

The list includes all the persons officially working on the \madx project.

Laurent Deniau: project manager since june 2011.\\
Andrea Latina: since october 2011.\\
Ghislain Roy: since march 2013.\\
Piotr Skowronski: since january 2012.\\


\section{Module keepers and contributors}

This list includes all the volunteers who accepted to maintain and
develop some of the \madx modules.

Fanouria Antoniou: ibs.\\
Helmut Burkhardt: makethin.\\
Riccardo De Maria: match2.\\
Laurent Deniau: survey.\\
Valery Kapin (FNAL, Chicago): ptc\_track.\\
Emanuele Laface (ESS, Sweden): match.\\
Andrea Latina: trac, twiss.\\
Yannis Papaphilippou: dynap.\\
Ghislain Roy: aperture, error, cororbit, touschek, c6t, plot.\\
Piotr Skowronski: (ptc\_trac), ptc\_twiss, ptc\_normal, ptc\_module.\\
Rogelio Tomas: emit.

We are looking for volunteers, any help is welcome!


\section{Special contributors}

This list includes all the persons who have contributed exceptionally to the project in the past.

Frank Schmidt: project custodian from 2002 to 2011. \\
Etienne Forest: author of PTC \& FPP (written in F90). \\
Hans Grote: author of the core (written in C).


\section{Other contributors}

The list below includes the persons who have contributed significantly
to the \madx project in the past. Unless noted they were affiliated with
CERN at the time of their contributions. 

Ralph Assmann  (emit) \\
Oliver Bruning (match) \\
Hans Grote (Core in C, plot) \\
Werner Herr (error, cororbit) \\
Bernard Jeanneret (aperture) \\
Alex Koschik (thintrack) \\
Nikolay Malitsky - BNL, New-York  (sxf) \\
Eric McIntosh (memory leaks) \\
Jean Luc Nougaret (twiss, ptc\_twiss, C/Fortran wrappers, bug tracker) \\
Thys Risselada (threader, closed orbit) \\
Frank Schmidt (c6t, twiss, sodd, ptc\_twiss, ptc\_module, ptc\_normal) \\
Yipeng Sun (thintrack) \\
Frank Tecker (survey) \\
Andr\'e Verdier (survey) \\
Lingyun Yang (tpsa) \\
Frank Zimmermann (dynap, touschek, ibs) \\

\textbf{\madnine contributors}\\
Christoph Iselin is the author of MAD-9 and a major contributor to the Classic library.

\textbf{\madeight contributors}\\
Hans Grote and Christoph Iselin are the authors of \madeight.

\newpage

%% EOF
