%%%\title{Orbit Correction}
%  Changed by: Hans Grote, 13-Sep-2000 
%  Changed by: Werner Herr, 19-Jun-2002 
%  Changed by: Werner Herr, 02-Sep-2002 
%  Changed by: Werner Herr, 01-Oct-2002 
%  Changed by: Werner Herr, 01-May-2003 
%  Changed by: Werner Herr, 05-April-2004 
%  Changed by: Werner Herr, 01-Dec-2004 
%  Changed by: Werner Herr, 22-Oct-2008 
%  Changed by: Ghislain Roy, 2013-06-26
%  Changed by: Ghislain Roy, 2014-01-06
%  Changed by: Ghislain Roy, 2014-03-03

\chapter{Orbit Correction}  

This chapter describes the commands to correct a closed orbit or a
trajectory. The initial orbit or trajectory to be corrected can be
obtained  from an internal or external TFS table.   

The purpose of this orbit module is to provide some basic tools to
assess the performance of an orbit correction system of a machine in the
design phase.  

Although some interface is available, it cannot and does not provide the
full functionality expected from a dedicated online orbit correction and
steering program.  

%\input{cororbit/co_correct}
\section{CORRECT}  
\label{sec:correct}

The CORRECT statement makes a complete closed orbit  or trajectory
correction using the \textbf{computed} values at the monitors  from the
Twiss table.   

The CORRECT command has the following format (some options are valid
only for special algorithms):   

\madbox{
CORRECT, \=SEQUENCE=seqname, \\        
         \>FLAG=string, MODE=string, PLANE=string, \\
         \>COND=integer, NCORR=integer, \\
         \>SNGVAL=real, SNGCUT=real, \\
         \>MONERROR=integer, MONON=real, MONSCALE=real, \\
         \>CORRLIM=real, TWORING=logical, UNITS=real, \\           
         \>CORZERO=integer, ERROR=real,\\         
         \>ORBIT=table, MODEL=table, TARGET=table, \\
         \>BEAM1TAB=table, BEAM2TAB=table, \\
         \>EXTERN=logical, \\
         \>NAME\_COL=string, X\_COL=string, Y\_COL=string, \\
         \>CLIST=filename, MLIST=filename, \\
         \>RESOUT=integer, TWISSUM=integer;         
}

%% "correct: correct correct 0 0 "
%% "error    = [r, 1.e-5], "   v
%% "ncorr    = [i, 0], "       v
%% "monerror = [i, 0], "       v 
%% "monscale = [i, 0], "       v
%% "monon    = [r, 1], "       v
%% "moncut   = [r, 0], "       v
%% "corrlim  = [r, 1], "       v
%% "resout   = [i, 0], "       v
%% "sequence = [s, {none}], "  v
%% "clist    = [s, none], "    v
%% "mlist    = [s, none], "    v
%% "plane    = [s, x], "       v
%% "flag     = [s, ring], "    v
%% "mode     = [s, micado], "  v
%% "cond     = [i, 0], "       v
%% "twissum  = [i, 0], "       v
%% "orbit    = [s, none], "    v
%% "model    = [s, none], "    v
%% "corzero  = [i, 0], "       v
%% "tworing  = [l, false, true], " v
%% "beam1tab = [s, none], "    v
%% "beam2tab = [s, none], "    v
%% "target   = [s, none], "    v
%% "name_col = [s, none], "    v
%% "x_col    = [s, none], "    v
%% "y_col    = [s, none], "    v
%% "units    = [r, 1], "       v
%% "sngcut   = [r, 50], "      v
%% "sngval   = [r, 2], "       v
%% "extern   = [l, false, true]; " v


The command \texttt{CORRECT} is set up with defaults which should allow a
reasonable correction for most cases with a minimum of required options
(see Example 1 below).  

The orbit correction must always be preceded by \texttt{TWISS} commands  which
generate Twiss tables. The most recent Twiss table is assumed to contain
the optical parameters and the distorted orbits. 

The options used in the \texttt{CORRECT} command are: 

\begin{madlist}
   \ttitem{FLAG} can be "ring" or "line", either a circular machine
     or a trajectory is corrected.   
     \\ (Default: ring) 

   \ttitem{MODE} defines the method to be used for corrections. 
     \\ Available modes are \texttt{LSQ, MICADO} and \texttt{SVD}.  
     The first performs a
     least squares minimization using all available correctors. The mode
     \texttt{SVD} uses a Singular Value Decomposition to compute a correction
     using all available correctors. The latter can also be used to
     condition the response matrix for the modes \texttt{LSQ} or \texttt{MICADO} (using
     \texttt{COND=1}). It is highly recommended to precede a \texttt{LSQ} correction by a
     \texttt{SVD} conditioning (set \texttt{COND=1}).  
     \\ The mode \texttt{MICADO} is a "best kick" algorithm. Naive use or using
     it with a large number of correctors (see option \texttt{NCORR}) can get
     unexpected results. To avoid the creation of local bumps, it is
     recommended to precede a \texttt{MICADO} correction by a \texttt{SVD} conditioning
     (set \texttt{COND=1}).  
     \\ (Default: \texttt{MICADO})

   \ttitem{PLANE} With \texttt{PLANE=x}, the orbit correction
     is performed for the horizontal plane; With \texttt{PLANE=y}, 
     the correction is performed for the vertical plane. made. (This
     differs from the \madeight implementation).  
     \\ (Default: X for horizontal plane) 

   \ttitem{COND} When \texttt{COND=1}, a Singular Value Decomposition is
     performed and  the response matrix is conditioned to avoid linearly
     dependent correctors. This can be used to avoid creation of
     artificial bumps during a \texttt{LSQ} or \texttt{MICADO} correction.  
     Please note: this option is not robust since it
     depends on parameters which control the determination of singular
     values and redundant correctors. These can be set with the attributes
     \texttt{SNGVAL} and \texttt{SNGCUT}. Both parameters depend on the machine and may
     need adjustment. Default values are adjusted to large machines and give
     "reasonable" performance for smaller machines.

   \ttitem{NCORR} Defines the number of correctors to be used by the 
     \texttt{MICADO} algorithm. If \texttt{NCORR=0} all available
     correctors are used. Only used by the \texttt{MICADO} algorithm.
     \\ (Default: 0 i.e. all available correctors)

   \ttitem{SNGVAL}  Used to set the threshold for finding singular values
     with the \texttt{COND} attribute. (Hint: smaller number finds fewer singular
     values).  
     \\ \textbf{Use with care !}
     \\ (Default: 2.0)

   \ttitem{SNGCUT}  Used to set the threshold for finding redundant
     correctors with the \texttt{COND} attribute. (Hint: larger number finds fewer
     redundant correctors).  
     \\ \textbf{Use with extreme care !}
     \\ (Default: 50.0) 

   \ttitem{MONERROR} When \texttt{MONERROR} is 1, the \texttt{MREX} and \texttt{MREY} 
     alignment errors on monitors assigned  by \hyperref[sec:ealign]{EALIGN} 
     are taken into account, otherwise they are ignored.
     \\ (Default: 0)

   \ttitem{MONSCALE} When \texttt{MONSCALE} is 1, the MSCALX and
     MSCALY scaling errors on monitors assigned  by \hyperref[sec:ealign]{EALIGN}  
     are taken into account, otherwise they are ignored.
     \\ (Default: 0)

   \ttitem{MONON} takes a real number between 0.0 and 1.0. It
     determines the number of available monitors. If the command is
     given, each monitor is considered valid with a probability
     \texttt{MONON}. In the average a fraction $(1.0 - \texttt{MONON})$ 
     of the monitors will be disabled for the correction, i.e. they are 
     considered not existing.  This allows to study the effect of missing monitors.
     \\ (Default: 1.0 i.e. 100 \%) 

   \ttitem{CORRLIM}  A limit on the maximum corrector strength can be
     given and a WARNING is issued if it is exceeded by one or more
     correctors.  Please note: the strengths computed by the correction
     algorithms are NOT limited, only a warning is printed ! 
     \\ (Default: 1.0 mrad)

   \ttitem{TWORING} When true, the correction will be done on two rings at
     once. The only correction mode available in this case is
     \texttt{MICADO}. The attribute \texttt{ORBIT=table} is ignored and \texttt{BEAM1TAB=table} and
     \texttt{BEAM2TAB=table} are used instead.  
     \\ (Default: false)

   \ttitem{UNITS} when this parameter is set the value is a multiplier [TO BE COMPLETED] 
     \\ Default unknown

   \ttitem{CORZERO} an integer value to specify whether corrector settings 
     should be all reset to zero before starting the orbit correction 
     (\texttt{CORZERO}$>0$) 
     or corrector settings calculated by the orbit correction should be added to 
     existing corrector settings (\texttt{CORZERO}$=0$, Default).

   \ttitem{ERROR} specifies the maximum RMS value, in meters, of the 
     orbit to be reached by the correction algorithm, 
     e.g. \texttt{ERROR=1.e-3} for a 1~mm RMS target value. 
     \\ (Default: 1.e-5 m)

\end{madlist}

Normally the last active table provides the orbit to be
corrected and the model for the correction. This can be overwritten
by the appropriate options. Optionally, these tables can be given
names like in \hyperref[chap:twiss]{\texttt{TWISS, TABLE=name;}}. 
To use these named tables, one of the following optional
parameters must be  used:  

\begin{madlist}
   \ttitem{ORBIT} When this parameter is given, the orbit to be corrected
     is taken from a named table. The default is the last (named or
     unnamed) Twiss table.  

   \ttitem{MODEL} When this parameter is given, the model for the
     correction is taken from a named Twiss table. The default is the
     last (named or unnamed) Twiss table.  

   \ttitem{TARGET} When this parameter is given, the correction is made to
     a named target orbit, pre-computed with a \texttt{TWISS} command. Default is
     correction to the zero orbit.  

   \ttitem{EXTERN} (default: false): When false, the \texttt{ORBIT} and 
     \texttt{TARGET} table
     are assumed to be computed by \mad with a previous \texttt{TWISS}
     command. When set to true, that option allows to use twiss tables
     imported from an external file (with the \texttt{READTABLE} command), for
     example to use measured BPM data. In that case, the imported twiss
     table is allowed to contain coordinate data only at the location of
     the monitors. Note that the list has to be in order and that you
     cannot skip any locations of the monitors you have included. 
     
   \ttitem{NAME\_COL} The name of the column that contains the name of 
     monitors in the tables. (Default: "\texttt{name}")
   \ttitem{X\_COL} The name of the column that contains the \texttt{X} position 
     readings of monitors in the tables. (Default: "\texttt{X}")
   \ttitem{Y\_COL} The name of the column that contains the \texttt{Y} position 
     readings of monitors in the tables. (Default: "\texttt{Y}")
     
\end{madlist}

Example of use of \texttt{CORRECT} to reproduce a measured orbit: 
\begin{verbatim}
! To have a refererence optical model
twiss, table=twiss_ref;

! The bpm.tsv is a reduced Twiss file containing only lines for the BPMs
readtable, file="bpm.tsv", table="twiss_bpm";

! correct orbit using external measurements
correct, flag=ring, mode=micado, ncorr=5, cond=1 ,plane=x, extern,
         model=twiss_ref, orbit=twiss_ref, target=twiss_bpm, 
         error=1.0e-21;
\end{verbatim}


Two attributes affect the printing of tables and results: 
\begin{madlist}
   \ttitem{CLIST} the name of the file where corrector settings (in units of rad)
     before and after correction are printed.

   \ttitem{MLIST} the name of the file where monitor readings (in units of m) before
     and after correction are printed.

   \ttitem{RESOUT} This command outputs the results for all monitors and
     all correctors in a computer readable format if its integer
     argument is larger than 0. The argument is added to the
     output. Useful to analyze runs with loops to produce large
     statistics. \\
     \textbf{ATTENTION: May produce gigantic outputs for large
       machines.} 

   \ttitem{TWISSUM} If the attribute value is larger than 0, \texttt{CORRECT}
     prints maximum orbit and r.m.s. values for both planes, taken from the
     Twiss summary table, in computer readable form. This allows to analyze
     orbits etc. at elements that are not monitors or correctors. The
     argument is added to the output.  This attributes is used only to produce 
     output: no correction is made, and all other attributes are ignored.  
\end{madlist}

Obsolete commands or options:
\begin{verbatim}
ITERATE, ITERMAX             /* Done with loop feature in MAD commands */
THREADER, THRTOL, WRORBIT    /* Not part of orbit correction module */
M1LIST, M2LIST               /* Replaced by MLIST */
C1LIST, C2LIST               /* Replaced by CLIST */
GETORBIT, PUTORBIT           /* Replaced by generic TFS access */
GETKICK, PUTKICK             /* Replaced by generic TFS access */
\end{verbatim}

\textbf{EXAMPLES:}\\ 
for complete MAD input files see section on examples:

Example 1: correct orbit in horizontal plane, taken from most recent
Twiss table, using default algorithm (MICADO)
\begin{verbatim}
 CORRECT, PLANE = x; 
\end{verbatim}

Example 2: no correction, only output of Twiss summary 
\begin{verbatim}
CORRECT, TWISSUM = 1; 
\end{verbatim}

Example 3: correct orbit in horizontal plane, corrector and monitor
output on table 
\begin{verbatim}
CORRECT, PLANE = x, MODE = lsq, CLIST = corr.out, MLIST = mon.out;   
\end{verbatim}

Example 4: correct orbit in horizontal plane, use alignment and scaling
errors, 15\% of orbit correctors faulty
\begin{verbatim}
CORRECT, PLANE = x, MONERROR = 1, MONSCALE = 1, MONON = 0.85; 
\end{verbatim}

Example 5: correct orbit in horizontal plane for a two-beam machine,
using MICADO with no SVD conditioning, zeroing correctors before the
correction. 
\begin{verbatim}
CORRECT, FLAG=ring,
         TWORING, BEAM1TAB=twb1, BEAM2TAB=twb2,
         MODE=micado, COND=0, NCORR=4, ERROR=1E-6, PLANE=x,
         MLIST='mx12.tab', CLIST='cx12.tab', RESOUT=1, CORZERO=1;
\end{verbatim}




%\input{cororbit/co_activate}
\section{USEMONITOR, USEKICK}
\label{sec:activate}
To provide more flexibility with orbit correction two commands are
provided:  

\madbox{
USEMONITOR, \=STATUS=flag, SEQUENCE=sequence, RANGE=range, \\
            \>CLASS=class, PATTERN=regex; \\
\\
USEKICK,    \>STATUS=flag, SEQUENCE=sequence, RANGE=range, \\
            \>CLASS=class, PATTERN=regex;
}

The command \texttt{USEMONITOR} activates or deactivates a
selection of beam position monitor. 
This command affects elements of types \texttt{MONITOR, HMONITOR},
and \texttt{VMONITOR}.    

The command  \texttt{USEKICK} activates or deactivates a selection
of orbit correctors. This command
affects elements of types \texttt{KICKER, HKICKER} and \texttt{VKICKER}. 


Both commands have the same attributes: 
\begin{madlist}
   \ttitem{STATUS} If this flag is true (on), the selected elements
     are activated. Active orbit monitor readings will be
     considered, and active correctors can change their strengths
     in subsequent correction commands. Inactive elements will be
     ignored subsequently.  
   \ttitem{SEQUENCE} The sequence can be specified, otherwise the
     currect sequence is used for this operation.  
   \ttitemn{RANGE, CLASS, PATTERN}: The usual selection commands are
     used to identify the elements for this operation.  
\end{madlist} 

Example:
{\small
\begin{verbatim}
USE,...                                ! set working beam line 
...                                    ! define imperfections 
USEKICK, STATUS = OFF, RANGE = ...;    ! deactivate selected correctors 
USEMONITOR, STATUS = OFF, RANGE = ...; ! deactivate selected monitors   
CORRECT, NCORR = 32;                   ! uses different set of correctors
USEKICK, STATUS = OFF, RANGE = ...;    ! deactivate different set of correctors 
CORRECT, NCORR = 32;                   ! uses different set of correctors
\end{verbatim}
}

%\input{cororbit/co_corrsave}
\section{CSAVE}
\label{sec:csave}
 This section is under construction, options presently only available in
 MADX development version.  

\section{SETCORR}
\label{sec:setcorr}
 This section is under construction, options presently only available in
 MADX development version.  


%\input{cororbit/co_option} 
\section{COPTION}
\label{sec:coption}

\madbox{
COPTION, SEED=integer, PRINT=integer;      
}

In the orbit program monitors can be randomly disabled and the correct
option command specifies different seeds for random values:  

\begin{madlist}
   \ttitem{SEED} Seeds or starts a particular sequence of random values. 
   A valid \texttt{SEED} value is an integer in the range [0...999999999], or
   an expression that evaluates to an integer in the same range 
   (Default: 123456789).\\
   Internally the code takes as the effective seed the value
   \texttt{ABS(seed)\%1.e10} hence normalizing the provided seed to an
   appropriate value in the range irrespective of the provided value. \\
   Note that the seed set with this command is shared with the
   \hyperref[sec:track]{\texttt{TRACK}}  and
   \hyperref[sec:eoption]{\texttt{EOPTION}}  commands.
   See also: \hyperref[subsubsec:random]{Random values}.
     
   \ttitem{PRINT} A flag taking integer values to control the printout. 
     \\ In general: the higher the value the more printout is produced.  
     For \texttt{PRINT=0} no output is produced. 
     The default value is 1 (Correction summary is given). 
\end{madlist}

\textbf{Example:}
\madxmp{
COPTION, SEED=987456321, PRINT=2;
}

Note that the random generator for \mad is taken from \cite{knuth1981}.

%% EOF
