%%\title{Command Format}
%  Changed by: Chris ISELIN, 24-Jan-1997 
%  Changed by: Hans Grote, 25-Sep-2002 
%  Changed by: Ghislain ROY, 27-Jan-2014 

\chapter{Command Format}


\section{Input Statements}

Input for \madx follows in broad lines the
\href{http://cern.ch/mad9}{\madnine}\index{MAD-9} format, \textsl{i.e.} 
free format with commas (,) as separators, although blanks may be used as
separators outside \{...\} enclosures. 

Blank input lines do not affect program execution. 

The input is not case sensitive except for strings enclosed in double
quotes (" "). 

The input file consists of a sequence of statements. 
A statement may occupy any number of input lines. 
Several statements may be placed on the same line.
A statement must be terminated by a semicolon (;).

Several statements may be grouped into a block enclosed inside a \{...\}
enclosure. In this case the terminating semmicolon can be omitted.

\madxmp{if (a $<$ 3) \{ a=b*b; b=2*b+4; \};}
or
\madxmp{if (a $<$ 3) \{ a=b*b; b=2*b+4; \}}
are both valid.

\section{Comments}
When an exclamation mark (\texttt{!}) or double forward slash (\texttt{//}) is
found in the input line, the remaining characters until the end of the
line are treated as a comment and are skipped. 

A comment spreading over multiple lines starts with a string \texttt{/*}
and ends with a string \texttt{*/}.

\section{Commands}
The general format for a command is 
\madbox{
label: keyword \{,attribute\} ;
}
where the  \texttt{\{ \}} are not part of the command and the items
enclosed in \texttt{\{ \}} can be omitted or repeated any number of times. 


A command contains three parts:
\begin{madlist}
   \ttitem{label}\index{label} 
   A label is required for a definition statement. 
   A label gives a name to the stored command.
     
   \ttitem{keyword}\index{keyword} 
   A keyword identifies the action desired.
     
   \ttitem{attributes}\index{attribute} 
   The attributes are normally entered in the
   form \madxmp{attribute-name = attribute-value}
   and serve to define data for the command, where:
   \begin{madlist}
       \ttitem{attribute-name} selects the attribute, and
       \ttitem{attribute-value} provides its value.       
   \end{madlist}
\end{madlist}

If a value is to be assigned to an attribute, the \texttt{attribute-name} is
mandatory.

Whenever an attribute is not explicitly given a value, the default 
\texttt{attribute-value} specified in the command dictionary is assumed. 

%% In the case of attributes taking logical values, it is sufficient to
%% enter the \texttt{attribute-name} only; the attribute is then given the
%% default \texttt{attribute-value} specified in the command dictionary.

\section{Keywords}
\label{subsec:keyword}

A keyword begins with a letter and consists of letters and digits. 

The \madx keywords are protected; attempting to use a protected 
keyword as a label results in a fatal error.

The keyword \texttt{VERSION} has been introduced since version 5.02.07. 
It contains the version number of the \madx release quoted as a decimal.
For example:
\madxmp{
X:> value VERSION;\\
version =              50207 ;
}
This allows testing of the version number of the running \madx process. 
Note that any version prior to version 5.02.07 reports the value 
\texttt{version = 0;}


\section{Attribute Types}
The command attributes can have one of the following types:
\begin{itemize}
  \item String attribute,
  \item Logical attribute,
  \item Integer attribute,
  \item Real attribute,
  \item Expression,
  \item Range selection,
\end{itemize}

Any integer or real attribute can be replaced by a real expression;
expressions are normally deferred\index{deferred} (see
\hyperref[sec:defer]{deferred expression}), except in the   
definition of constants where they are evaluated immediately.

When a command has a \hyperref[sec:label]{label}, \madx keeps this
command in memory. This allows repeated execution of the same command by
simply entering \texttt{EXEC label;}. This construct may be nested. 

%For an exhaustive list of valid declarations of constants or variables
%see \href{declarations.html}{declarations}.


\section{Variable Declarations}
\label{subsec:var_declarations}

In the following, "=" means that the variable on the left receives the
current value of the expression on the right, but does not depend on it
any further, whereas ":=" makes the variable on the left depend on the
expression on the right, \textsl{i.e.} every time the expression
changes, the variable is re-evaluated, except for "const" variables.

\madxmp{
var = expression; \\
var := expression; \\
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
real var = expression;        \>// identical \\
real var := expression;       \>// to above \\
int var = expression;         \>// truncated if expression is real \\
int var := expression; \\
const var = expression; \\
const var := expression; \\
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
const real var = expression;     \>// identical \\
const real var := expression;    \>// to above \\
const int var = expression;      \>// truncated if expression is real \\
const int var := expression;
}


\section{Identifiers or Labels}
\label{sec:label}
A label begins with a letter, followed by up to fifteen letters, digits,
decimal points (.), or underscores (\_). Characters beyond the sixteenth
are dropped, but should be avoided, and the resulting sequence must be
unique. 

A label may refer to a keyword, an element, a beam-line, a sequence,
etc.   

The \madx keywords are protected; using one of them as a label
results in a fatal error.  


%\input{Introduction/attribute}
\section{Command Attributes}
\label{sec:attribute}

\begin{itemize}
  \item A name or string attribute refers to an object, or a string. 
  \item A logical attribute selects or deselects an option. 
  \item An integer attribute defines a value stored as integer data.
  \item A real attribute defines a value stored as double precision data.  
  \item A real expression defines a datum for a command, it may be
    varied in matching. An expression is built of a combination of
    operator and operand. 
  \item A constraint specifies a matching constraint. 
  \item A variable name selects a variable to be matched. 
\end{itemize}


%\input{Introduction/name}
\section{Name or String Attributes}
\label{sec:name}\label{sec:string}
A name or string attribute often selects one of a set of options: 
\madxmp{
xxxxxxxxxxxxxxxxxxxx\= \kill
USE, PERIOD=lhc;    \>// expand the LHC sequence
}

It may also refer to a user-defined object: 
\madxmp{
xxxxxxxxxxxxxxxxxxxxxxx\= \kill
TWISS, FILE=optics;    \>// specifies the name of the OPTICS output file
}

It may also define a string: 
\madxmp{
TITLE, "LHC version 6.2";
}

The case of letters is only significant if a string is enclosed in
quotes, otherwise all characters are converted to lowercase at
reading. 

On the other hand, strings that do not contain blanks do not
need to be enclosed in quotes.

Example:
\madxmp{
CALL, FILE = "my.file"; \\
CALL, FILE = my.file; \\
CALL, FILE = MY.FILE; \\
CALL, FILE = "MY.FILE"; \\
CALL, FILE = 'MY.FILE'; 
} 
In the first three cases, \madx will try to read a file named my.file, in the
last two it will try to read the file named MY.FILE.  


%\section{String Attributes}

A string attribute makes alphanumeric information available, e.g. a
title or a file name. It can contain any characters, enclosed in single
(') or double (") quotes, except for quotes of the type used as its
delimiter.  

Examples: 
\madxmp{
TITLE, 'This is a title for the program run "test"'; \\
\\
SYSTEM, "ln -fns some-lengthy-directory-name local-link";
}


\section{Logical Attributes}
\label{sec:logical}
Many commands in \madx require the setting of logical values (flags) to
represent the on/off state of an option. A logical value "flag" can be
set in two ways:  
\madxmp{flag | flag = true}

It can be reset by: 
\madxmp{-flag | flag = false}

Example: 
\madxmp{
OPTION, -ECHO;  // switch off copying the input to the standard output
}

The default for a logical flag is normally \texttt{false}, but can be found e.g. for options by the command  
\madxmp{HELP, option;}


\section{Integer Attributes}
\label{sec:integer}
An integer attribute usually denotes a count. Example: 
\madxmp{myline: LINE = ( -3 * (a,b,c) );}

In this case, a litteral integer is requested; however, in the following 
\madxmp{rfc: RFCAVITY, HARMON = 12345;}
or 
\madxmp{rfc: RFCAVITY, HARMON = num;}
"num" may be an integer variable, a real variable, or an expression.  In
the two latter cases, the value is truncated. 


\section{Real Attributes}
\label{sec:real}
Most attributes are of type REAL and are treated internally as double
precision values. They may be set by integer values, real values,  or
expressions. 

Example:  
\madxmp{
xxxxxxx\= \kill
ddd:   \>drift, l = 1.2345; \\
dddd:  \>drift, l = ddd-\textgreater l-0.3;
}


%\input{Introduction/expression}
%  Changed by: Ghislain Roy, 12-Dec-2013: updated physicial constants
%  to PDG2012 values; very old values were pre PDG2010.
\section{Real Expressions}
\label{sec:expression}
To facilitate the definition of interdependent quantities, any real
value and integer value can be entered as an arithmetic expression. When
a value used in an expression is redefined by the user or changed in a
matching process, the expression is reevaluated. Expression definitions
may be entered in any order. \madx evaluates them in the correct order
before it performs any computation. At evaluation time all operands used
must have values assigned.  

An expression is built from a combination of
\hyperref[subsec:operator]{operator} and \hyperref[subsec:operand]{operand}, and it
may contain \hyperref[subsubsec:random]{random generators}.   

\subsection{Operators in Arithmetic Expressions}
\label{subsec:operator}
An expression can be formed using the following operators: 

\subsubsection{Arithmetic operators}
\begin{madlist}
  \ttitem{+} Addition, 
  \ttitem{-} Subtraction, 
  \ttitem{*} Multiplication, 
  \ttitem{/} Division, 
  \ttitem{\^\ } Exponentiation. 
\end{madlist}

\subsubsection{Ordinary Functions}
\label{subsubsec:function}

\begin{madlist}
  \ttitem{SQRT(x)} square root, 
  \ttitem{LOG(x)}  natural logarithm, 
  \ttitem{LOG10(x)} logarithm base 10, 
  \ttitem{EXP(x)} exponential, 
  \ttitem{SIN(x)} trigonometric sine, 
  \ttitem{COS(x)} trigonometric cosine, 
  \ttitem{TAN(x)} trigonometric tangent, 
  \ttitem{ASIN(x)} arc sine, 
  \ttitem{ACOS(x)} arc cosine, 
  \ttitem{ATAN(x)} arc tangent, 
  \ttitem{SINH(x)} hyperbolic sine, 
  \ttitem{COSH(x)} hyperbolic cosine, 
  \ttitem{TANH(x)} hyperbolic tangent, 
  \ttitem{SINC(x)} cardinal sine function,
  \ttitem{ABS(x)} absolute value, 
  \ttitem{ERF(x)} Gauss error,
  \ttitem{ERFC(x)} complementary error,
  \ttitem{FLOOR(x)} floor, largest previous integer,
  \ttitem{CEIL(x)} ceiling, smallest next integer,
  \ttitem{ROUND(x)} round, closest integer;
  \ttitem{FRAC(x)} fractional part of number;
\end{madlist}

\subsubsection{Random Number Generators}
\label{subsubsec:random}
\begin{madlist}
  \ttitem{RANF()} random number, uniformly distributed in [0,1], 
  \ttitem{GAUSS()} random number, gaussian distribution with unit
  standard deviation,  
  \ttitem{TGAUSS(x)} random number, gaussian distribution with unit
  standard deviation, truncated at x standard deviations;  
\end{madlist}

in the above cases, "x" can be any expression, i.e. contain other
functions, variable or constant expressions. To initialize the \madx
random generator use the \hyperref[sec:eoption]{\texttt{EOPTION}} or  \hyperref[sec:coption]{\texttt{COPTION}} commands. Note that the command \hyperref[sec:track]{\texttt{TRACK}} with \texttt{QUANTUM} attribute set to true may set or reset the seed too. 

The command  \hyperref[sec:option]{\texttt{OPTION}} has two attributes to control the new random generator of MAD-X. The attribute \texttt{RAND} can take the value \texttt{default} (for backward compatibility) or \texttt{best} for the new generator. The latter supports 10 multiple independent generators numbered from 1 to 10, and can be selected with the attribute \texttt{RANDID=id} in combination to \texttt{RAND}. If \texttt{RANDID} is not specified and \texttt{RAND=best}, then the last \texttt{id} set is used (init: 1). The user can switch between the 10 independent generators at will to use different streams for the next generated random numbers. When a new seed is set through to command \hyperref[sec:eoption]{\texttt{EOPTION}},  \hyperref[sec:coption]{\texttt{COPTION}} or \hyperref[sec:track]{\texttt{TRACK}}, it will only affect the currently selected stream. The special initial seed for the \texttt{best} generator can be restored by using the value zero instead of the commands default 123456789. 

The following example displays the first 10 random numbers of the 10 streams of the \texttt{best} generator, using the default initial seed.

\begin{verbatim}
option, -info;
i = 1;
while (i <= 10) {
  option, rand=best, randid=i;
  j = 1;
  while(j <= 10) {
    value, i, j, ranf();
    j = j+1;
  }
  i = i+1;
}
\end{verbatim}

\subsubsection{Table Access Functions}
\label{subsubsec:table}

\begin{madlist}
  \ttitem{TABLE(x,z)} identical to TABLE(x,z,1); example: table(summ,q1) returns the hor. tune of
  the Twiss summary table "summ".

  \ttitem{TABLE(x,y,z)} accesses value of the named table column "z"
  for element "y" (first table row with that name) of table "x";
  example: table(twiss,mb.12,betx) returns the beta\_x at
  element mb.12 from the Twiss table "twiss".   When the element
  is called with its proper name, as in the example above, the
  value is returned at the first occurrence of the element of
  this name. If the value is needed at an occurrence number: NNN,
  then "[NNN]" has to be appended to the name, e.g. in the above
  example one obtains the betx of the 23th occurrence of the
  element "mb.12" by changing the example to:
  "table(twiss,mb.12[23],betx)". Mind you that the old, but
  little known, form: "table(twiss,mb.12-\textgreater 23,betx)"
  continues to work. Lastly, if NNN exceeds the maximum
  occurrence number the return value is arbitrarily small.  

  \ttitem{TABLE(x,z,N\_row)} accesses the value of the named table
    column  "z" at the "N\_row" number of rows of table "x" (row
    numbers start at  1); example: table(twiss,betx,370) returns
    the beta\_x at row number  "370" of the Twiss table
    "twiss". The return value is zero if "N\_row"  is outside of
    the allowed range.

  \ttitem{TABINDEX(x,z,key)} identical to TABINDEX(x,z,1,key).  

  \ttitem{TABINDEX(x,z,N\_row,key)} returns the index of the first occurrence of a string starting by "key" in table column "z" (must be a column of strings) starting at the "N\_row" number of rows of table "x"; example: tabindex(survey,name,102,bpm) returns the row number of the first element found starting from row 102 with a "name" beginning by "bpm" in the table "survey".

\ttitem{TABSTRING(x,z,N\_row)} accesses value of the named table
 column "z" (must be a column of strings) at the "N\_row" number of rows of table "x";
example: tabstring(survey,name,10) returns the name of
the 10th element of the survey table "survey". Row count starts at 1. The
function can be used in any context where a string appears; in case
there is no match, it returns "\_void\_".  
\end{madlist}

Note that "N\_row" has to be a number and not a  variable. However, the
\hyperref[sec:macro]{Macro} facility in \madx allows
the use of a variable instead.   

\begin{verbatim}
gettab(tblname, colname, rowidx) : macro = {
  gettab.val = table(tblname, colname, rowidx) ;
}
getidx(tblname, colname, rowidx, keyname) : macro = {
  getidx.val = tabindex(tblname, colname, rowidx, keyname) ;
}
row = 10 ;
exec, gettab(twiss,betx,$row) ;
exec, getidx(twiss,name,$row,qf) ;
value, gettab.val, getidx.val ;
\end{verbatim}

Beware: 
\begin{itemize}
	\item  Unnamed Drifts are not included in the table name
	database, so as to speed up the search for "real"
	elements. Therefore, those  unnamed drifts cannot be
	found. However, named drifts can be searched for.  
	\item  Due to the importance of finding elements in the table
	for a proper functioning of the \madx runs, the programs
	throws a "fatal\_error" if an element cannot be found in the
	table.   
\end{itemize}

A typical example could look like this, in fact the square root of betx
(user defined variable myvar) is added to the twiss table:  
\begin{verbatim}
myvar := sqrt(table(twiss,betx));
select, flag=twiss, column=name, s, myvar, betx;
twiss, file;
\end{verbatim}

Or another arbitrary test case of adding k1l taken from the Twiss table: 

Define macro: 
\begin{verbatim}
mymacro(xx,yy,zz): macro = {myval = table(xx,yy,zz);};
\end{verbatim}

Use macro in loop: 
\begin{verbatim}
i = 0;
incval = 0;
while (i < 100) {
  value, i;
  exec, mymacro(twiss,k1l,$i);
  incval = incval + myval;
  value, i, myval, incval;                
  i = i + 1;
}
\end{verbatim}

\subsection{Operands in Arithmetic Expressions}
\label{subsec:operand}
An expression may contain the following operands:  

\subsubsection{Literal Constants} 
Numerical values are entered like FORTRAN constants. Real values are
accepted in INTEGER or REAL format. The use of a decimal exponent,
marked by the letter D or E, is permitted.  

Examples: 
\madxmp{
1, 10.35, 5E3, 314.1592E-2
}

\subsubsection{Symbolic constants}
\label{subsubsec:symbolic_const}
\madx recognizes some \hyperref[tab-constants]{mathematical and physical
constants}. Their names must not be used for user-defined labels.  

Additional symbolic constants may be defined to simplify their repeated
use in statements and expressions.  
\madbox{
CONST name=constant-expression;
}
defines a real constant with the name given. An existing symbolic
constant can be redefined, but it cannot change in a matching procedure.  

Example: 
\madxmp{CONST IN = 0.0254;}

A number of predefined symbolic constants exist in \madx and can be used
in expressions. The values of physical constants are regularly updated
to reflect the values published by the Particle Data Group \cite{PDG2014}. 

\begin{table}[ht]
  \caption{Predefined Symbolic Constants in \madx}
  \label{tab-constants}
\vspace{1ex}
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{\madx name} & \textbf{symbol} & \textbf{value} & \textbf{unit} \\ 
\hline
PI & $\pi$ & 4 * atan(1) & 1 \\ 
TWOPI & $2\pi$ & 2 * PI & 1 \\ 
DEGRAD & $180/\pi$ & 180 / PI  & deg/rad \\ 
RADDEG & $\pi/180$ & PI / 180 & rad/deg \\ 
E & e & exp(1) & 1 \\ 
EMASS & $m_e$ & $0.51099895000\times 10^{-3}$& GeV \\ 
PMASS & $m_p$ & 0.93827208816 & GeV \\ 
NMASS & $u$ & 0.93956542052 & GeV \\
MUMASS & $m_\mu$ & 0.1056583715 & GeV \\ 
CLIGHT & $c$ & $2.99792458\times 10^{8}$ & m/s \\ 
QELECT & $e$ & $1.602176634\times 10^{-19}$ & A.s \\ 
HBAR & $\hbar$ & $6.582119569\times 10^{-25}$ & MeV.s\\
ERAD & $r_e$ & $2.8179403262\times 10^{-15}$ & m\\
PRAD & $r_e (m_e / m_p)$ & ERAD*EMASS/PMASS & m \\
\hline
\end{tabular}
\end{table}

Note that the \texttt{NMASS} constant in \madx is the unified atomic mass unit and 
not the neutron mass.

\subsubsection{Parameter labels} 
Often a set of numerical values depends on a common variable
parameter. Such a parameter must be defined as a global parameter. A
global parameter always has a current value; however, this value may be
re-evaluated or not, depending on the parameter definition:  
\begin{verbatim}
x = a;
\end{verbatim} 
\texttt{x} is set to the current value of \texttt{a} and not changed,
even if \texttt{a} changes. This makes assignments such as  
\begin{verbatim}
x = x + 1;
\end{verbatim} 
perfectly valid (this replaces the deprecated \texttt{SET} instruction).

The definition of the deferred expression  
\begin{verbatim}
x := a;
\end{verbatim} 
assign the current value of \texttt{a} to \texttt{x} every time
\texttt{x} is used, \textsl{i.e.} it is re-evaluated using the latest
value of \texttt{a}; therefore,   
\begin{verbatim}
x := x + 1;
\end{verbatim} 
results in an infinite loop (!) when \texttt{x} is used, which results
in abnormnal termination of \madx. 

Of course, the following definitions are equivalent:  
\begin{verbatim}
x = 0.1;
x := 0.1;
\end{verbatim}

When such a parameter is used in an expression, \madx uses the current
value of the parameter if the expression is deferred:  

Example: 
\madxmp{
x := 1.0; \\
d1: drift, l =  x; \\
d2: drift, l := 2.0 - x; \\
}
When the value of \texttt{x} is changed, the length of the drift
\texttt{d1} remains unchanged, while the length of the drift \texttt{d2}
is recalculated.

\subsubsection{Element or Command Attributes} 
In arithmetic expressions the attributes of physical elements or
commands can occur as operands. They are named respectively by  
\madbox{
element-name->attribute-name\\
command-name->attribute-name
}

Values are assigned to attributes in element definitions or commands. 

Example: 
\madxmp{
xxxxxxxx\= \kill
D1:     \>DRIFT, L= 1.0; \\
D2:     \>DRIFT, L= 2.0 - D1->L;
}
\texttt{D1-\textgreater L} refers to the length \texttt{L} of the drift
space \texttt{D1}.   

\subsection{Expressions and Random Values}
\label{subsec:expr_rnd}
The definition of random machine imperfections requires evaluation of
expressions containing random functions. These are evaluated like any
other expression when the expression is used, i.e. only once if a "="
assignment refers to it, or every time if the assignment is ":="; this
latter case is used by the error generation routines.  

Example: 
\madxmp{
a := 3*ranf();
}
Every time \texttt{a} is used, it gets a random value assigned from a uniform
distribution between 0 and 3.  

\madxmp{
error: ealign, range, dx:=sigma*gauss();
}
All elements in range are assigned independent random displacements
sampled from a Gaussian distribution with standard deviation sigma.  



%\input{Introduction/constraint}
%\section{Constraints}
%\label{sec:constraints}
\section{Logical Expressions}
\label{sec:logicalexpr}

In matching it is desired to specify equality constraints, as well as
lower and upper limits for a quantity. \madx accepts the following forms
of constraints:  

%% \begin{verbatim}
%% ! equality constraint:
%% name = expression

%% ! upper limit:
%% name < expression

%% ! lower limit:
%% name > expression

%% ! both upper and lower limit for the same name:
%% name < expression, name > expression
%% \end{verbatim}

\madbox{
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
name = expression                    \>! equality constraint \\
name < expression                    \>! upper limit \\
name > expression                    \>! lower limit \\
name < expression, name > expression \>! both upper and lower limit \\
                                     \>! for the same name
}

%\input{Introduction/variable}
\section{Variable Names}
\label{sec:variable}
A variable name can have one of the formats: 
\madbox{
 parameter-name \\
 element-name->attribute-name \\
 command-name->attribute-name \\
 beam\%sequence-name->attribute-name \\
 table(table-name,..,..)
}

The first format refers to the value of the global parameter
\texttt{parameter-name}.

The second and third formats refer to the real attribute
\texttt{attribute-name} of the element  \texttt{element-name}, or the
command \texttt{command-name}.  

The fourth format is specific
to beams belonging to a particular sequence (for details see
\hyperref[chap:sequence]{sequences} and \hyperref[chap:beam]{beams}). 

The fifth format allows extraction of variables from existing tables, as specified in
\hyperref[chap:tables]{table access}.  



%\input{Introduction/wildcard}
\section{Regular Expressions}
\label{sec:regex}
Some commands allow selection of items via "regular expression"
strings. Such a pattern string \textbf{must} be enclosed in single or
double quotes. \madx follows regexp (Unix regular expression patterns)
for matching. The following features are implemented:  

A "search string" below is the string containing the pattern, a "target
string" is the string being searched for a possible match with the
pattern. 

\begin{madlist}
   \ttitem{"\textasciicircum"} at the start of the search string: Match
     following search string at the start of the target string;
     otherwise the search string can start anywhere in the target
     string. To search for a  genuine "\textasciicircum" anywhere, use
     "$\backslash$\textasciicircum".  
   \ttitem{"\$"} at the end of the search string: Match preceding search
     string at the end of the target string; otherwise the search string
     can end anywhere in the target string. To search for a  genuine
     "\$" anywhere, use "$\backslash$\$".
   \ttitem{"."} stands for an arbitrary character; to search for a genuine
     ".", use "$\backslash$."
   \ttitem{"[xyz]"} stands for one character belonging to the string
     contained in brackets (example: "[abc]" means one of a, b, c).  
   \ttitem{"[a-ex-z]"} stands for ranges of characters (example:
     "[a-zA-Z]" means any letter).  
   \ttitem{"[\textasciicircum xyz]"} (i.e. a "\textasciicircum" as first
     character in a square bracket) stands for exclusion of all
     characters in the list, i.e. "[\textasciicircum a-z]" means "any
     character but a lower case letter". 
   \ttitem{"*"} allows zero or more repetitions of the preceding
     character, either specified directly, or from a list. (examples:
     "a*" means zero or more occurrences of "a",  "[A-Z]*" means zero or
     more upper-case letters).  
   \ttitem{"$\backslash$c"}  (e.g. "$\backslash$.") removes the special
     meaning of character c.  
\end{madlist}

All other characters stand for themselves. 


Example: 
\madxmp{
SELECT, FLAG=twiss, PATTERN="\textasciicircum d..\$" ; \\
SELECT, FLAG=twiss, PATTERN="\textasciicircum k.*qd.*$\backslash$.r1\$" ;
}
 
The first command selects all elements whose names have exactly three
characters and begin with the letter "D". The second command selects
elements beginning with the letter "K", containing the string "QD", and
ending with the string ".R1". The two occurrences of ".*" each stand for
an arbitrary number (including zero) of any character, and the
occurrence "$\backslash$." stands for a literal period.  


%%\title{Parameters}
%  Changed by: Hans Grote, 17-Jun-2002 
%\chapter{Parameter Statements}
%\label{chap:parameter}

\section{Relations between Variable Parameters}
\label{sec:relations}\label{sec:defer}
A relation is established between variables by one of two statements 
\madbox{
xxxxxxxxxxxxxxxx\= \kill
parameter-name \>=  expression; \\
parameter-name :\>=  expression;
}
The first form evaluates the expression on the right immediately and
assigns its value to the parameter. It is an immediate assignment.

The second form assigns the value by evaluating  the expression on the
right every time the parameter is actually used. It is a deferred
assignment.  

This mechanism holds as well for element parameters that can be defined
with either immediate or deferred assignments.

Attention! If you want
to modify e.g. the strength of a quadrupole later (e.g. in a match,  or
by entering a new value for a parameter on which it depends) then the
defition has to be  
\madxmp{QD: QUADRUPOLE, K1 := ak1;}
and not 
\madxmp{QD: QUADRUPOLE, K1 = ak1;}
In the latter case, K1 will be set to the current value of ak1 at the
time of declaration, and will not change when ak1 later changes.  

Parameters that have not yet been defined at time of evaluation have a
zero value. 

Example: 
\madxmp{
gev = 100; \\
BEAM, ENERGY=gev;
}

The parameter on the left may appear on the right as well in the
computer science form of assignments: 
\madxmp{x = x+1;}
increases the value of x by 1. \\
%As a result, the SET statement of \madeight is no longer necessary and
%is not implemented in \madx.   


Successive definitions are allowed in the first form of relations or 
immediate assignments:
\madxmp{
a = b + 2; \\
b = a * b;
}
But circular definitions in the second form of relations, or
deferred assignments, are forbidden: 
\madxmp{
a := b + 2; \\
b := a * b;
}
results in an error.

%\href{http://www.cern.ch/Hans.Grote/hansg_sign.html}{hansg} 11.9.2000



%% EOF

