%%\title{Matching Module}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 
 
\chapter{Matching Module}
\label{chap:match}

Before a match operation at least one sequence must be selected by means
of a \hyperref[sec:use]{\texttt{USE}} command. Matching is 
initiated by the \hyperref[sec:match]{\texttt{MATCH}} command. The matching
module can act on more than one sequence simultaneously by  specifying
more than one sequence when initiating the matching mode.  
From this command to the corresponding
\hyperref[sec:endmatch]{\texttt{ENDMATCH}} command \madx accepts the
matching commands listed below. For a mathematical description of the
minimisation procedures see \cite{MINUIT}. 

In particular one may do the following: 
\begin{itemize}
 \item Define the sequence(s) the matching module will work on 
 \item Set initial conditions for transfer line matching 
 \item Define constraints 
 \item Define the parameters to be varied 
 \item Match by different methods. 
\end{itemize}

The matching commands are described in detail below. Some other commands
can also be issued during matching.  

\begin{itemize}
\item Enter and Leave Matching Mode
  \begin{itemize}
  \item \hyperref[sec:match]{\texttt{MATCH}}: Initiate Matching Mode
  \item \hyperref[sec:endmatch]{\texttt{ENDMATCH}}: Leave Matching Mode
  \end{itemize}
\item Define Variable Parameter
  \begin{itemize}
  \item \hyperref[sec:vary]{\texttt{VARY}}: Set Parameter to Vary
  \end{itemize}
\item Define Constraint
  \begin{itemize}
  \item \hyperref[sec:constraint]{\texttt{CONSTRAINT}}: Simple Constraint 
  \item \hyperref[sec:userconstraint]{\texttt{CONSTRAINT}}: User Defined Variables 
  \item \hyperref[sec:weight]{\texttt{WEIGHT}}: Matching Weights
  \item \hyperref[sec:global]{\texttt{GLOBAL}}: Global Constraints
  \item \hyperref[sec:gweight]{\texttt{GWEIGHT}}: Weights for Global Constraints
  \end{itemize}
\item \hyperref[sec:match-methods]{Matching Methods}
  \begin{itemize}
  \item \hyperref[subsec:match-lmdif]{\texttt{LMDIF}}: Fast Gradient Minimisation
  \item \hyperref[subsec:match-migrad]{\texttt{MIGRAD}}: Gradient Minimisation
  \item \hyperref[subsec:match-simplex]{\texttt{SIMPLEX}}: Simplex Minimisation
  \item \hyperref[subsec:match-jacobian]{\texttt{JACOBIAN}}: Newton Minimisation
  \end{itemize}
\item Expression Matching with \hyperref[sec:use-macro]{\texttt{USE\_MACRO}}
\end{itemize}

%\line(1,0){300}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002; 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006. 

% add other files to the end of this file


%\input{match/match_main}
%%\title{MATCH / ENDMATCH}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 
%  Changed by: Riccardo de Maria, 9-Jan-2008 


%% Before matching at least one
%% \href{../Introduction/sequence.html}{SEQUENCE} must be selected by means
%% of a \href{../control/general.html#use}{USE} command. The matching
%% module can act on more than one sequence simultaneously by specifying
%% more than one sequence in the \texttt{MATCH} command.


\section{MATCH \ldots ENDMATCH}
\label{sec:match}\label{sec:endmatch}

The \texttt{MATCH} command is used for matching either a \hyperref[sec:cell]{periodic cell}
or an \hyperref[sec:insertion]{insertion} to another part of the machine. 

Both matching modes are initiated by the \texttt{MATCH} command using
one of several forms outlined below. 

The \texttt{ENDMATCH} command terminates the matching section and deletes
all tables related to the matching run. 
\madbox{
ENDMATCH;
}


\section{Cell Matching}
\label{sec:cell}

The matching of a periodic cell is initiated by a \texttt{MATCH} command of
the form:

\madbox{
MATCH, \=SEQUENCE='name1', 'name2', \ldots , 'name-n', \\
       \>SLOW=logical;
}

In this first mode the matching routine is initiated only with one
attribute specifying the sequence(s) the matching module will work
on. In this matching mode the periodicity of the optics functions is
enforced at the beginning and end of the selected range. 
\madxmp{MATCH, SEQUENCE='name1', 'name2', \ldots ,'name-n';}
    


\section{Insertion Matching}
\label{sec:insertion}
The matching of an insertion to another part of the machine is initiated
by a \texttt{MATCH} command taking one of the following forms:

\madbox{
MATCH, \=SEQUENCE= 'name1', 'name2', \ldots , 'name-n',\\
       \>BETA0= 'beta01', 'beta02', \ldots , 'beta0n', \\
       \>SLOW=logical;
}

or

\madbox{
MATCH, \=SEQUENCE='seqname', \\ 
       \>BETX=real, ALFX=real, MUX=real, \\
       \>BETY=real, ALFY=real, MUY=real, \\
       \>X=real, PX=real, Y=real, PY=real, \\
       \>DX=real, DY=real, DPX=real, DPY=real, \\
       \>DELTAP=real, SLOW=logical;
}

In the second mode, called insertion matching, the matching routine is
initiated with two attributes: one specifying the sequence(s) the
matching module will work on and one specifying the initial conditions
of the optic functions for each sequence. In this case the initial
values are assumed as exact. 
    
\textbf{Specification of Initial Values:}
The initial values of the optical
functions  for the insertion matching can either be specified in form of
a \hyperref[sec:savebeta]{\texttt{SAVEBETA}} command or by
explicitly stating the individual optic function values after the
\texttt{MATCH} command. The two options can be entered as         
\madxmp{
  MATCH, \=SEQUENCE= 'name1', 'name2',.., 'name-n',\\
         \>BETA0= 'beta01', 'beta02',..., 'beta0n';
}
or
\madxmp{
  MATCH, \=SEQUENCE='seqname', \\ 
         \>BETX=real, ALFX=real, MUX=real, \\
         \>BETY=real, ALFY=real, MUY=real, \\
         \>X=real, PX=real, Y=real, PY=real, \\
         \>DX=real, DY=real, DPX=real, DPY=real, \\
         \>DELTAP=real;
}

\textbf{Example 1:}   
\begin{verbatim}
CELL: SEQUENCE=(...) ;
INSERT: SEQUENCE=(...) ;
USE, PERIOD=cell;
SAVEBETA, LABEL=bini, PLACE=#e;
TWISS, SEQUENCE=cell;
USE, PERIOD=insert;
MATCH, SEQUENCE=insert, BETA0=bini;
CONSTRAINT, SEQUENCE=insert, RANGE=#e, MUX=9.345, MUY=9.876;
\end{verbatim}

This matches the sequence 'INSERT' with initial conditions to a new
phase advance. The initial conditions are given by the periodic solution
for the sequence CELL1. 
    
\textbf{Example 2:}	
\begin{verbatim}
USE, PERIOD=insert;
MATCH, SEQUENCE=insert;
CONSTRAINT, SEQUENCE=insert, RANGE=#e, MUX=9.345, MUY=9.876;
\end{verbatim}
This matches the beam line 'INSERT' with periodic boundary conditions to
a new phase advance. 

The initial conditions can also be transmitted by a combination of a
\hyperref[sec:savebeta]{\texttt{SAVEBETA}} command and explicit
optic function specifications: 
\begin{verbatim}
USE, cell1;
SAVEBETA, LABEL=bini, PLACE=#E;
TWISS, SEQUENCE=cell1;
USE, PERIOD=line1;
MATCH, SEQUENCE=line1, BETA0=bini, MUX=1.234, MUY=4.567;
\end{verbatim}

This example transmits all values of the \texttt{SAVEBETA} array 'bini' as
initial values to the \texttt{MATCH} command and overrides the initial phase
values by the given values.

An additional \hyperref[sec:constraint]{\texttt{CONSTRAINT}} may be
imposed in other places, i.e. intermediate or end values of the optics
functions at the transition point.  
 
\section{More than one active sequence}

The matching module can act on more than one sequence simultaneously by
specifying more than one sequence after the \texttt{MATCH} command: 
\madxmp{
MATCH, SEQUENCE=line1, CELL1, BETA0=bini1, bini2;
}
This example initiates the matching mode for the 'LINE1' and the 'CELL1'
sequence. The \hyperref[chap:twiss]{Twiss} functions of the
two sequences are calculated with fixed initial conditions. 

The \texttt{SAVEBETA} array 'bini1' is used for calculating the optics
functions of sequence 'LINE1' and the \texttt{SAVEBETA} array 'bini2'
for calculating the optics functions of sequence 'CELL1'. Without the
initial conditions the matching module will work in the
\hyperref[sec:cell]{\texttt{CELL}} mode. 
 
\section{SLOW attribute}

The \texttt{SLOW} logical flag enforces the old and slow matching procedure
which allows to use the special columns \texttt{mvar1, ..., mvar4}, if they
are added to the twiss table.

Recently a number of parameter, like \texttt{RE56}, have been added to the
list of parameters that can be matched. 

Nevertheless, some parameters might only be available when using the
\texttt{SLOW} attribute.
 
\section{Useful TWISS attributes}

Some of the attributes of the \hyperref[chap:twiss]{\texttt{TWISS}} command
can be used in the \texttt{MATCH} command and are transmitted to the
\texttt{TWISS} command when it is internally invoked during the matching process. 

The main \texttt{TWISS} attributes that can be used also in the \texttt{MATCH}
command are:

\begin{madlist}
  \ttitem{RMATRIX} If this flag is used the one-turn map at the location of every
  element is calculated and prepared for storage in the TWISS table.
 
  Target values for the matrix elements at certain positions in the
  sequence can be specified with the help of the
  \hyperref[sec:constraint]{\texttt{CONSTRAINT}} command and the keywords:\\
  \textbf{RE, RE11...RE16...RE61...RE66}, where \textbf{REij} refers to
  the "ij" matrix component.
  

  \ttitem{CHROM} This logical flag sets the matching process to transmit
  the \texttt{CHROM} attribute to the \hyperref[chap:twiss]{\texttt{TWISS}}
  command when it is invoked, enforcing the calculation of chromatic
  functions and synchrotron radiation integrals, and the alternative
  calculation of chromaticities as documented in
  \hyperref[chap:twiss]{\texttt{TWISS}}.

  If this flag is used the chromatic functions at the location of
  every element are calculated and prepared for storage in the TWISS
  table. 
  
  Target values for the chromatic functions at certain positions in the
  sequence can be specified with the help of the
  \hyperref[sec:constraint]{\texttt{CONSTRAINT}} command and the
  repective keywords 
  \hyperref[subsec:tables-chrom]{\texttt{WX, PHIX, WY, PHIY,...}} for
  the chromatic functions. 
\end{madlist}


\textbf{Examples:}

\begin{verbatim}
MATCH, RMATRIX, SEQUENCE='name', BETA0='beta-block-name';
CONSTRAINT, SEQUENCE=insert, RANGE=#e, RE11=-2.808, RE22=2.748;
VARY, NAME=kqf, STEP=1.0e-6;
VARY, NAME=kqd, STEP=1.0e-6;
\end{verbatim}

This matches the sequence 'name' with initial conditions to new values
for the matrix elements \texttt{RE11} and \texttt{RE22} by varying the
strength of the main quadrupole circuits.


%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, October, 2003;
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, January, 2008.


%\input{match/match_vary}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 19-Jun-2002 
 
\section{VARY}
\label{sec:vary}
A parameter to be varied is specified by the command 

\madbox{
VARY, \=NAME=variable, STEP=real, LOWER=real, UPPER=real \\
      \>SLOPE=integer, OPT=real;
}

It has six attributes: 
\begin{madlist}
  \ttitem{NAME} The name of the parameter or attribute to be varied,  

  \ttitem{STEP} The approximate initial step size for varying the
  parameter. If the step is not entered, \madx tries to find a
  reasonable step, but this may not always work.  

  \ttitem{LOWER} Lower limit for the parameter (optional), 
  \ttitem{UPPER} Upper limit for the parameter (optional). 

  \ttitem{SLOPE} allowed change rate (optional, available only using
  \hyperref[subsec:match-jacobian]{JACOBIAN} routine). Limit the
  parameter to increase (\texttt{SLOPE=1}) or decrease (\texttt{SLOPE=-1}) only.

  \ttitem{OPT} optimal value for the parameter (optional, available
  only using \hyperref[subsec:match-jacobian]{JACOBIAN} routine).  
\end{madlist}

\textbf{Examples:}
\madxmp{
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
VARY, NAME=PAR1, STEP=1.0E-4;          \>! vary global parameter PAR1 \\
VARY, NAME=QL11->K1, STEP=1.0E-6;      \>! vary attribute K1 of the QL11 \\ 
VARY, NAME=Q15->K1, STEP=0.0001, LOWER=0.0, UPPER=0.08; ! vary with limits
}

If the upper limit is smaller than the lower limit, the two limits are
interchanged. If the current value is outside the range defined by the
limits, it is brought back to range. 

If a parameter comes outside the
limits during the matching process the matching module resets the
parameter to a value inside the limits and informs the user with a
message. If such a 'rescaling' occurs more than 20 times the matching
process terminates. The user should either eliminate the corresponding
parameters from the list of varied parameters or change the
corresponding upper and lower limits before restarting the matching
process. 

After a matching operation all varied attributes retain their
value after the last successful matching iteration. Using
\hyperref[subsec:match-jacobian]{JACOBIAN} routine, \texttt{STRATEGY=3}, in case
the number of parameters is greater than the number of constraint, if a
parameter comes outside the limits, it is excluded automatically from
the set of variables and a new solution is searched.  

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002. 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006.



%\input{match/match_con}
%%\title{Matching Constraints}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 25-Sep-2002 
% Changed by: Ghislain Roy, 11 Dec 2013: removed duplicate paragraph in user defined matching constraint 

\section{CONSTRAINT}
\label{sec:constraint}

Simple constraints are imposed by the \texttt{CONSTRAINT} command. 
The \texttt{CONSTRAINT} command has three attributes:   
\begin{itemize}
 \item  the \texttt{SEQUENCE} entry specifies the sequence for which the
   constraint applies.  
 \item  the RANGE entry specifies the position where the
   constraint must be satisfied. The \texttt{RANGE} can either be the name
   of a single element in the sequence or a range between two
   elements. In the later case the two element names must be
   separated by a '/': \texttt{RANGE=name1/name2}  
 \item the optics functions to be constrained. 
\end{itemize} 

The optic functions can be constrained in four different ways: 
\begin{enumerate}
 \item lower limit: \texttt{BETX $>$ value}
 \item upper limit: \texttt{BETX $<$ value}
 \item lower and upper limits: \texttt{BETX $<$ value1, 
   BETX $>$ value2} 
 \item target value: \texttt{BETX=value}
\end{enumerate} 

In case one element is affected by more than one constraint command the
last CONSTRAINT will be chosen. For example, one can specify the
maximum acceptable beta function over a range of the sequence and
specify the target beta function for one element that lies inside this
range. In this case one must first specify the constraint that affects
the whole range and then the constraint for the single element. This way
the constraint of the target value overrides the previous constraint on
the upper limit for the selected element. 

For example, the following
constraint statements limit the maximum horizontal beta function between
'marker1' and 'marker2' to 200 meter and require a horizontal beta
function of 100 meter at element 'name1':  
\madxmp{
CONSTRAINT, SEQUENCE=seqname, RANGE='marker1'/'marker2', BETX<200.0;\\
CONSTRAINT, SEQUENCE=seqname, RANGE='name1'/'marker2', BETX=100.0;
}

When the two constraint statements are interchanged, and supposing that
name1 is an element in the range marker1/marker2, the horizontal beta 
function at element 'name1' will only be limited to less than 200 meter
and NOT constrained to 100 meter! 

The CONSTRAINTS can either be specified with explicit values for the
constraints of the optic functions or via a pre-calculated
\hyperref[sec:savebeta]{\texttt{SAVEBETA}} module. The first
options has the form: 
\madbox{
CONSTRAINT, \=SEQUENCE=seqname, RANGE=position,\\
            \>BETX=real, ALFX=real, MUX=real, \\
            \>BETY=real, ALFY=real, MUY=real, \\
            \>X=real, PX=real, Y=real, PY=real, \\
            \>DX=real, DY=real, DPX=real, DPY=real;
}

Here all \hyperref[subsec:tables-linear]{linear lattice functions} 
(BETX, BETY, ALFX, ALFY, MUX, MUY, DX, DY, DPX, DPY)
or \hyperref[subsec:tables-chrom]{chromatic lattice functions}
(WX, XY, PHIX, PHIY, DMUX, DUMY, DDX, DDY, DDPX, DDPY)
are constrained at the selected range to the corresponding values.

The second form of the CONSTRAINT command has the form
\madbox{
CONSTRAINT, \=SEQUENCE=seqname, RANGE=position, \\
            \>BETA0=beta0-name, MUX=real, MUY=real;
}

Here all of (\texttt{BETX, BETY, ALFX, ALFY, MUX, MUY, DX, DY, DPX, DPY})
are constrained in the selected points to the corresponding values
of a pre-calculated \hyperref[sec:savebeta]{\texttt{SAVEBETA}} module.
In the above example the phases (\texttt{MUX, MUY}) are overridden by the 
numerical values specified via \texttt{MUX=real} and \texttt{MUY=real}.
Normally \hyperref[sec:range]{\texttt{RANGE}} refers to a single position.

\section{User Defined Matching Constraints}
\label{sec:userconstraint}

In addition to the nominal \hyperref[chap:twiss]{\texttt{TWISS}} variables, 
the user can define a limited set of 'user-defined' variables in the
constraint statement.
This allows, for example, the matching of the normalized dispersion or the
mechanical aperture. The \texttt{MATCH} module allows four user defined
variables called: \texttt{MVAR1, MVAR2, MVAR3} and \texttt{MVAR4}. 
The variables can be defined according to the general variable
declaration rules of \hyperref[sec:defer]{deferred expressions}.
For example, in order to match the normalized dispersion at a certain
location in the sequence one would first define a variable:

\madxmp{
MVAR1 := table(twiss,dx)/sqrt(table(twiss,betx));
}

After that the user has to select the variable for output in the TWISS
statement (see details in the \hyperref[chap:twiss]{TWISS module} and 
\hyperref[sec:select]{SELECT} statement):
\madxmp{
SELECT, FLAG=twiss, CLEAR; \\
SELECT, FLAG=twiss, COLUMN=keyword, name, s, betx, dx, mvar1; \\
TWISS, SEQUENCE=seqname, FILE=twissfile;
}

The variable can now be referenced like any other \texttt{TWISS} variable
in the constraint command:
\madxmp{
CONSTRAINT, SEQUENCE=seqname, RANGE=range, MVAR1=targetvalue;
}

\section{GLOBAL}
\label{sec:global}

In addition to conventional matching constraints that specify the optics 
functions at a certain position in the sequence the user can also constrain 
global optics parameters such as, for example, the overall machine tune
and the machine chromaticity. Such global optics parameters can be
constraint via the \texttt{GLOBAL} command, having the following syntax:

\madbox{
GLOBAL, \=SEQUENCE=seqname, \\
        \>Q1=real, DQ1=real, \\ %, DDQ1=real, \\
        \>Q2=real, DQ2=real;    %, DDQ2=real, \\
%       \>DQ1DE1=real, DQ2DE2=real, DQ1DE2=real, \\
%       \>GAMMATR=real, ALFA=real;
}


All attributes are optional and have the following meaning:
\begin{madlist}
  \ttitem{SEQUENCE} the name of the sequence on which to operate the matching.
  \ttitem{Q1, Q2} enable a correction of tunes in presence of
  magnetic imperfections or misalignments
  \ttitem{DQ1, DQ2}
  enable a correction of chromaticities in presence of
  magnetic imperfections or misalignments
%  \ttitem {DDQ1, DDQ2} enable a correction of nonlinear chromaticities
%  \ttitem {DQ1DE1, DQ2DE2, DQ1DE2} enable a correction of anharmonicities,   
% \textsl{ie} the detuning with amplitudes
%  \ttitem {GAMMATR} enable a correction of the transition energy of the lattice
%  \ttitem {ALFA} enable a correction of the momentum compaction factor of the  
%lattice.
\end{madlist}


\section{WEIGHT, GWEIGHT}
\label{sec:weight}\label{sec:gweight}

The matching procedures try to fulfil the constraints
in a least square sense.
A penalty function is constructed which is the sum of the
squares of all errors, each multiplied by the specified weight.
The larger the weight, the more important a constraint becomes.
The weights are taken from a table of current values.
These are initially set to default values listed in the table below,
and may be reset at any time to different values.
Values set in this way remain valid until changed again.

The \texttt{WEIGHT} command changes the weights for subsequent
constraints: 
\madbox{
WEIGHT, \=BETX=real, ALFX=real, MUX=real, \\
        \>BETY=real, ALFY=real, MUY=real, \\
        \>X=real, PX=real, Y=real, PY=real, \\
        \>DX=real, DPX=real, DY=real, DPY=real;
}

The weights are entered with the same name as the
\hyperref[subsec:tables-linear]{linear lattice functions} and
\hyperref[subsec:tables-canon]{orbit coordinate} 
to which the weight applies.
Frequently the matching weights serve to select a restricted
set of functions to be matched.

Matching weights for global matching constraints can be set by the 
\texttt{GWEIGHT} command, having attributes identical to those of
\texttt{GLOBAL}.

\madbox{
GWEIGHT, \=Q1=real, DQ1=real, \\ %DDQ1=real, \\
         \>Q2=real, DQ2=real;  % DDQ2=real, \\
%        \>DQ1DE1=real, DQ2DE2=real, DQ1DE2=real, \\
%        \>GAMMATR=real, ALFA=real;
}


Default values for matching weights are given in the table below.

\begin{table}[ht]
  \centering
  \caption{Default Matching Weights}
  \vspace{1ex}
  \begin{tabular}{|lr|lr|lr|}
    \hline
    NAME   & WEIGHT & NAME   & WEIGHT & NAME   & WEIGHT \\
    \hline
    BETX   &   1.0  & ALFX   &  10.0  & MUX    &  10.0  \\
    BETY   &   1.0  & ALFY   &  10.0  & MUY    &  10.0  \\ 
    X      &  10.0  & PX     & 100.0  &  & \\
    Y      &  10.0  & PY     & 100.0  &  & \\
    T      &  10.0  & PT     & 100.0  &  & \\ 
    DX     &  10.0  & DPX    & 100.0  &  & \\
    DY     &  10.0  & DPY    & 100.0  &  & \\ 
    WX     &   1.0  & PHIX   &   1.0  & DMUX   &   1.0  \\
    WY     &   1.0  & PHIY   &   1.0  & DMUY   &   1.0  \\ 
    DDX    &   1.0  & DDPX   &   1.0  &  & \\
    DDY    &   1.0  & DDPY   &   1.0  &  & \\ 
    MVAR\textit{i}  &  10.0  & & & & \\
    Q1     &  10.0  & DQ1    &   1.0  &  & \\ %& DDQ1    &  0.1  \\
    Q2     &  10.0  & DQ2    &   1.0  &  & \\ %& DDQ2    &  0.1  \\
    %      DQ1DE1 &    ??  & DQ2DE2 &    ??  & DQ1DE2  &   ??  \\
    %      GAMMATR&    ??  & ALFA   &    ??  &  & \\
    \hline
    %% name   & weight & name   & weight & name   & weight & name   & weight & name   & weight & name   & weight \\ 
    %% BETX   &   1.0  & ALFX   &  10.0  & MUX    &  10.0  & BETY   &   1.0  & ALFY   &  10.0  & MUY    &  10.0  \\ 
    %% X      &  10.0  & PX     & 100.0  & Y      &  10.0  & PY     & 100.0  & T      &   0.0  & PT     &   0.0  \\ 
    %% DX     &  10.0  & DPX    & 100.0  & DY     &  10.0  & DPY    & 100.0  & - \\ 
    %% WX     &  10.0  & PHIX   &  10.0  & DMUX   & 100.0  & WY     &  10.0  & PHIY   &  10.0  & DMUY   & 100.0  \\ 
    %% DDX    &  10.0  & DDPX   & 100.0  & DDY    &  10.0  & DDPY   & 100.0  & - \\ 
    %% MVAR1    & 10.0  & MVAR2    & 10.0  & MVAR3    & 10.0  & MVAR4    & 10.0  & -
  \end{tabular}
\end{table}



%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002


%\input{match/match_xeq}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 

\section{Matching Methods}
\label{sec:match-methods}

\madx currently supports four different matching algorithms each associated to 
a command with its own attributes. 

\subsection{LMDIF: Fast Gradient Minimisation}
\label{subsec:match-lmdif}
The \texttt{LMDIF} command minimises the sum of squares of the constraint
functions using their numerical derivatives. It is the fastest
minimisation method available in \madx.
 
\madbox{
LMDIF, CALLS=integer, TOLERANCE=real;
}

The command has two attributes:  
\begin{madlist}
   \ttitem{CALLS} The maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} The desired tolerance for the minimum. 
   (Default:~1E-6)  
\end{madlist}

\subsection{MIGRAD: Gradient Minimisation}
\label{subsec:match-migrad}
The \texttt{MIGRAD} command minimises the penalty function using the numerical
derivatives of the sum of squares.  

\madbox{
MIGRAD, CALLS=integer, TOLERANCE=real, STRATEGY=integer;
}

The command has three attributes: 
\begin{madlist}
   \ttitem{CALLS} the maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} the desired tolerance for the minimum. 
   (Default:~1E-6)
   \ttitem{STRATEGY} specifies the strategy to be used. (Default:~1) \\
   Details are given in \cite{MINUIT}.  
\end{madlist} 

\subsection{SIMPLEX: Simplex Minimisation}
\label{subsec:match-simplex}
The \texttt{SIMPLEX} command minimises the penalty function by the simplex
method. Details are given in \cite{MINUIT}.

\madbox{
SIMPLEX, CALLS=integer, TOLERANCE=real;
}

The command has two attributes: 
\begin{madlist}
   \ttitem{CALLS} The maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} The desired tolerance for the minimum. 
   (Default:~1E-6)
\end{madlist} 

\subsection{JACOBIAN: Newton Minimisation}
\label{subsec:match-jacobian}
The \texttt{JACOBIAN} command minimises the penalty function calculating the
Jacobian and solving the linear problem. A $Q R$ or $L Q$  decomposition is
performed when the system is over or under-determined. Before starting
the matching routine two optional transformations (\texttt{COOL} and
\texttt{RANDOM}) are performed. 

\madbox{
JACOBIAN, \=CALLS=integer, TOLERANCE=real, REPEAT=integer, \\
          \>STRATEGY=integer, COOL=real, BALANCE=real, \\
          \>RANDOM=real;
}

The command has the folowing attributes: 
\begin{madlist}
  \ttitem{CALLS} The maximum number of calls to the penalty
  function. (Default:~30)  

  \ttitem{TOLERANCE} The desired tolerance for the minimum. 
  (Default:~1E-6)

  \ttitem{REPEAT} The number of calls of the JACOBIAN routine. 
  (Default:~1) 

  \ttitem{BISEC} Selects the maximum number of iteration used to
  determine the step length which reduces the penalty function during
  the main iteration. A large number (i.e. 6) reduces the probability
  to diverge from the solution, but increases the probability of being
  trapped in a local minimum.  

  \ttitem{STRATEGY} A code for the strategy to be used. (Default:~3)\\
  If STRATEGY=1 the routine resets the values of the variables which
  exceeds the limits. If STRATEGY=2 the routine print the Jacobian
  and exit without matching. If STRATEGY=3 the routine  disables the
  variables which exceeds the limits keeping however the number of
  variables greater or equal to the number of the constraints.  

  \ttitem{COOL, BALANCE} The factors which specify the following
  transformation: 
  \madxmp{
   xxx\=xxxxxxxxx\= \kill
   if \>"balance" >=0 \\
      \>newval = (1-cool)*oldval +  \& \\
      \>       \>cool*((1-balance)*maxval + balance*minval)\\
   else \\
      \>newval = (1-cool)*oldval + cool*optval
  }
  where \texttt{newval} is the new value after the transformation,
  \texttt{oldval} is the previous value, \texttt{maxval, minval,
    optval} are the maximum, minimum and optimal value of
   the variable specified in the \texttt{VARY} command. 

   \ttitem{RANDOM} The factors which specify the following transformation:
   \madxmp{newval = ( 1 + random * rand() ) * oldval}
   where \texttt{newval} is the new value after the transformation,
   \texttt{oldval} is the previous value, \texttt{rand()} is a stochastic
   variable with a uniform (-0.5,0.5) distribution.   
\end{madlist} 


%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002. 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006. 


%\input{match/match_um}
%%\title{Expression Matching with USE\_MACRO}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 

\section{USE\_MACRO}
\label{sec:use-macro}
 
It is possible to match user defined expressions with the
\texttt{USE\_MACRO} keyword. 

The general input structure for a \texttt{MATCH} command is the following:

\begin{verbatim}
MATCH,USE_MACRO;
... VARY statements ...
USE_MACRO, NAME=macro1;
     or
macro1: MACRO={ ... madx statements};
CONSTRAINT, expr=  "lhs1 < | = | > rhs1";
CONSTRAINT, weight=2, expr=  "lhs2 < | = | > rhs2";
...  CONSTRAINT statements ...
MACRO 2 definition
... CONSTRAINT statements ...
MACRO n definition
... CONSTRAINT statements ...
... METHODS statements ...
ENDMATCH;
\end{verbatim}
 
The algorithm for evaluating the penalty function is the following:
 
\begin{itemize}
   \item  execute the first macro,
   \item  evaluate and compute the difference between the left hand side
     (lhs) and the right hand side (rhs) of the first set of expressions, 
   \item in case of other macros, evaluates in order the macro and the
     expressions 
   \item  the set of differences are  minimized by the selected method
     using the variables defined in the VARY statements. 
\end{itemize}

\subsection{Initiating the Matching Module with \texttt{USE\_MACRO}}
 
With:
\madxmp{
MATCH, USE\_MACRO;
}
the \texttt{MATCH} command can be used for matching any expression which
can be defined through expression. It requires a slightly different syntax.

\subsection{VARY statements}
In the \texttt{USE\_MACRO} mode the \texttt{VARY} statement follows the
same rules of the other modes explained in the section
\hyperref[sec:vary]{Define Variable Parameter} 

\subsection{Macro definitions}
The macro to be used in the matching routine can be defined in two ways:
 
\begin{itemize}
\item using USE\_MACRO statement:
  \madxmp{USE\_MACRO, NAME=macro1;}
  defining a new macro on the fly using the usual syntax for
  \hyperref[sec:macro]{\texttt{MACRO}}s.  
\end{itemize}
 
After a macro definition a set of constraints should be defined, with
the following syntax for the \texttt{CONSTRAINT} command:
 
\madxmp{
CONSTRAINT, expr = "lhs = rhs"; \\
CONSTRAINT, expr = "lhs < rhs"; \\
CONSTRAINT, expr = "lhs > rhs"; 
}

where "lhs" and "rhs" are well defined \madx expressions. 
Another set of macro and constraints can be defined afterwards. 

\subsection{Examples}
The \texttt{USE\_MACRO} mode can emulate a matching script
which uses the normal syntax. 

Normal syntax:
\begin{verbatim}
MATCH,SEQUENCE=LHCB1,LHCB2;
    VARY, NAME=KSF.B1, STEP=0.00001;
    VARY, NAME=KSD.B1, STEP=0.00001;
    VARY, NAME=KSF.B2, STEP=0.00001;
    VARY, NAME=KSD.B2, STEP=0.00001;
    GLOBAL,SEQUENCE=LHCB1,DQ1=QPRIME;
    GLOBAL,SEQUENCE=LHCB1,DQ2=QPRIME;
    GLOBAL,SEQUENCE=LHCB2,DQ1=QPRIME;
    GLOBAL,SEQUENCE=LHCB2,DQ2=QPRIME;
    LMDIF, CALLS=10, TOLERANCE=1.0E-21;
ENDMATCH;
\end{verbatim}

USE\_MACRO syntax:

\begin{verbatim}
MATCH,USE_MACRO;
    VARY, NAME=KSF.B1, STEP=0.00001;
    VARY, NAME=KSD.B1, STEP=0.00001;
    VARY, NAME=KSF.B2, STEP=0.00001;
    VARY, NAME=KSD.B2, STEP=0.00001;
    M1: MACRO={ TWISS,SEQUENCE=LHCB1; };
    CONSTRAINT, EXPR= TABLE(SUMM,DQ1)=QPRIME;
    CONSTRAINT, EXPR= TABLE(SUMM,DQ2)=QPRIME;
    M2: MACRO={ TWISS,SEQUENCE=LHCB2; };
    CONSTRAINT, EXPR= TABLE(SUMM,DQ1)=QPRIME;
    CONSTRAINT, EXPR= TABLE(SUMM,DQ2)=QPRIME;
    LMDIF, CALLS=10, TOLERANCE=1.0E-21;
ENDMATCH;
\end{verbatim}

%\line(1,0){300}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning},October, 2003;
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006.


%\input{match/match_xmpl}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 16-Sep-1997 
%  Changed by: Oliver Bruning, 27-Jun-2002 
%  Changed by: Hans Grote, 25-Sep-2002 
 
\section{Matching Examples}
\label{sec:match-examples}

All matching examples and the related files for executing the \madx
sample jobs can be found in the examples directory
(\href{https://github.com/MethodicalAcceleratorDesign/madx-examples}{https://github.com/MethodicalAcceleratorDesign/madx-examples})
of the \madx website.

\begin{itemize}
   \item Simple Periodic Cell\\
     Match a simple cell to given phase advances:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/5cell/job.5cell.madx}
     {FIVE-CELL}
     
   \item Simple Periodic Cell\\
     Match the matrix elements of the linear transfer matrix at the
     end of a sequence 5 periodic cells:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/r-matrix/job.r-matrix.madx}
     {RMATRIX}
     
   \item Transfer line with initial conditions\\
     Match a sequence of 5 periodic cells with initial conditions  to
     given beta-functions at the end of the sequence:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/line/job.line.madx}
     {Transfer line}
     
   \item Global tune matching in a sequence of 5 periodic cells \\
     Match the global tune of a sequence of 5 periodic cells:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/global-tune/job.global-tune.madx}
     {Global tune}
     
   \item Global tune matching for the LHC\\
     Match the global tune for beam1 of the LHC:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.tune/job.lhc.tune.madx}
     {Global tune for the LHC}
     
   \item Global chromaticity matching for the LHC\\
     Match the global chromaticity for beam1 of the LHC:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.chromaticity/job.lhc.chromaticity.madx}
     {Global chromaticity for the LHC}
     
   \item Global chromaticity matching for both beams of the LHC\\
     Match the global chromaticity for beam1 and beam2 of the LHC:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.2chromaticity/job.lhc.2chromaticity.madx}
     {Global chromaticity for both beams of the LHC}
     
   \item IR8 insertion matching for beam1 of the LHC\\
     Match the insertion IR8 with initial conditions to given values
     of the optics  functions at the IP and the end of the insertion:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.insertion/job.lhc.insertion.madx}
     {IR8 insertion matching} for beam1 of the LHC
     
   \item IR8 insertion matching for beam1 of the LHC with upper
     limits on the optics functions\\
     Match the insertion IR8 with initial conditions to given values
     of the optics  functions at the IP and the end of the insertion
     while limiting the maximum acceptable beta functions over the
     whole insertion:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.insertion-upper/job.lhc.insertion-upper.madx}
     {IR8 insertion matching}
     for beam1 of the LHC with upper limits for
     all beta functions inside the insertion
     
   \item Simultaneous orbit matching at IP8 for beam1 and beam2 of
     the LHC\\
     Match simultaneously the orbit of beam1 and beam of the LHC at
     IP8  with initial conditions to the same given values at the IP:
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.iporbit/job.lhc.iporbit.madx}
     {Orbit matching at IP8} for beam1 and beam2 of the LHC
     
   \item IR8 beta squeeze for beam1 using JACOBIAN matching routine\\
     Try to find a beta squeeze for IR8 starting from 10 meters.
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhcV65.ir8squeeze/job.lhcV65.ir8squeeze.madx}
     {Beta squeeze for IR8}
     
   \item Matching first and second order chromaticity of the LHC
     using USE\_MACRO option. \\
     Match simultaneously the first and second order chromaticity by
     defining macros which compute them using the TWISS command or
     PTC.
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/lhc.qpp/job.lhc.qpp.madx}
     {Second order chromaticity}
     
   \item Matching s position using VLENGTH flag.\\
     match the positions of elements and the total sequence length
     for a simple sample sequence.
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/s-match/job.s-match.madx}
     {s position matching}
     
   \item Matching s position using USE\_MACRO.\\
     match the positions of elements and the total sequence length for a simple sample sequence using USE\_MACRO.
     \href{https://github.com/MethodicalAcceleratorDesign/madx-examples/blob/master/match/s-match-usemacro/job.s-match-usemacro.madx}
     {s position matching}
     
\end{itemize}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002; 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, August, 2007. 

%% EOF
