%%\title{PTC Set-up Parameters}
%  Created by: Valery KAPIN, 21-Mar-2006 
%  Changed by: ____________, ___________ 

\chapter{\ptc Set-up Parameters}
\label{chap:ptc-setup}

The Polymorphic Tracking Code \cite{forest2002} of Etienne
Forest is a kick code, allowing a symplectic integration through all
accelerator elements giving the user full control over the precision
(number of   steps and integration type) and exactness (full or extended
Hamiltonian) of the   results. 
The degree of exactness is determined by the user and the speed of his
computer.  
The main advantage is that the code is inherently based on the map
formalism and provides users with all associated tools. 

The \ptc code is actually a library that can be used in many different
ways to create an actual module that calculates some property of
interest. 

{\bf Attention:}
\ptc exists inside of \madx as a library. \madx offers the interface to
\ptc, i.e. the \madx input file is used as input for \ptc. Internally,
both \ptc and \madx have their own independent databases which are
linked via the interface. With the \hyperref[sec:ptc-create-layout]{\tt
  PTC\_CREATE\_LAYOUT} command, only numerical values are transferred
from the \madx data structures to the \ptc data structures. Any
modification to the \madx data structure is unknown to \ptc until the
next call to \hyperref[sec:ptc-create-layout]{\tt PTC\_CREATE\_LAYOUT}. 
For example, a \hyperref[sec:defer]{deferred expression} of \madx is
only evaluated at the time of the \hyperref[sec:ptc-create-layout]{\tt
  PTC\_CREATE\_LAYOUT} command and is ignored within \ptc afterwards.


Several modules using the PTC code have been presently implemented in
\madx. These {\madx}-{\ptc} modules\cite{schmidt2005} are executed by
the following commands: 
\hyperref[chap:ptc-twiss]{\tt PTC\_TWISS},
\hyperref[chap:ptc-normal]{\tt PTC\_NORMAL},
\hyperref[chap:ptc-track]{\tt PTC\_TRACK},
\hyperref[sec:ptc-trackline]{\tt PTC\_TRACK\_LINE}. 

To perform calculations with these {\madx}-{\ptc} commands, the \ptc
environment must be initialized, handled and turned off by special
commands within the \madx input script. 

\section{Command Synopsis}
\label{sec:ptc-synopsis}

A typical set of commands to invoke \ptc is given below: 

\madxmp{
PTC\_CREATE\_UNIVERSE, \=SECTOR\_NMUL\_MAX= integer, SECTOR\_NMUL= integer,  \\
                       \>NTPSA= logical, SYMPRINT= logical; \\
\\
PTC\_CREATE\_LAYOUT, \=TIME= logical, MODEL= integer, \\
                     \>METHOD= integer, NST= integer, EXACT= logical, \\
                     \>OFFSET\_DELTAP= double, ERRORS\_OUT= logical, \\
                     \>MAGNET\_NAME= string, RESPLIT= logical,  \\
                     \>THIN= double, XBEND= double,  \\
                     \>EVEN = logical; \\
... \\
PTC\_MOVE\_TO\_LAYOUT, INDEX= integer; \\
...\\
PTC\_READ\_ERRORS, OVERWRITE= logical; \\
... \\
PTC\_ALIGN; \\
...\\
PTC\_END;
}

\section{PTC\_CREATE\_UNIVERSE}
\label{sec:ptc-create-universe}

The {\tt PTC\_CREATE\_UNIVERSE} command is required to set-up the
\ptc environment.  

\madbox{
PTC\_CREATE\_UNIVERSE, \=SECTOR\_NMUL\_MAX=integer, SECTOR\_NMUL=integer, \\
                       \>NTPSA=logical, SYMPRINT=logical;
}

The attributes are:
\begin{madlist}

   \ttitem{SECTOR\_NMUL\_MAX} a global variable in \ptc needed for exact
   sector bends defining up to which order Maxwell's equation are solved
   (see \cite{forest2002}~page~76-77). 
   The value of {\tt SECTOR\_NMUL\_MAX} must not be smaller than
   {\tt SECTOR\_NMUL} otherwise \madx stops with an error. \\ (Default:~10)

   \ttitem{SECTOR\_NMUL} a global variable in PTC needed for exact
   sector bends defining up to which order the multipole are included in
   solving Maxwell's equation up to order {\tt SECTOR\_NMUL\_MAX}. 
   Multipoles of order N with N $>$ {\tt SECTOR\_NMUL} and N $\leq$ 
   {\tt SECTOR\_NMUL\_MAX} are treated similar to {\it SixTrack}. \\ 
   (Default:~10)  

   \ttitem{NTPSA} invokes the Differential Algebra (DA) package
   written in C++ and kindly provided by Lingyun Yang (lyyang@lbl.gov). \\ 
   Etienne Forest has written the wrapper to allow the use of both
   the legendary DA package written in Fortran by Martin Berz
   (default) and this new DA package of Lingyun~Yang. 
   It is expected that this DA package will allow for the efficient
   calculation of a large number of DA parameters. \\ (Default:~false)

   \ttitem{SYMPRINT} a flag to enable the printing of the check of
   symplecticity. It is recommended to leave this flag set to TRUE. \\
   (Default: true)
\end{madlist}


\section{PTC\_CREATE\_LAYOUT}
\label{sec:ptc-create-layout}

The {\tt PTC\_CREATE\_LAYOUT} command creates the \ptc-layout
according to  the specified integration method and fills it with the
current \madx sequence defined in the latest
\hyperref[sec:use]{\tt USE} command.

\madbox{
PTC\_CREATE\_LAYOUT, \=TIME=logical, MODEL=integer, METHOD=integer,  \\
                     \>NST=integer, EXACT=logical, OFFSET\_DELTAP=double, \\
                     \>ERRORS\_OUT=logical, MAGNET\_NAME=string, \\
                     \>RESPLIT=logical, THIN=double, XBEND=double, \\ 
                     \>EVEN=logical; 	
}

The attributes are:
\begin{madlist}

  \ttitem{TIME} a logical flag to control which coordinate system 
  is being used. \\ (Default=~true) \\ \\
  This option changes the canonical coordinate system depending
  whether the calculation is done in 5D or 6D: 
  \begin{madlist}
    \ttitem{5D} if {\tt TIME} is true, the fifth coordinate is
    \hyperref[subsec:tables-canon]{\tt PT}, $p_t = \Delta E / p_0 c$ \\ 
    if {\tt TIME} is false, the fifth coordinate is
    \hyperref[subsec:tables-canon]{\tt DELTAP}, $\delta_p = \Delta p / p_0$
    
    \ttitem{6D} if {\tt TIME} is true, the \hyperref[subsec:tables-canon]{\madx
      coordinate system} \{$-ct$, $p_t$\} is used. \\   
    if {\tt TIME} is false, the second \ptc coordinate system
    \{-pathlength, $\delta_p$\} is used. 
  \end{madlist}
  
 {\bf Note:} at small energy ($\beta_0 << 1$),
 momentum-dependent variables like dispersion will depend  strongly on
 the choice of  the logical input variable "time".  In fact, the
 derivative ($\frac{\partial}{\partial \delta_p}$)  and
 ($\frac{\partial}{\partial p_t}$)  are different by the
 factor $\beta_0$. One would  therefore typically  choose
 the option "time=false",  which sets the fifth variable to
 the relative momentum deviation $\delta_p$.

   
  \ttitem{MODEL} an integer to switch between models:\\
  1 for "Drift-Kick-Drift";  (Default value)\\ 
  2 for "Matrix-Kick-Matrix" and \\ 
  3 for "Delta-Matrix-Kick-Matrix" (SixTrack-code model).

  \ttitem{METHOD} the integration order: 2, 4, or 6 (See
  \cite{forest2002}~Chapter~K) \\ (Default: 2)

  \ttitem{NST} the number of integration steps.  (Default:~1)\\
  The body of each element is divided into {\tt NST} equal slices and
  Forest-Yoshida integration is carried out on each slice.
  For best results {\tt NST} should increase with strength
  and length of elements. The optimum {\tt NST} value corresponds to 
  the value beyond which the studied properties no longer change. 
  However, for time consuming calculations the user may
  reduce {\tt NST}. (See below the {\tt RESPLIT} option for automatic
  adjustment.)\\
  This attribute sets the same {\tt NST} value  for all "thick" elements
  ($l > 0$) of a beam-line; however each individual element may also have
  its own {\tt NST} value defined independently 
  (\hyperref[sec:add-option-PTC]{see below}).

  \ttitem{EXACT} a logical flag to turn on calculations with an exact
  Hamiltonian, otherwise the expanded Hamiltonian is used. \\
  (Default:~false)

  \ttitem{OFFSET\_DELTAP} {\bf [ Beware: Expert attribute! ]}\\
  provides relative momentum deviation of the reference particle (6D case
  ONLY). This option implies {\tt TOTALPATH=true}. \\
  (Default: 0.0)

  \ttitem{ERRORS\_OUT} a logical flag to write-out multipolar errors
  in \hyperref[sec:efcomp]{\tt EFCOMP} table format. 
  \\ (Default: false) \\
  Two tables are created and filled: "errors\_field" contains only
  field errors, "errors\_total" contains also desired field
  components, which can include the strength of correctors.  
  The choice of magnets is defined by the {\tt MAGNET\_NAME}
  attribute (see below). 
  The tables can be \hyperref[sec:write]{written} to file, and can be
  read back via the {\tt ERRORS\_IN} flag.\\ 
  The {\tt ERRORS\_IN} flag has precedence over this {\tt ERRORS\_OUT} flag.

  \ttitem{MAGNET\_NAME} a string giving a simple selection for the
  names of magnet to be used for an error write-out using the {\tt
  ERRORS\_OUT} flag (see above). The errors are recorded for all
  magnets with names starting with the exact string given here, which
  would be equivalent to the ??? regular expression.\\
  (Default:~nil)

  \ttitem{RESPLIT} a logical flag to apply the \ptc resplit
  procedure. This is meant to create an "adaptive" setting of the {\tt
    METHOD} and {\tt NST} attributes according to the strengths of
  quadrupoles (using the {\tt THIN}  attribute) and dipoles (using the
  {\tt XBEND} attribute). The {\tt EVEN} attribute further controls the
  number of splits.  \\
  (Default:~false)

  \ttitem{THIN} is the main {\tt RESPLIT} attribute and is meant for
  splitting quadrupoles according to their strength. The default value
  {\tt THIN=0.001} has shown in practice to work well without costing
  too much with respect of performance.
  
  \ttitem{XBEND} is an optional {\tt RESPLIT} attribute and is meant for
  splitting dipoles. A value {\tt XBEND=0.001} is also advisable for
  dipoles. \\
  (Default: -1.0 for no splitting)

  \ttitem{EVEN} a logical switch to ensure even number of splits when
  using the {\tt RESPLIT}  procedure of \ptc, which is particularly
  useful when one attempts to calculate {\tt PTC\_TWISS} with the {\tt
    CENTER\_MAGNETS} option, i.e. to calculate the {\tt TWISS}
  parameters in the center of the element. 
  Uneven number of splits is ensured with {\tt EVEN=false}.  
  (Default:~true)
\end{madlist}

\section{PTC\_MOVE\_TO\_LAYOUT}
\label{sec:ptc-move-to-layout}

Several \ptc layouts can be created within a single \ptc-"universe". 
The layouts are automatically numbered with sequential integers by the 
\madx code. The {\tt PTC\_MOVE\_TO\_LAYOUT} command is used to
activate a specific layout, and the next \ptc commands will be
applied to this active \ptc layout until a new \ptc layout is created
or activated. 

\madbox{
PTC\_MOVE\_TO\_LAYOUT, INDEX=integer;
}

The only attribute is:
\begin{madlist}
	\ttitem{INDEX} is the numeric index of the \ptc layout to be
	activated.\\ (Default:~1)
\end{madlist}

\section{PTC\_READ\_ERRORS}
\label{sec:ptc-read-errors}

The {\tt PTC\_READ\_ERRORS} command reads any number of
\textbf{"errors\_read"} table through the 
\hyperref[sec:readmytable]{\tt READMYTABLE} mechanism.

\madbox{
PTC\_READ\_ERRORS, OVERWRITE=logical; 
}

The only attribute is 
\begin{madlist}
   \ttitem{OVERWRITE} a flag to specify that the read-in errors
   overwrite previous errors instead of adding the read-in errors to
   existing errors, ie multipole components already present.\\
   (Default:~false)  
\end{madlist}

{\bf Note:}\\
Because of the way the table is read in memory, a warning will always be
issued by default in the form:
\begin{verbatim}
warning: string_from_table_row: row out of range: errors_read->name[1>=n+1<=n]
\end{verbatim}
where {\tt n} is  the number of records read from the table. 
This warning has no consequence on the errors read and the following
calculation. \\
The warning is purely the result of the way that the reading loop is
programmed with a break based on the return value of the routine
{\sl string\_from\_table\_row}.  
But if {\sl string\_from\_table\_row} tries to read in a row ({\tt n+1})
past the last row ({\tt n}) of the table, it prints a warning before
returning a value that will effectively break the loop. Of course this
will only happen if the {\tt WARN} option is true and this can be turned
off with \madxmp{OPTION, -WARN;}

       

\section{PTC\_ALIGN}
\label{sec:ptc-align}

The {\tt PTC\_ALIGN} command is used to apply the \madx alignment
errors to the current PTC layout, and takes no attributes.

\madbox{
PTC\_ALIGN;
}


\section{PTC\_END}
\label{sec:ptc-end}

The {\tt PTC\_END} command turns off the \ptc environment,
which releases all memory and returns control to the \madx world proper.

\madbox{
PTC\_END;
}


\section{Additional Options for Physical Elements}
\label{sec:add-option-PTC}

For some of the \madx elements, additional attributes can be defined
that are available to \ptc only. \ptc also uses standard \madx
attributes in a slightly different way.

\madbox{
SBEND | \=RBEND | QUADRUPOLE | SEXTUPOLE | OCTUPOLE | SOLENOID ,\\
        \>L=real, ... , TILT=real, ... , NST=integer, ... ,\\
        \>KNL=\{0, real, real,...\}, KSL=\{0, real, real,...\}; 
}

These attributes are:
\begin{madlist}
  \ttitem{L} the length of the element. \\ \ptc treats bending magnets
  (\hyperref[sec:bend]{\tt SBEND} or \hyperref[sec:bend]{\tt RBEND})
  as \hyperref[sec:marker]{\tt MARKER} if their length is equal to zero. 
     
  \ttitem{NST} gives a specific \hyperref[sec:ptc-twiss]{\tt NST}
  values for a particular "thick" element ($L > 0$). \\
  \\
  For example RF cavities are represented in \madx 
  by a single kick, while \ptc splits the RF cavity into
  {\tt NST} segments thereby taking into account properly the
  transit-time effects of the cavity. Specifying explicitly {\tt NST=1}
  for RF cavity reproduces in \ptc the approximate results of \madx,
  ignoring transit time effects.
  \\

  \ttitem{KNL, KSL} The full range of  normal and skew multipole
  components on the bench can be  specified for the following physical
  elements: 
  \hyperref[sec:bend]{\tt sbend}, 
  \hyperref[sec:bend]{\tt rbend},
  \hyperref[sec:quadrupole]{\tt quadrupole},
  \hyperref[sec:sextupole]{\tt sextupole},  
  \hyperref[sec:octupole]{\tt octupole} and 
  \hyperref[sec:solenoid]{\tt solenoid}. 
  {\tt KNL} and {\tt KSL} multipole coefficients are specified as the
  integrated value ($\int K ds$) of the field components along the
  magnet axis. The multipole components in \ptc are spread over
  the length of thick elements. This is a considerable
  advantage of \ptc input compared  to \madx  which allows only
  \hyperref[sec:multipole]{thin multipoles}. 
  \begin{madlist}
    \ttitem{KNL} is an array representing the normal multipole
    coefficients. \\ (Default:~0~m$^{-1}$)
    \ttitem{KSL} is an array representing the skew multipole
    coefficients. \\ (Default:~0~m$^{-1}$)
  \end{madlist}
  To preserve the reference orbit of straight elements, dipole
  components for those elements are ignored and must be specified as
  zero: KNL(0)=0, KSL(0)=0. 
   
  A full range of additional multipole \hyperref[sec:efcomp]{field
    errors} can be additionally specified with the
  \hyperref[sec:efcomp]{\tt EFCOMP} command. Errors are added to the
  above multipole fields on the bench. 

\end{madlist}

%\href{mailto:kapin@itep.ru}{  V.Kapin}(ITEP) and 
%\href{mailto:Frank.Schmidt@cern.ch}{  F.Schmidt}, March  2006

