%%\title{RLOT}

\section{RPLOT}
\label{sec:rplot}

\texttt{RPLOT} is a \madx plug-in that privides additional functionality
using \href{http://root.cern.ch}{ROOT}.  It contains several tools.

\begin{description}

\item[\textbf{ RVIEWER }] 
\textit{ plotting tool that handles the results in paramremtric form }

What makes it different from the standard PLOT module of \madx is
that it is also able to deal with the parmateric results. \texttt{RPLOT} 
provides graphical user interface that allows to choose which
functions shall be drawn, set its ranges and adjust all the details
of the plot formatting. Of course, the result is immendiately
visible on the screen, in contrary to the standard plot tool that is
able to work solely in the batch mode. The user can choose several
formats to save his plot, including postscript, gif, pdf, root macro and
many others.       

\texttt{RVIEWER} is able to draw the lattice functions     
\begin{enumerate}
   \item  along the layout 
   \item  at given position in function of one or two knobs  
\end{enumerate}     

It provides a convienient way to set the knob values. As the value is
set, the plotted functions are immediately drawn for the new value.            

In order to run \texttt{RVIEWER} simpy issue "rviewer;" command        

\item[\textbf{RTRACKSTORE}] 
\textit{enables storage of the tracking data in ROOT NTuple/Tree format}

Ntuple and its modern extension called Tree are formats designed
for storing particle tracking data. It is proven to provide the
fastest data writing and reading thanks to column wise I/O
operations. It is commonly used for data storage by HEP
experiments. Additionally, ROOT provides automatical ZIP data
compression that is transparent for the user algorithms.
Morover, ROOT provides wide set of very comfortable tools for
advanced analysis and plotting of the data stored in Trees.    

Addtionally, we plan to extend \texttt{RVIEWER} functionality that would
provide intuitive graphical user interface to most commonly used
features in particle tracking in accelerators. Thanks to that, the user is
not forced to learn how to use the ROOT package.    

Currently the feature is enabled only for tracking using the
\texttt{PTC\_TRACKLINE} command, however, it will be extended to other
tracking modes.           
\end{description}

%\textbf{Download} \\
%The newest version is available \href{download/}{here}\\

\textbf{Installation}\\
Prerequisite: ROOT must be installed before compilation and whenever
the user wants to use the plug-in. See explanations on
\href{http://root.cern.ch}{ROOT webpage}.  \\ 

To install \texttt{RPLOT} 
\begin{enumerate}
   \item  Unpack the archive, it will create directory rplot    
\begin{verbatim}
   tar xvzf rplot-X.XX.tgz
\end{verbatim}

   \item Change to rplot directory    
\begin{verbatim}
   cd rplot
\end{verbatim}

   \item Type     
\begin{verbatim}
   make install
\end{verbatim}
\end{enumerate}

\textbf{Examples}


\textbf{SYNOPSIS}
\begin{verbatim}
RVIEWER; 
\end{verbatim}


\textbf{PROGRAMMERS MANUAL  } \\

 To be continued... 

