%% TRACKING chapter
\chapter{Particle Tracking}
\label{chap:tracking}\label{chap:thintrack}

%%\title{Overview of MAD-X Tracking Modules}

\section{Introduction to \madx Tracking Modules}
\label{sec:trackintro}

A number of particles with given initial conditions can be tracked
through a beam-line or a ring. The particles can be tracked either for a
single passage or for many turns.  


While \madx keeps most of the functionality of \madeight, the
trajectory tracking in \madx is considerably modified compared to
\madeight. 
The reason is that in \madeight the thick lens tracking is inherently not
symplectic, which implies that the phase space volume is not preserved
during the tracking, i.e. contrary to the real particle the tracked
particle amplitude is either growing or decreasing. 


The non-symplectic tracking as in \madeight has been completely excluded
from \madx by taking out the thick lens part from the tracking
modules. Instead two types of tracking modules (both symplectic) are
implemented into \madx. 


The first part of this design decision is the thin-lens tracking module
(\texttt{\href{../thintrack/thintrack.html}{thintrack}})  which tracks
symplecticly through drifts and kicks and by replacing the end effects
by their symplectic part in form of an additional kick on either  end of
the element. This method demands a preliminary conversion of a sequence
with thick elements into one composed of thin elements (see the
\texttt{\href{../makethin/makethin.html}{MAKETHIN}} command). The
details of its usage are given on the page
"\href{../thintrack/thintrack.html}{thintrack}". 


The second part of this design decision is to produce a thick lens
tracking module based on the PTC code of E.~Forest that
allows a symplectic treatment of all accelerator elements giving the
user full control over the precision (number of steps and integration
type) and exactness (full or extended Hamiltonian) of the results. 


The first PTC thick-lens tracking module is named
\href{../ptc_track/ptc_track.html}{ptc\_track}. 
It has the same features as the thin-lens tracking code
(\texttt{\href{../thintrack/thintrack.html}{thintrack}}) except that it
treats thick-lenses in a symplectic manner. 


There is a second PTC tracking module called the line tracking module
(\href{../ptc_track_line/ptc_track_line.html}{ptc\_track\_line}). It is
meant for tracking particles in
\href{http://clic-study.web.cern.ch/CLIC-Study/}{CLIC}, in fact it
treats beam-lines containing traveling-wave cavities and includes a beam
acceleration. 


%See Also \href{../ptc_general/ptc_general.html}{PTC Set-up Parameters}

%\href{mailto:kapin@itep.ru}{ V.Kapin}(ITEP) and 
%\href{mailto:Frank.Schmidt@cern.ch}{ F.Schmidt}, March  2006


%%\title{Thin-Lens Tracking Module (thintrack)}
%  Created by: Andre VERDIER, 21-Jun-2002 
%  Changed by: Andre Verdier, 26-Jun-2002 
%  Changed by: Alexander Koschik, 07-Mar-2006 
%  Changed by: Alexander Koschik, 29-Mar-2006 
%  Changed by: Alexander Koschik, 02-Feb-2007 

\section{Overview of Thin-Lens Tracking} % Module (thintrack)}
\label{sec:trackoverview}

The \textbf{thin-lens tracking module} of \madx performs element per
element tracking of one or several particle trajectories in the last
\href{../control/general.html#use}{\texttt{use}}d sequence.  
 
%  either for single passage (option <var class="option">onepass</var>)
%  or for many turns (default option).  

Only thin elements are allowed (apart from the element \texttt{drift}),
which guarantees the symplecticity of the coordinate transformation. Any
lattice can be converted into a "thin element" lattice by invoking the
\href{../makethin/makethin.html}{\texttt{makethin}} command. 

Several commands are actually required to complete a tracking run:

\madbox{
xxxxxxx\= \kill
TRACK, \>DELTAP=real, ONEPASS=logical, DAMP=logical; \\
       \>QUANTUM=logical, DUMP=logical, APERTURE=logical, \\
       \>ONETABLE=logical, RECLOSS=logical, FILE=filename,\\
       \>UPDATE=logical;\\
xxxx\=xxxxxx\= \kill
  \>\ldots \\
  \>START, X=real, PX=real, Y=real, PY=real, T=real, PT=real;  \\
  \>START, FX=real, PHIX=real, FY=real, PHIY=real, FT=real, PHIT=real;\\  
  \>\ldots \\
  \>OBSERVE, PLACE=string; \\  
  \>\ldots \\
  \>RUN, TURNS=integer, MAXAPER=double\_array, FFILE=integer; \\
  \>\ldots\\
  \>DYNAP, \>TURNS=real, FASTUNE=logical, LYAPUNOV=real,\\
  \>       \>MAXAPER=real\_array, ORBIT=logical;\\
  \>\ldots \\
ENDTRACK;
}


%{\bf Commands:}
%TRACK, DELTAP=double, ONEPASS, DUMP, ONETABLE, FILE= string;  ! (MAD-X version 1)
%TRACK, DELTAP=double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE=string; ! (MAD-X version 2)
%TRACK, DELTAP=double, ONEPASS, DAMP, QUANTUM, DUMP, APERTURE,ONETABLE, RECLOSS, FILE= string; ! (MADX version 3)
%% \begin{verbatim}
%% TRACK, DELTAP=real, ONEPASS, DAMP, QUANTUM, DUMP, APERTURE, ONETABLE, RECLOSS, FILE= string;
%% ...
%% START, X=real, PX=real, Y=real, py=real, t=real, pt=real;
%% START, FX=double, PHIX=double, FY=double, PHIY=double, FT=double, PHIT=double;  
%% ...
%% OBSERVE, PLACE=string;  
%% ...
%% RUN, MAXAPER=double_array, TURNS=integer, FFILE=integer;  
%% ...

%% ENDTRACK;
%% \end{verbatim}


Inside the block \texttt{TRACK}-\texttt{ENDTRACK} a series 
of initial trajectory coordinates can be specified by the \texttt{START} 
command (as many commands as trajectories). This will be usually done in a 
\texttt{while}-loop. \textbf{Note} that the coordinates are either 
\textbf{canonical} coordinates or \textbf{action-angle} variables!

For usual tracking (single/multi-turn), all coordinates are specified
with respect to the actual closed orbit (possibly off-momentum, with
magnet errors) and \textbf{NOT} with respect to the reference orbit. 
            
If the option \texttt{onepass} is used, the coordinates are specified
with respect to the reference orbit. The name "onepass" might be
misleading: Still tracking can be single- or multi-turn!   
            
The tracking is actually started with the \texttt{RUN} command, where
the option \texttt{turns} defines for how many turns the particles will
be tracked in the given sequence. 
          
If the option \texttt{dump} is used, the particle coordinates are
written to files at each turn. The output files are named
automatically. The name given by the user is followed by
.obsnnnn(observation point), followed by .pnnnn(particle number). Hence
filenames look like \texttt{track.obs0001.p0001}.  

Tracking creates a number of internal tables and can create files on disk: 
{\texttt TRACKSUMM, TRACKLOSS}, and \texttt{TRACKONE} or
\texttt{TRACK.OBS\$\$\$\$.P\$\$\$\$} (depending on the attribute
\texttt{ONETABLE} of the \texttt{RUN} command).

These internal tables can be accessed via the
\href{../Introduction/expression.html#table}{\texttt{table}}-access
functions.      

Plotting of particle coordinates or other data in these tables is
possible in \madx. Plotting can also be done with external programs by
using the files created by \texttt{TRACK}.  

{\bf See Also:}\\
\href{../Introduction/aperture.html}{\texttt{APERTURE}}, 
\href{../makethin/makethin.html}{\texttt{MAKETHIN}}



\section{TRACK}
\label{sec:track}

The \texttt{TRACK} command initiates trajectory tracking by entering the 
thin-lens tracking module. 

\madbox{
xxxxxxx\= \kill
TRACK, \>DELTAP=real, ONEPASS=logical, DAMP=logical; \\
       \>QUANTUM=logical, DAMP=logical, APERTURE=logical, \\
       \>ONETABLE=logical, RECLOSS=logical, FILE=filename, \\
       \>UPDATE=logical;
}

The attributes of the TRACK command are:

\begin{madlist}
  \ttitem{DELTAP} relative momentum offset for reference closed orbit (switched
  off for onepass) \\  
  Defining a non-zero \texttt{deltap} results in a change of the beam
  momentum/energy without changing the magnetic properties in the
  sequence, which leads to an off-momentum closed orbit different from
  the on-momentum reference orbit. Particle coordinates are then given
  with respect to this new closed orbit, unless the option
  \texttt{onepass=true} is used! \\  
  (Default:~0.0)

  \ttitem{ONEPASS} flag to ensure that no closed orbit search is done,
  which also means that no stability test is done. This is always the
  case for transfer lines, but this option can also be enabled for
  multi-turn tracking of a circular machine. ONEPASS=true does {\bf NOT}
  restrict tracking to a single turn. \\
  With ONEPASS=true, the particle coordinates are specified with
  respect to the reference orbit. \\  
  With ONEPASS=false, the closed orbit is calculated and the particle
  coordinates are given with respect to the closed orbit coordinates.\\
  The name of this attribute is misleading but was kept for backwards
  compatibility.  \\ 
  (Default:~false)

  \ttitem{DAMP} flag to introduce synchrotron damping (needs RF cavity
  and flag RADIATE in the BEAM command). \\ (Default:~false)

  \ttitem{QUANTUM} flag to introduce quantum excitation via random
  number generator and tables for photon emission. \\ (Default:~false)

  \ttitem{DUMP} flag to write the particle coordinates in files, whose
  names are generated automatically. \\ (Default:~false)

  \ttitem{APERTURE} flag to trigger particle loss if its trajectory is
  outside the aperture of the current element. \\ 
  (Default:~false) \\
  If the \texttt{aperture} option is applied, the \texttt{apertype} 
  and \texttt{aperture} information of each element in the sequence
  is used to check whether the particle is lost or not. For further
  information on the definition of apertures and different aperture
  types, see the documentation of the
  \href{../Introduction/aperture.html}{\texttt{APERTURE}} module. \\  
  Note that if no aperture information was specified for an element, 
  the following procedure still takes place:
  \\
  $\rightarrow$ No aperture definition for element $\rightarrow$ 
  Default apertype/aperture assigned (currently this is   
  \texttt{apertype= circle, aperture = \{0\}}) 
  \\ $\rightarrow$  
  If tracking with \texttt{aperture} is used and an
  element with \texttt{apertype= circle} AND \texttt{aperture= \{0\}}  
  is encountered, then the first value of the \texttt{maxaper} vector
  is assigned as the circle's radius (no permanent assignment!). 
  See option \hyperlink{run}{\texttt{maxaper}} for the default values. 
  \\ $\Rightarrow$
  Hence even if no aperture information is specified by the user for
  certain elements, default values will be used! 


  \ttitem{ONETABLE} flag to write all particle coordinates in a single
  file instead of one file per particle. \\ (Default:~false)

  \ttitem{RECLOSS} flag to create in memory a table named "trackloss"
  containing the coordinates of lost particles.\\
  (Default:~false) \\
  Traditionally, when a particle is lost on the aperture, this information
  is written to stdout. To allow more flexible tracking studies, the
  coordinates of lost particles and additional information can also be
  saved in a table in memory. Usually one would save this table to a
  file using the \texttt{WRITE} command after the tracking run has
  finished. The following information is available in the TFS table
  "trackloss":          
  \begin{itemize}
  \item Particle ID (number)
  \item Turn number
  \item Particle coordinates (x,px,y,py,t,pt)
  \item Longitudinal position in the machine (s)
  \item Beam energy
  \item Element name, where the particle is lost
  \end{itemize}

  \ttitem{FILE} name for the track table. The default name is different
  depending on the value of the ONETABLE attribute. \\ 
  (Default: "track" if ONETABLE=true, "trackone" if ONETABLE=false)

  \ttitem{UPDATE} flag to trigger parameter update per turn. \\  
  (Default:~false) \\
  Specifying UPDATE=true gives access to the following additions:   
  \begin{madlist}
  \ttitem{tr\$turni} this special variable contains the turn number;
  it can be used in expressions like 'KICK := SIN(tr\$turni)' and is
  updated at each turn during tracking.     
  \ttitem{tr\$macro}  this special macro can be
    user-defined and is executed/updated at each turn, during tracking, e.g. \\
    'tr\$macro(turn): macro=\{commands depending on turnnumber;\};' \\ 
    A macro structure is necessary to provide for table access.
  \end{madlist}


\end{madlist}

{\bf Remarks}\\
\emph{IMPORTANT:} If an RF cavity has a non-zero voltage, synchrotron
oscillations are automatically included. If tracking with constant
momentum is desired, then the voltage of the RF cavities has to be set
to zero. If an RF cavity has a no zero voltage and DELTAP is non zero, 
tracking is done with synchrotron oscillations around an off-momentum
closed orbit.


%% \begin{tabular}{c p{5cm} p{3cm} c}
%%   \hline 
%%   \textbf{Option} & \textbf{Meaning} & \textbf{Default Value} &
%%   \textbf{Value Type} \\  
%%   \hline
%%   DELTAP & relative momentum offset for reference closed orbit (switched
%%   off for onepass) &  0.0 & double \\  
%%   \hline
%%   ONEPASS & the sequence is treated as transfer line (no stability test,
%%   ie. no closed-orbit search) & .FALSE.= closed-orbit search & logical
%%   \\  
%%   \hline
%%   DAMP & introduce synchrotron damping (needs RF cavity, RADIATE in
%%   BEAM)  & .FALSE.= no damping & logical \\  
%%   \hline
%%   QUANTUM & introduce quantum excitation via random number generator and
%%   tables for photon emission & .FALSE.= no excitation & logical \\  
%%   \hline
%%   DUMP & write the particle coordinates in files (names generated
%%   automatically)  & .FALSE.= no file generated & logical \\  
%%   \hline
%%   APERTURE & particle is lost if its trajectory is outside the aperture
%%   of the current
%%   element. \hyperlink{track:remarks:aperture:notes}{Notes}. & .FALSE.=
%%   no aperture check & logical \\  
%%   \hline
%%   ONETABLE & write all particle coordinates in a single file & .FALSE.=
%%   one file per particle & logical \\  
%%   \hline
%%   RECLOSS & create a table named "trackloss" in memory with lost
%%   particles' coordinates & .FALSE.= no table & logical \\  
%%   \hline
%%   FILE & name for the track table   & "track", "trackone" & string \\ 
%%   \hline
%%   UPDATE & parameter update per turn   & .FALSE.= no update & string \\  
%%   \hline
%% \end{tabular}

          

\section{START}          
\label{sec:start}

After the \texttt{TRACK} command, initial trajectory coordinates must be
provided for each trajectory or particle to be tracked, with one
{\texttt START} command per trajectory or particle.

The coordinates can be expressed as either
\href{../Introduction/tables.html#canon}{\textbf{canonical}}
or \textbf{action-angle} coordinates.

\madbox{
START, X=real, PX=real, Y=real, PY=real, T=real, PT=real;  \\
START, FX=real, PHIX=real, FY=real, PHIY=real, FT=real, PHIT=real;
}
          
For the case of action-angle coordinates, the normalised amplitudes are
expressed in number of r.m.s. beam size F$_X$, F$_Y$, F$_T$ (the actions being computed with
the emittances given in the \texttt{BEAM} command) \textbf{in each mode
  plane}. The phases are PHI$_X$, PHI$_Y$ and PHI$_T$ expressed in
radian. In the uncoupled case, we have in the plane mode labelled z, and
with E$_z$ being the r.m.s. emittance in that plane:\\
\\
Z = F$_z$ sqrt(E$_z$) cos(PHI$_z$),    \\
\\
P$_z$= F$_z$ sqrt(E$_z$) sin(PHI$_z$)


The attributes of the START command are:
\begin{madlist}
  \ttitem{X, PX, Y, PY, T, PT} canonical coordinates. 
  \ttitem{FX, FY, FT, PHIX, PHIY, PHIT} action-angle coordinates.
\end{madlist}

%% {\bf Options}
%% \begin{tabular}{p{3cm} cccc}
%%    \hline 
%%    \textbf{Option} & \textbf{Meaning} & \textbf{Default Value} & \textbf{Value Type} & \textbf{Unit} \\ 
%%    \hline
%%    X, PX, Y, PY, T, PT & canonical coordinates & 0.0 & double & m \\ 
%%    \hline
%%    FX, PHIX, FY, PHIY, FT, PHIT & action-angle coordinates & 0.0 & double & rad \\ 
%%    \hline
%% \end{tabular}

{\bf Remarks} \\
For usual tracking (single/multi-turn), all coordinates are specified
with respect to the actual closed orbit (possibly off-momentum, with
magnet errors) and \textbf{NOT} with respect to the reference orbit.  

If the option \texttt{onepass} of the \texttt{TRACK} is used, the
coordinates are specified with respect to the reference orbit. 
            


\section{OBSERVE}
\label{sec:observe}

During the tracking process, particle coordinates at specific named
locations along the machine can be printed to file(s). The declaration of
an observation point is with the OBSERVE command: 

\madbox{
OBSERVE, PLACE=string;  
}

The single attirbute of \texttt{OBSERVE} is:
\begin{madlist}
  \ttitem{PLACE} the name of the observation point. 
\end{madlist}

     
%% {\bf Options} 
%% \begin{tabular}{cccc}
%%    \hline 
%%    \textbf{Option} & \textbf{Meaning} & \textbf{Default Value} & \textbf{Value Type} \\ 
%%    \hline
%%    PLACE & name of the observation point &  & string \\ 
%%    \hline   
%% \end{tabular}

Several \texttt{OBSERVE} commands can be given for the same tracking
job, one per observation point. 

If no \texttt{OBSERVE} command is given in a tracking job, but the
\texttt{DUMP} option in the \texttt{TRACK} command is used, the
trajectory coordinates are still recorded and one observation point is
provided at the starting point of the sequence. 
     
The output files are named automatically. The name given by
the user (attribute \texttt{FILE} of the \texttt{TRACK} command) is
followed by ".obsnnnn", where nnnn is the observation point number, and followed by 
".pnnnn"  wherer nnnn is now the particle number. Hence the default
filename for the first obseration point and first particle looks like
\texttt{track.obs0001.p0001}.


\section{RUN}
\label{sec:run}

The actual tracking is triggerred by the \texttt{RUN} command. 

\madbox{
RUN, TURNS=integer, MAXAPER=real\_array, FFILE=integer;  
}

The \texttt{RUN} command has three attributes:  
     
\begin{madlist}
  \ttitem{TURNS} number of turns to be tracked.
  \ttitem{MAXAPER} upper limits for the six coordinates.\\
  (Default: \{0.1, 0.01, 0.1, 0.01, 1.0, 0.1\} \\
  The limits defined by the \texttt{maxaper} option are only being taken
  into account if the \texttt{aperture} option of the \texttt{TRACK}
  command is used. 
  \ttitem{FFILE} defines the turn periodicity for printing coordinates at
   observation points. (Default:~1)\\
   \texttt{FFILE=n} will print coordinates every n-th turn only. 
\end{madlist}


%%\title{DYNAP}
%  Changed by: Hans Grote, 17-Jun-2002 
%  Changed by: Frank Zimmermann, 18-Jun-2002 
%  Inserted in THINTRACK by ghislain, 2014-Aug-07  14:43:45  

\section{DYNAP}

The \texttt{DYNAP} command calculates tunes, tune footprints, smear and
Lyapunov exponent from trackin data. \texttt{DYNAP} can be called
instead of \texttt{RUN} inside a \texttt{TRACK} command environment.

\madbox{
DYNAP, \=TURNS=integer, FASTUNE=logical, LYAPUNOV=real,\\
       \>MAXAPER=real\_array, ORBIT=logical;
}
 
For each previously entered start command, \texttt{DYNAP} tracks two
close-by particles over a selected number of turns, from which it
obtains the betatron tunes with error, the action smear, and an estimate
of the lyapunov exponent. 
Many such companion particle-pairs can be tracked at the same time,
which speeds up the calculation. 

The \textit{smear} is 
defined as  \textit{2.0 (wxy$_{max}$ - wxy$_{min}$) / (wxy$_{max}$ + wxy
  $_{min}$)}, where the \textit{wxy$_{min,max}$} refer to the  miminum and
maximum values of the sum of the transverse betatron invariants
\textit{wx+wy} during the tracking. 

The tunes are computed by using an FFT and either formula (18) or formula (25) of CERN SL/95-84 (AP),
depending on whether the number of turns is less-equal or larger than 64. 
 
\texttt{DYNAP} has the following attributes: 
\begin{madlist}
   \ttitem{TURNS} the number of turns to be tracked (Default:~64,
   present maximum:~1024). 
     
   \ttitem{FASTUNE} a logical flag to compute the tunes. (Default:~false)
 
   \ttitem{MAXAPER} a vector of 6 real numbers defining the maximum
   aperture beyond which the particle is considered to be lost.\\
   (Default: \{0.1, 0.01, 0.1, 0.01, 1.0, 0.1\} \\
     
   \ttitem{LYAPUNOV} the initial distance which is added to the
   \textit{x} coordinate of the companion particle of every particle
   declared with \texttt{START} commands. (Default:~1.e-7~m)
     
   \ttitem{ORBIT} A logical flag. If set, the flag \textit{orbit} 
   is true during the tracking and its initialization
   (default: true).
   \textbf{ This flag should be set to be true, if 
     normalized coordinates are to be entered.}
\end{madlist}

%% Example:
%% \begin{verbatim}
%% BEAM,PARTICLE=ELECTRON,ENERGY=50,EX=1.E-6,EY=1.E-8,ET=0.002,SIGT=1.E-2;
%% ...
%% USE,PERIOD=FODO;
%% ...
%% TRACK;
%% START,X=0.0010,Y=0.0017,PT=0.0003;
%% DYNAP,FASTUNE,TURNS=1024,LYAPUNOV=1.e-7;
%% ENDTRACK;
%% ...
%% \end{verbatim}

The first command defines the beam parameters. It is  essential that the
longitudinal emittance \textit{ET} is set. The command \textit{use}
selects the beam line or sequence. The \textit{track} activates the
tracking module, \textit{start } enters the starting coordinates (more
than one particle can be defined),  \textit{dynap} finally tracks two
nearby particles  with an initial distance \textit{lyapunov}  for each
\textit{start} definition over \textit{ turns } revolutions, and
\textit{endtrack} terminates the execution of the tracking module. 

The results are stored in the \textit{DYNAP} and \textit{DYNAPTUNE}
tables, and can be obtained by the two commands  
 
\madxmp{
VALUE, \=TABLE(dynap,smear); \\
VALUE, \>TABLE(dynaptune,tunx), \\
       \>TABLE(dynaptune,tuny), \\
       \>TABLE(dynaptune,dtune);
}

More generally, all results can be printed to a file, using the commands 
\madxmp{
WRITE, TABLE=dynap, file; \\
WRITE, TABLE=dynaptune, file;
}
The output file 'lyapunov.data' lists the turn number and phase distance
between the  two Lyapunov partners, respectively, allowing for visual
inspection of chaoticity.  
 
%\href{http://consult.cern.ch/xwho/people/62690}{frankz} 20.03.2006



\section{ENDTRACK}
\label{sec:endtrack}

Tracking is terminated by the command \texttt{ENDTRACK} with no
attributes. 

\madbox{ENDTRACK;}



%\href{http://consult.cern.ch/xwho/people/74251}{A. Koschik},  February  2007

%% EOF
