%%\title{PTC\_NORMAL Module (Non-Linear Machine Parameters)}
%  Created by: Valery KAPIN, 21-Mar-2006 


\section{PTC\_NORMAL Module: Non-Linear Machine Parameters}

The \textbf{PTC\_NORMAL module} of MAD-X
[\hyperlink{F._Schmidt}{a},\hyperlink{d Amico}{b}] is based on PTC
code. This module takes full advantage of the PTC Normal Form analysis
which is  a  considerable upgrade of what was available with
the Lie Algebra   technique used in MAD8. It allows to calculate
dispersions, chromaticities,   anharmonicities and Hamiltonian terms to
very high order. In fact, the order is   only limited by the RAM memory
of your computer and your patience to wait for   the results.  

The number of terms per order increases with some power law. The
internal MAD-X   tables are not adequate to keep such large amounts of
data. On the other hand,   only a reduced set of this data is actually
needed by the user. Thus a much   easier and flexible solution is to
gather the users  requirements with a series   of special MAD-X command
called \texttt{SELECT\_PTC\_NORMAL}. A   special MAD-X table is
dynamically built using just those commands and it will   be filled by
the next call to the \texttt{PTC\_NORMAL}-command.  

Another essential advantage of this table is the fact that it is
structured to   facilitate exchange of Normal Form (including
Hamiltonian terms of high order)   between MAD-X modules. The immediate
goal is to use this table to allow   non-linear matching inside the
present MAD-X MATCHING module. 


\begin{description}
   \item[Synopsis]
     \text{ \\}
     \textit{
       \texttt{PTC\_CREATE\_UNIVERSE;}}\\
     \textit{\texttt{PTC\_CREATE\_LAYOUT, model=integer,method=integer,  nst=integer, [exact];}}\\
     \texttt{..............................\\
       SELECT\_PTC\_NORMAL, dx,..., gnfu;}\\
     \texttt{..............................}\\
     \texttt{PTC\_NORMAL;\\
       WRITE, table=normal\_results, file=normal\_results;}\\
     \texttt{..............................}\\
     \textit{\texttt{PTC\_END; }}\\
\end{description}

\begin{description}
   \item[Commands] 

   \item {\texttt{\textbf{\textbf{SELECT\_PTC\_NORMAL, }}}\\
     \textbf{\textbf{dx}=integer\textbf{, dpx}=integer\textbf{, dy}=integer\textbf{, dpy}=integer\textbf{,\\
         q1}=0\textbf{, dq1}=integer\textbf{, q2}=0\textbf{, dq2}=integer\textbf{,         
         \\ anhx}=integer array, \textbf{anhy}=integer array\textbf{, 
         \\ gnfu}=integer,0,0\textbf{, haml}=integer,0,0\textbf{,
         \\ eign}=integer, integer\textbf{;}}}
     
   \item[Description] 
     % <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br /> 
     % <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
     The \texttt{SELECT\_PTC\_NORMAL }command selects parameters to be
     calculated by the next \texttt{PTC\_NORMAL }command. The dispersion
     and tune parameters are defined by a name and an integer number
     specifying their order. For example, the notation "dx=2" means the
     horizontal dispersion to second order
     \textit{D$_x$}$^{(2)}$=$\partial$$^{(2)}$\textit{x}$_{co}$/$\partial$\textit{$\delta$}$_\textit{p}$$^{(2)}$, 
     where "co" is abbreviation of "closed orbit". The anharmonisities
     are defined by a name and three integer numbers: the first is the
     order of \textit{$\epsilon$}$_1$, the second is the order of
     \textit{$\epsilon$}$_2$, the third is the order of
     \textit{\nolinebreak$\delta$}$_\textit{p}$. For example, the
     notation "anhx=2,0,0" means second order in
     \textit{$\epsilon$}$_1$:
     $\partial$$^{(2)}$\textit{q}$_1$/$\partial$\textit{$\epsilon$}$_1$$^{(2)}$.  \\       
     Components of the eigenvectors at the end of the structure can be
     specified by two integers: the first integer defines the
     eigenvector number, the second integer defines the coordinate
     \{\textit{x}, \textit{p$_x$}, \textit{y}, \textit{p$_y$},
     \textit{t},\textit{p$_t$}\}. \\
     The Generating Function can be specified by \{ \textit{n}, 0,
     0\}. The positive and negative values of \textit{n} define the
     order of upright or skew resonances, respectively. The integers
     \textit{n}$_2$ and \textit{n}$_3$ are reserved for a future
     upgrade. For example, "gnfu=-5, 0, 0" will calculate all Generating
     Function terms for skew decapoles. In the output table, one finds
     the cosine, sine and amplitude coefficients as denoted by "GNFC",
     "GNFS", and "GNFA", respectively. \\
     Similarily, the Hamiltonian terms can be specified by \{
     \textit{n}, 0, 0\}. The positive and negative values of \textit{n}
     define the order of upright or skew resonances, respectively. For
     example, "haml=3, 0, 0" will calculate all Hamiltonian terms for
     upright sextupoles. In the output table, one finds the cosine, sine
     and amplitude coefficients as denoted by "HAMC", "HAMS", and
     "HAMA", respectively. 
\end{description}

\begin{description}
   \item[Parameters] 
     \begin{tabular}{p{2cm} p{7cm} p{2cm}}
       \hline 
       \textbf{Notation} & \textbf{Meaning} & \textbf{Value} \\ 
       \hline
       DX, DPX, DY,DPY & dispersions, \textit{D$_x$}$^{(n)}$,
       \textit{D$_px$}$^{(n)}$, \textit{D$_y$}$^{(n)}$,
       \textit{D$_py$}$^{(n)}$ &  
       \textit{n} \\ 
       \hline
       Q1, Q2 & horizontal and vertical tunes \textit{q}$_1$$^{(0)}$,
       \textit{q}$_2$$^{(0)}$ & 0 \\  
       \hline
       DQ1, DQ2 & derivatives of horizontal and vertical tunes
       $\partial$$^{(n)}$\textit{q}$_1$/$\partial$\textit{$\delta$}$_\textit{p}$$^{(n)}$,
       $\partial$$^{(n)}$\textit{q}$_2$/$\partial$\textit{$\delta$}$_\textit{p}$$^{(n)}$
       & \textit{n} \\  
       \hline
       ANHX, ANHY & Anharmonicities &
       \textit{n}(\textit{$\epsilon$}$_1$),
       \textit{n}(\textit{$\epsilon$}$_2$),
       \textit{n}(\textit{$\delta$}$_\textit{p}$) \\  
       \hline
       GNFU & Generating Function & \textit{n}, 0, 0 \\ 
       \hline
       HAML & Hamiltonian  & \textit{n}, 0, 0 \\ 
       \hline
       EIGN & Eigenvector (the \textit{n}$_2$-th component of the
       \textit{n}$_1$-th eigenvector) & \textit{n}$_1$, \textit{n}$_2$
       \\  
       \hline
     \end{tabular}
\end{description}

\begin{description}
    \item {\texttt{\textbf{\textbf{PTC\_NORMAL, 
            \\   icase=}integer, }}\textbf{\texttt{normal}, \textbf{closed\_orbit}, \textbf{
          \\   no}=integer,         \textbf{map\_table, deltap}=double\textbf{;}}}

    \item[Description] 
      % <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br /> 
      % <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
      The calculation of the parameters specified by the preceding
      \texttt{SELECT\_PTC\_NORMAL } commands is initiated by the
      \texttt{PTC\_NORMAL }command, which operates on the working beam
      line defined in the latest \href{../control/general.html#use}{
        USE} command. The options for \texttt{PTC\_NORMAL }command are
      described in the table below.  

    \item[Options]
      \begin{tabular}{l p{5cm} p{2cm} p{2cm}}
        \hline 
        \textbf{Option} & \textbf{Meaning} & \textbf{Default Value} & \textbf{Value Type} \\ 
        \hline
        ICASE & the user-defined dimensionality of the phase-space (4, 5 or 6) & 4 & integer \\ 
        \hline
        NO & the order of the map. & 1 & integer \\ 
        \hline
        CLOSED\_ORBIT & the switch to turn on the closed orbit calculation. & .FALSE. & logical \\ 
        \hline
        DELTAP & relative momentum offset for reference closed orbit & 0.0 & double \\ 
        \hline
        MAPTABLE & turn on the map-table in memory & .FALSE. & logical \\ 
        \hline
        NORMAL & turn on the calculation of the Normal Form & .FALSE. & logical \\ 
        \hline
      \end{tabular}

    \item[Remarks] 
      \textbf{MAPTABLE}:\nolinebreak(requires no=1) creates the
      one-turn matrix which can be used by the next
      \href{../ptc_twiss/ptc_twiss.html}{PTC\_TWISS} command. 

    \item[Example]
      The simple example is located on the Web-page for the
      \href{http://cern.ch/frs/mad-X_examples/ptc_normal}{\texttt{PTC\_NORMAL}
        example}. 
\end{description}


\begin{description}
\item[References for PTC\_NORMAL] 
\end{description}

\begin{enumerate}
\item \href{F._Schmidt}{F. Schmidt},
  "`\href{http://cern.ch/madx/doc/MPPE012.pdf}{MAD-X PTC Integration}'',
  Proc. of the 2005 PAC Conference in Knoxville, USA, pp.1272. 
\item E.T. \href{d Amico}{d Amico}, "Nonlinear parameters from PTC",
  \href{http://cern.ch/frs/MAD-X_minutes/Meeting-7}{MAD-X Meeting 7
    (29.11.2004)}, notes
  (\href{http://cern.ch/frs/MAD-X_minutes/Meeting-7/Nonlinear_parameters_from_PTC.doc}{doc}-file). 
\item \href{A._Schoch}{A. Schoch}, "Theory of linear and non-linear
  perturbations of betatron oscillations in alternating-gradient
  synchrotrons ",
  \href{http://cern.ch/madx/doc/yellow-report-1957.pdf}{CERN-27-21},
  1958. 
\end{enumerate}

\begin{description}
\item[See Also]
  \href{http://cern.ch/frs/mad-X_examples/ptc_normal}{\texttt{PTC\_NORMAL}
    example}, \href{../ptc_general/ptc_general.html}{PTC Set-up
    Parameters}. 
\end{description}



%\href{mailto:kapin@itep.ru}{  V.Kapin}(ITEP) and 
%\href{mailto:Frank.Schmidt@cern.ch}{  F.Schmidt}, March  2006

