%%\title{Matching Module}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 
 
\chapter{Matching Module}
\label{chap:match}

Before a match operation at least one sequence must be selected by means
of a \href{../control/general.html#use}{USE} command. Matching is 
initiated by the \href{match_main.html}{MATCH} command. The matching
module can act on more than one sequence simultaneously by  specifying
more than one sequence when  \href{match_main.html#match}{initiating}
the matching mode.  From this command to the corresponding
\href{match_main.html#endmatch}{ENDMATCH} command MAD accepts the
matching commands listed below. For a mathematical description of the
minimisation procedures see \cite{MINUIT}. 

In particular one may do the following: 
\begin{itemize}
 \item Define the sequence(s) the matching module will work on 
 \item Set initial conditions for transfer line matching 
 \item Define constraints 
 \item Define the parameters to be varied 
 \item Match by different methods. 
\end{itemize}

The matching commands are described in detail below. Some other commands
can also be issued during matching.  

\begin{itemize}
\item \href{match_main.html}{Enter and Leave Matching Mode}
  \begin{itemize}
  \item \href{match_main.html#match}{MATCH: Initiating the Matching Mode}
  \item \href{match_main.html#endmatch}{ENDMATCH: Leave Matching Mode}
  \end{itemize}
\item \href{match_vary.html}{Define Variable Parameter}
  \begin{itemize}
  \item \href{match_vary.html#vary}{VARY: Set Parameter to Vary}
  \end{itemize}
\item \href{match_con.html}{Constraint}
  \begin{itemize}
  \item \href{match_con.html#constraint}{CONSTRAINT}: Simple Constraint 
  \item \href{match_con.html#user-var}{CONSTRAINT}: User Defined Variables 
  \item \href{match_con.html#weight}{WEIGHT: Matching Weights}
  \item \href{match_con.html#global}{GLOBAL: Global Constraints}
  \item \href{match_con.html#global}{GWEIGHT: Weights for Global Constraints}
  \end{itemize}
\item \href{match_xeq.html}{Matching Methods}
  \begin{itemize}
  \item \href{match_xeq.html#lmdif}{LMDIF: Fast Gradient Minimisation}
  \item \href{match_xeq.html#migrad}{MIGRAD: Gradient Minimisation}
  \item \href{match_xeq.html#simplex}{SIMPLEX: Simplex Minimisation}
  \item \href{match_xeq.html#jacobian}{JACOBIAN: Newton Minimisation}
  \end{itemize}
\item \href{match_um.html}{Expression Matching with USE\_MACRO}
\item \href{match_xmpl.html}{Matching Examples}
\end{itemize}

%\line(1,0){300}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002; 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006. 

% add other files to the end of this file


%\input{match/match_main}
%%\title{MATCH / ENDMATCH}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 
%  Changed by: Riccardo de Maria, 9-Jan-2008 


Before matching at least one
\href{../Introduction/sequence.html}{SEQUENCE} must be selected by means
of a \href{../control/general.html#use}{USE} command. The matching
module can act on more than one sequence simultaneously by specifying
more than one sequence in the \texttt{MATCH} command.
 

\section{MATCH \ldots ENDMATCH}
\label{sec:match}\label{sec:endmatch}

The \texttt{MATCH} command is used for matching either a periodic cell
or an insertion to another part of the machine. 

Both matching modes are initiated by the \texttt{MATCH} command using
one of several forms:

\madbox{
MATCH, SEQUENCE='name1', 'name2', \ldots , 'name-n';
}

\madbox{
MATCH, \=SEQUENCE= 'name1', 'name2', \ldots , 'name-n',\\
       \>BETA0= 'beta01', 'beta02', \ldots , 'beta0n'; \\
\\
or\\
\\
MATCH, \=SEQUENCE='seqname', \\ 
       \>BETX=real, ALFX=real, MUX=real, \\
       \>BETY=real, ALFY=real, MUY=real, \\
       \>X=real, PX=real, Y=real, PY=real, \\
       \>DX=real, DY=real, DPX=real, DPY=real, \\
       \>DELTAP=real;
}


\begin{itemize}
  \item Cell matching:
    In the first mode the matching routine is initiated only with one
    attribute specifying the sequence(s) the matching module will work
    on. In this matching mode the periodicity of the optics functions is
    enforced at the beginning and end of the selected range. 
    \madxmp{MATCH, SEQUENCE='name1', 'name2', \ldots ,'name-n';}
    
  \item Insertion matching:
    In the second mode, called insertion matching, the matching routine is
    initiated with two attributes: one specifying the sequence(s) the
    matching module will work on and one specifying the initial conditions
    of the optic functions for each sequence. In this case the initial
    values are assumed as exact. 
 
    \begin{itemize}
      \item Specification of Initial Values: The initial values of the optical
        functions  for the insertion matching can either be specified in form of
        a \href{../control/general.html#savebeta}{SAVEBETA} command or by
        explicitly stating the individual optic function values after the
        'MATCH' command. The two options can be entered as         
        \madxmp{
          MATCH, \=SEQUENCE= 'name1', 'name2',.., 'name-n',\\
                 \>BETA0= 'beta01', 'beta02',..., 'beta0n';
        }
        or
        \madxmp{
          MATCH, \=SEQUENCE='seqname', \\ 
                 \>BETX=real, ALFX=real, MUX=real, \\
                 \>BETY=real, ALFY=real, MUY=real, \\
                 \>X=real, PX=real, Y=real, PY=real, \\
                 \>DX=real, DY=real, DPX=real, DPY=real, \\
                 \>DELTAP=real;
        }

{\bf Examples:}
 
Example 1:
\begin{verbatim}
CELL: SEQUENCE=(...) ;
INSERT: SEQUENCE=(...) ;
USE,PERIOD=cell;
SAVEBETA,LABEL=bini,place=#e;
TWISS,SEQUENCE=cell;
USE,PERIOD=insert;
MATCH,SEQUENCE=insert,BETA0=bini;
CONSTRAINT,SEQUENCE=insert,RANGE=#e,MUX=9.345,MUY=9.876;
\end{verbatim}
This matches the sequence 'INSERT' with initial conditions to a new
phase advance. The initial conditions are given by the periodic solution
for the sequence CELL1. 

Example 2:	
\begin{verbatim}
USE,PERIOD=INSERT;
MATCH,SEQUENCE=insert;
CONSTRAINT,SEQUENCE=insert,RANGE=#e,MUX=9.345,MUY=9.876;
\end{verbatim}
This matches the beam line 'INSERT' with periodic boundary conditions to
a new phase advance. 


 
The initial conditions can also be transmitted by a combination of a
\href{../control/general.html#savebeta}{SAVEBETA} command and explicit
optic function specifications: 
\begin{verbatim}
USE,CELL1;
SAVEBETA,LABEL=bini,PLACE=#E;
TWISS,SEQUENCE=CELL1;
USE,PERIOD=LINE1;
MATCH,SEQUENCE=LINE1,BETA0=bini,MUX=1.234,MUY=4.567;
\end{verbatim}

This example transmits all values of the SAVEBETA array 'bini' as
initial values to the MATCH command and overrides the initial phase
values by the given values.

\end{itemize}

An additional \href{match_con.html#constraint}{CONSTRAINT} may be
imposed in other places, i.e. intermediate or end values of the optics
functions at the transition point.  
 
\item More than one active sequence:

The matching module can act on more than one sequence simultaneously by
specifying more than one sequence after the MATCH command: 
\begin{verbatim}
MATCH, SEQUENCE=LINE1, CELL1, BETA0=bini1, bini2;
\end{verbatim}
This example initiates the matching mode for the 'LINE1' and the 'CELL1'
sequence. The \href{../twiss/twiss.html}{Twiss module} function of the
two sequences are calculated with fixed initial conditions. The SAVEBETA
array 'bini1' is used for calculating the optics functions of sequence
'LINE1' and the SAVEBETA array 'bini2' for calculating the optics
functions of sequence 'CELL1'. Without the initial conditions the
matching module will work in the \href{match_main.html#cell}{CELL}
mode. 
 
	\item  Special flag:

The "slow" attribute enforces the old and slow matching procedure which
allows to use the special columns \texttt{mvar1, ..., mvar4}, if they
are added to the twiss table. Recently a number of parameter, like
"RE56", have been added to list of matchable parameters. Nevertheless,
some parameters might only be available when using the "slow" attribute. 
 
\end{itemize}
 
\subsection{Useful TWISS attributes}

\begin{itemize}
\item RMATRIX: If this flag is used the one-turn map at the location of every
  element is calculated and prepared for storage in the TWISS table.\\ 
  Target values for the matrix elements at certain positions in the sequence
  can be specified with the help of the \href{match_con.html#constraint}{CONSTRAINT}
  command and the
  keywords: \textbf{RE, RE11...RE16...RE61...RE66}, where \textbf{REij} refers
  to the "ij" matrix component.
  
  \textgreater Examples:
 
\begin{verbatim}
     MATCH, RMATRIX, SEQUENCE='name',BETA0='beta-block-name';
     CONSTRAINT, SEQUENCE=insert, RANGE=#e, RE11=-2.808058321, RE22=2.748111197;
     VARY, NAME=kqf, STEP=1.0e-6;
     VARY, NAME=kqd, STEP=1.0e-6;
\end{verbatim}

  This matches the sequence 'name' with initial conditions to new values
  for the matrix elements 'RE11' and 'RE22' by varying the strength of the
  main quadrupole circuits.

\item CHROM: If this flag is used the chromatic functions at the location of
  every element are calculated and prepared for storage in the TWISS
  table. 
  
  Target values for the chromatic functions at certain positions in the
  sequence can be specified with the help of the
  \href{match_con.html#constraint}{CONSTRAINT} command and the keywords
  \href{../Introduction/tables.html#normal}{WX, PHIX, WY, PHIY,...}. 
\end{itemize}

 
The ENDMATCH command terminates the matching section and deletes all
tables related to the matching run. 
\madbox{
ENDMATCH;
}



%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, October, 2003;
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, January, 2008.


%\input{match/match_vary}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 19-Jun-2002 
 
\section{VARY}
\label{sec:match_vary}
A parameter to be varied is specified by the command 

\madbox{
VARY, \=NAME=variable, STEP=real, LOWER=real, UPPER=real \\
      \>SLOPE=integer, OPT=??;
}

It has six attributes: 
\begin{madlist}
  \ttitem{NAME} The \href{../Introduction/variable.html}{name of the
    parameter or attribute} to be varied,  

  \ttitem{STEP} The approximate initial step size for varying the
  parameter. If the step is not entered, MAD tries to find a
  reasonable step, but this may not always work.  

  \ttitem{LOWER} Lower limit for the parameter (optional), 
  \ttitem{UPPER} Upper limit for the parameter (optional). 

  \ttitem{SLOPE} allowed change rate (optional, available only using
  \href{match_xeq.html#jacobian}{JACOBIAN} routine). Limit the
  parameter to increase (SLOPE=1) decrease (SLOPE=-1) only.  

  \ttitem{OPT} optimal value for the parameter (optional, available
  only using \href{match_xeq.html#jacobian}{JACOBIAN} routine).  
\end{madlist}

{\bf Examples:}
\madxmp{
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
VARY, NAME=PAR1, STEP=1.0E-4;          \>! vary global parameter PAR1 \\
VARY, NAME=QL11->K1, STEP=1.0E-6;      \>! vary attribute K1 of the QL11 \\ 
VARY, NAME=Q15->K1, STEP=0.0001, LOWER=0.0, UPPER=0.08; ! vary with limits
}

If the upper limit is smaller than the lower limit, the two limits are
interchanged. If the current value is outside the range defined by the
limits, it is brought back to range. 

If a parameter comes outside the
limits during the matching process the matching module resets the
parameter to a value inside the limits and informs the user with a
message. If such a 'rescaling' occurs more than 20 times the matching
process terminates. The user should either eliminate the corresponding
parameters from the list of varied parameters or change the
corresponding upper and lower limits before restarting the matching
process. 

After a matching operation all varied attributes retain their
value after the last successful matching iteration. Using
\href{match_xeq.html#jacobian}{JACOBIAN} routine, STRATEGY=3, in case
the number of parameters is greater than the number of constraint, if a
parameter comes outside the limits, it is excluded automatically from
the set of variables and a new solution is searched.  

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002. 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006.



%\input{match/match_con}
%%\title{Matching Constraints}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 25-Sep-2002 
% Changed by: Ghislain Roy, 11 Dec 2013: removed duplicate paragraph in user defined matching constraint 

\section{CONSTRAINT}
\label{sec:constaint}

Simple constraints are imposed by the CONSTRAINT command. The CONSTRAINT
command has three attributes:   
\begin{itemize}
 \item  the SEQUENCE entry specifies the sequence for which the
   constraint applies.  
 \item  the RANGE entry specifies the position where the
   constraint must be satisfied. The RANGE can either be the name
   of a single element in the sequence or a range between two
   elements. In the later case the two element names must be
   separated by a '/': RANGE=nam1/name2  
 \item the optics functions to be constrained. 
\end{itemize} 

The optic functions can be constrained in four different ways: 
\begin{itemize}
 \item lower limits: 'BETX $>$ value' -\textgreater type1 
 \item upper limits: 'BETX $<$ value' -\textgreater type2 
 \item lower and upper limits: 'BETX $<$ value1, 
   BETX $>$ value2' -\textgreater type3  
 \item target value: BETX=value -\textgreater type 4 
\end{itemize} 

In case one element is affected by more than one constraint command the
last CONSTRAINT will be chosen. For example, one can specify the
maximum acceptable beta function over a range of the sequence and
specify the target beta function for one element that lies inside this
range. In this case one must first specify the constraint that affects
the whole range and then the constraint for the single element. This way
the constraint of the target value overrides the previous constraint on
the upper limit for the selected element. 

For example, the following
constraint statements limit the maximum horizontal beta function between
'marker1' and 'marker2' to 200 meter and require a horizontal beta
function of 100 meter at element 'name1':  
\madxmp{
CONSTRAINT, SEQUENCE=seqname, RANGE='marker1'/'marker2', BETX<200.0;\\
CONSTRAINT, SEQUENCE=seqname, RANGE='name1'/'marker2', BETX=100.0;
}

When the two constraint statements are interchanged, and supposing that
name1 is an element in the range marker1/marker2, the horizontal beta 
function at element 'name1' will only be limited to less than 200 meter
and NOT constrained to 100 meter! 

The CONSTRAINTS can either be specified with explicit values for the
constraints of the optic functions or via a pre-calculated
\href{../control/general.html#savebeta}{SAVEBETA} module. The first
options has the form: 
\madbox{
CONSTRAINT, \=SEQUENCE=seqname, RANGE=position,\\
            \>BETX=real, ALFX=real, MUX=real, \\
            \>BETY=real, ALFY=real, MUY=real, \\
            \>X=real, PX=real, Y=real, PY=real, \\
            \>DX=real, DY=real, DPX=real, DPY=real;
}

Here all \href{../Introduction/tables.html#linear}{linear lattice functions} 
(BETX, BETY, ALFX, ALFY, MUX, MUY, DX, DY, DPX, DPY)
or \href{../Introduction/tables.html#chrom}{chromatic lattice functions}
(WX, XY, PHIX, PHIY, DMUX, DUMY, DDX, DDY, DDPX, DDPY)
are constrained at the selected range to the corresponding values.

The second form of the CONSTRAINT command has the form
\madbox{
CONSTRAINT, \=SEQUENCE=seqname, RANGE=position, \\
            \>BETA0=beta0-name, MUX=real, MUY=real;
}

Here all of (BETX, BETY, ALFX, ALFY, MUX, MUY, DX, DY, DPX, DPY)
are constrained in the selected points to the corresponding values
of a pre-calculated \href{../control/general.html#savebeta}{SAVEBETA} module.
In the above example
the phases (MUX, MUY) are overridden by the numerical values specified via
'MUX=real' and 'MUY=real'.
Normally ``RANGE'' refers to a single
\href{../Introduction/ranges.html#position}{position}.

\section{User Defined Matching Constraints}
\label{sec:userconstraint}

In addition to the nominal TWISS variables the user can define a limited
set of 'user-defined' variables in the constraint statement. This
allows, for example, the matching of the normalized dispersion or the
mechanical aperture. The MATCH module allows four user defined variables
called: mvar1, mvar2, mvar3 and mvar4. 
The variables can be defined according to the general variable
declaration rules of  
\href{../Introduction/expression.html#defer}{deferred expressions}.
For example, in order to match the normalized dispersion at a certain
location in the sequence one would first define a variable:

\madxmp{
mvar1 := table(twiss,dx)/sqrt(table(twiss,betx));
}

After that the user has to select the variable for output in the TWISS
statement (see \href{../twiss/twiss.html}{TWISS module} and
\href{../Introduction/select.html}{SELECT} for more
details on the TWISS module and SELECTION statements):
\madxmp{
SELECT, FLAG=twiss, CLEAR; \\
SELECT, FLAG=twiss, COLUMN=keyword, name, s, betx, dx, mvar1; \\
TWISS, SEQUENCE=seqname, FILE=twissfile;
}

The variable can now be referenced like any other TWISS variable
in the constraint command:
\madxmp{
CONSTRAINT, SEQUENCE=seqname, RANGE=range, MVAR1=targetvalue;
}

\section{GLOBAL}
\label{sec:global}

In addition to conventional matching constraints that specify the optics 
functions at a certain position in the sequence the user can also constrain 
global optics parameters such as, for example, the overall machine tune
and the machine chromaticity. Such global optics parameters can be
constraint via the  GLOBAL command, having the following syntax:

\madbox{
GLOBAL, \=SEQUENCE=sename, \\
        \>Q1=real, dQ1=real, ddQ1=real, \\
        \>Q2=real, dQ2=real, ddQ2=real;
}


All attributes are optional and have the following meaning:
\begin{madlist}
  \ttitem{Q1, Q2} enable a correction of tunes in presence of
  magnetic imperfections or misalignments, 
  \ttitem{dQ1, dQ2}
  enable a correction of chromaticities in presence of
  magnetic imperfections or misalignments, 
  \ttitem {ddQ1, ddQ2} enable a correction of nonlinear chromaticities
\end{madlist}


\section{WEIGHT, GWEIGHT}
\label{sec:weight}\label{sec:gweight}

The matching procedures try to fulfil the constraints
in a least square sense.
A penalty function is constructed which is the sum of the
squares of all errors,
each multiplied by the specified weight.
The larger the weight, the more important a constraint becomes.
The weights are taken from a table of current values.
These are initially set to default values listed in the table below,
and may be reset at any time to different values.
Values set in this way remain valid until changed again.

The \texttt{WEIGHT} command changes the weights for subsequent
constraints: 
\madbox{
WEIGHT, \=BETX=real,ALFX=real,MUX=real, \\
        \>BETY=real,ALFY=real,MUY=real, \\
        \>X=real,PX=real,Y=real,PY=real, \\
        \>DX=real,DPX=real,DY=real,DPY=real;
}

The weights are entered with the same name as the
\href{../Introduction/tables.html#linear}{linear lattice functions}
\href{../Introduction/closed_orbit.html}{orbit coordinate} 
to which the weight applies.
Frequently the matching weights serve to select a restricted
set of functions to be matched.


Matching weights for global matching constraints can be set by the 
\texttt{GWEIGHT} command, having attributes identical to those of
\texttt{GLOBAL}.

\madbox{
GWEIGHT, \=Q1=real, dQ1=real, ddQ1=real, \\
         \>Q2=real, dQ2=real, ddQ2=real;
}


Default values for matching weights are given in the table below.

%http://en.wikibooks.org/wiki/LaTeX/Tables#Resize_tables
%\resizebox{480px}{!}{
\begin{table}[ht]
  \begin{center}
    \caption{Default Matching Weights}
    \vspace{1ex}
    %\begin{tabular}{llllllllllll}
    \begin{tabular}{|lr|lr|lr|}
      \hline
      NAME   & WEIGHT & NAME   & WEIGHT & NAME   & WEIGHT \\
      \hline
      BETX   &   1.0  & ALFX   &  10.0  & MUX    &  10.0  \\
      BETY   &   1.0  & ALFY   &  10.0  & MUY    &  10.0  \\ 
      X      &  10.0  & PX     & 100.0  &  & \\
      Y      &  10.0  & PY     & 100.0  &  & \\
      T      &  10.0  & PT     & 100.0  &  & \\ 
      DX     &  10.0  & DPX    & 100.0  &  & \\
      DY     &  10.0  & DPY    & 100.0  &  & \\ 
      WX     &   1.0  & PHIX   &   1.0  & DMUX   &   1.0  \\
      WY     &   1.0  & PHIY   &   1.0  & DMUY   &   1.0  \\ 
      DDX    &   1.0  & DDPX   &   1.0  &  & \\
      DDY    &   1.0  & DDPY   &   1.0  &  & \\ 
      MVAR{\it i}  &  10.0  & & & & \\
      Q1   &   10.0  & dQ1   &  1.0  & ddQ1    &  0.1  \\
      Q2   &   10.0  & dQ2   &  1.0  & ddQ2    &  0.1  \\
      \hline
      %% name   & weight & name   & weight & name   & weight & name   & weight & name   & weight & name   & weight \\ 
      %% BETX   &   1.0  & ALFX   &  10.0  & MUX    &  10.0  & BETY   &   1.0  & ALFY   &  10.0  & MUY    &  10.0  \\ 
      %% X      &  10.0  & PX     & 100.0  & Y      &  10.0  & PY     & 100.0  & T      &   0.0  & PT     &   0.0  \\ 
      %% DX     &  10.0  & DPX    & 100.0  & DY     &  10.0  & DPY    & 100.0  & - \\ 
      %% WX     &  10.0  & PHIX   &  10.0  & DMUX   & 100.0  & WY     &  10.0  & PHIY   &  10.0  & DMUY   & 100.0  \\ 
      %% DDX    &  10.0  & DDPX   & 100.0  & DDY    &  10.0  & DDPY   & 100.0  & - \\ 
      %% MVAR1    & 10.0  & MVAR2    & 10.0  & MVAR3    & 10.0  & MVAR4    & 10.0  & -
    \end{tabular}
  \end{center}
\end{table}
%}


%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002


%\input{match/match_xeq}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 

\section{Matching Methods}
\label{sec:match_methods}

MADX currently supports four different matching algorithms: 

\subsection{LMDIF: Fast Gradient Minimisation}
\label{subsec:match_lmdif}
The LMDIF command minimises the sum of squares of the constraint
functions using their numerical derivatives. It is the fastest
minimisation method available in \madx.    
 
\madbox{
LMDIF, CALLS=integer, TOLERANCE=real;
}

The command has two attributes:  
\begin{madlist}
   \ttitem{CALLS} The maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} The desired tolerance for the minimum. 
   (Default:~1E-6)  
\end{madlist}

\subsection{MIGRAD: Gradient Minimisation}
\label{subsec:match_migrad}
The MIGRAD command minimises the penalty function using the numerical
derivatives of the sum of squares.  

\madbox{
MIGRAD, CALLS=integer, TOLERANCE=real, STRATEGY=integer;
}

The command has three attributes: 
\begin{madlist}
   \ttitem{CALLS} the maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} the desired tolerance for the minimum. 
   (Default:~1E-6)
   \ttitem{STRATEGY} specifies the strategy to be used. (Default:~1) \\
   Details are given in \cite{MINUIT}.  
\end{madlist} 

\subsection{SIMPLEX: Simplex Minimisation}
\label{subsec:match_simplex}
The SIMPLEX command minimises the penalty function by the simplex
method. Details are given in \cite{MINUIT}.

\madbox{
SIMPLEX, CALLS=integer, TOLERANCE=real;
}

The command has two attributes: 
\begin{madlist}
   \ttitem{CALLS} The maximum number of calls to the penalty
   function. (Default:~1000) 
   \ttitem{TOLERANCE} The desired tolerance for the minimum. 
   (Default:~1E-6)
\end{madlist} 

\subsection{JACOBIAN: Newton Minimisation}
\label{subsec:match_jacobian}
The JACOBIAN command minimises the penalty function calculating the
Jacobian and solving the linear problem. A $Q R$ or $L Q$  decomposition is
performed when the system is over or under-determined. Before starting
the matching routine two optional transformations (COOL and RANDOM) are
performed. 

\madbox{
JACOBIAN, \=CALLS=integer, TOLERANCE=real, REPEAT=integer, \\
          \>STRATEGY=integer, COOL=real, BALANCE=real, RANDOM=real;
}

The command has the folowing attributes: 
\begin{madlist}
  \ttitem{CALLS} The maximum number of calls to the penalty
  function. (Default:~30)  

  \ttitem{TOLERANCE} The desired tolerance for the minimum. 
  (Default:~1E-6)

  \ttitem{REPEAT} The number of call of the JACOBIAN routine. 
  (Default:~1) 

  \ttitem{BISEC} Selects the maximum number of iteratation used to
  determine the step length which reduces the penalty function during
  the main iteration. A large number (i.e. 6) reduces the probability
  to diverge from the solution, but increases the probability of being
  trapped in a local minimum.  

  \ttitem{STRATEGY} A code for the strategy to be used. (Default:~3)\\
  If STRATEGY=1 the routine resets the values of the variables which
  exceeds the limits. If STRATEGY=2 the routine print the Jacobian
  and exit without matching. If STRATEGY=3 the routine  disables the
  variables which exceeds the limits keeping however the number of
  variables greater or equal to the number of the constraints.  

  \ttitem{COOL, BALANCE} The factors which specify the following
  transformation: 
  \madxmp{
   xxx\=xxxxxxxxx\= \kill
   if \>"balance" >=0 \\
      \>newval = (1-cool)*oldval +  \& \\
      \>       \>cool*((1-balance)*maxval + balance*minval)\\
   else \\
      \>newval = (1-cool)*oldval + cool*optval
  }
  where \texttt{newval} is the new value after the transformation,
  \texttt{oldval} is the previous value, \texttt{maxval, minval,
    optval} are the maximum, minimum and optimal value of
   the variable specified in the \texttt{VARY} command. 

   \ttitem{RANDOM} The factors which specify the following transformation:
   \madxmp{newval = ( 1 + random * rand() ) * oldval}
   where \texttt{newval} is the new value after the transformation,
   \texttt{oldval} is the previous value, \texttt{rand()} is a stochastic
   variable with a uniform (-0.5,0.5) distribution.   
\end{madlist} 


%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002. 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006. 


%\input{match/match_um}
%%\title{Expression Matching with USE\_MACRO}
%  Changed by: Chris ISELIN, 27-Jan-1997 
%  Changed by: Oliver Bruning, 20-Jun-2002 
%  Changed by: Hans Grote, 30-Sep-2002 

\section{USE\_MACRO}
\label{sec:use_macro}
 
It is possible to match user defined expressions with the
\texttt{USE\_MACRO} keyword. 

The general input structure for a \texttt{MACTH} command is the following:

\begin{verbatim}
MATCH,USE_MACRO;
... VARY statements ...
USE_MACRO, NAME=macro1;
     or
macro1: MACRO={ ... madx statements};
CONSTRAINT, expr=  "lhs1 < | = | > rhs1";
CONSTRAINT, expr=  "lhs2 < | = | > rhs2";
...  CONSTRAINT statements ...
MACRO 2 definition
... CONSTRAINT statements ...
MACRO n definition
... CONSTRAINT statements ...
... METHODS statements ...
ENDMATCH;
\end{verbatim}
 
The algorithm for evaluating the penalty function is the following:
 
\begin{itemize}
   \item  execute the first macro,
   \item  evaluate and compute the difference between the left hand side
     (lhs) and the righ hand side (rhs) of the first set of expressions, 
   \item in case of other macros, evaluates in order the macro and the
     expressions 
   \item  the set of differences are  minimized by the selected method
     using the variables defined in the VARY statements. 
\end{itemize}

\subsection{Initiating the Matching Module with USE\_MACRO}
 
With:
\madxmp{
   MATCH, USE\_MACRO;
}
the 'match' command can be used for matching any expression which can be
defined through expression. It requires a slightly different syntax.

\subsection{VARY statements}
In the USE\_MACRO mode the vary statement follows the same rules of the
other modes explained in the section \href{match_vary.html}{Define
  Variable Parameter} 

\subsection{Macro definitions}
The macro to be used in the matching routine can be defined in two ways:
 
\begin{itemize}
\item using USE\_MACRO statement:
  \madxmp{USE\_MACRO, NAME=macro1;}
  defining a new macro on the fly using the usual syntax for
  \href{../control/special.html#macro}{ macros}.  
\end{itemize}
 
After a macro definition a set of constraints should be defined, with
the following syntax for the \texttt{CONSTRAINT} command:
 
\madxmp{
 CONSTRAINT, expr=  "lhs = rhs"; \\
 CONSTRAINT, expr=  "lhs < rhs"; \\
 CONSTRAINT, expr=  "lhs > rhs"; 
}

where "lhs" and "rhs" are well defined \madx expressions. 
Another set of macro and constraints can be defined afterwards. 

\subsection{Examples}
The following example the USE\_MACRO mode can emulate a matching script
which uses the normal syntax. 

Normal syntax:

\begin{verbatim}
MATCH,SEQUENCE=LHCB1,LHCB2;
    VARY, NAME=KSF.B1, STEP=0.00001;
    VARY, NAME=KSD.B1, STEP=0.00001;
    VARY, NAME=KSF.B2, STEP=0.00001;
    VARY, NAME=KSD.B2, STEP=0.00001;
    GLOBAL,SEQUENCE=LHCB1,DQ1=QPRIME;
    GLOBAL,SEQUENCE=LHCB1,DQ2=QPRIME;
    GLOBAL,SEQUENCE=LHCB2,DQ1=QPRIME;
    GLOBAL,SEQUENCE=LHCB2,DQ2=QPRIME;
    LMDIF, CALLS=10, TOLERANCE=1.0E-21;
ENDMATCH;
\end{verbatim}

USE\_MACRO syntax:

\begin{verbatim}
MATCH,USE_MACRO;
    VARY, NAME=KSF.B1, STEP=0.00001;
    VARY, NAME=KSD.B1, STEP=0.00001;
    VARY, NAME=KSF.B2, STEP=0.00001;
    VARY, NAME=KSD.B2, STEP=0.00001;
    M1: MACRO={ TWISS,SEQUENCE=LHCB1; };
    CONSTRAINT, EXPR= TABLE(SUMM,DQ1)=QPRIME;
    CONSTRAINT, EXPR= TABLE(SUMM,DQ2)=QPRIME;
    M2: MACRO={ TWISS,SEQUENCE=LHCB2; };
    CONSTRAINT, EXPR= TABLE(SUMM,DQ1)=QPRIME;
    CONSTRAINT, EXPR= TABLE(SUMM,DQ2)=QPRIME;
    LMDIF, CALLS=10, TOLERANCE=1.0E-21;
ENDMATCH;
\end{verbatim}

%\line(1,0){300}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning},October, 2003;
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, February, 2006.


%\input{match/match_xmpl}
%%\title{the mad program}
%  Changed by: Chris ISELIN, 16-Sep-1997 
%  Changed by: Oliver Bruning, 27-Jun-2002 
%  Changed by: Hans Grote, 25-Sep-2002 
 
\section{Matching Examples}
\label{sec:match_examples}

All matching examples and the related files for executing the MADX
sample jobs can be found in the examples directory. 
\begin{itemize}
	\item Simple Periodic Cell\\
	Match a simple cell to given phase advances: 
        \href{http://cern.ch/madx/madX/examples/match/5cell/job.5cell.madx}{FIVE-CELL}

	\item Simple Periodic Cell\\
	Match the matrix elements of the linear transfer matrix at the
        end of a sequence 5 periodic cells:  
        \href{http://cern.ch/madx/madX/examples/match/r-matrix/job.r-matrix.madx}{RMATRIX}

	\item Transfer line with initial conditions\\
	Match a sequence of 5 periodic cells with initial conditions  to
        given beta-functions at the end of the sequence:  
        \href{http://cern.ch/madx/madX/examples/match/line/job.line.madx}{Transfer line}

	\item Global tune matching in a sequence of 5 periodic cells \\
	Match the global tune of a sequence of 5 periodic cells: 
        \href{http://cern.ch/madx/madX/examples/match/global-tune/job.global-tune.madx}{Global tune}
	
	\item Global tune matching for the LHC\\
	Match the global tune for beam1 of the LHC: 
        \href{http://cern.ch/madx/madX/examples/match/lhc.tune/job.lhc.tune.madx}{Global tune for the LHC}
	
	\item Global chromaticity matching for the LHC\\
	Match the global chromaticity for beam1 of the LHC: 
        \href{http://cern.ch/madx/madX/examples/match/lhc.chromaticity/job.lhc.chromaticity.madx}{Global
          chromaticity for the LHC} 

	\item Global chromaticity matching for both beams of the LHC\\
	Match the global chromaticity for beam1 and beam2 of the LHC: 
        \href{http://cern.ch/madx/madX/examples/match/lhc.2chromaticity/job.lhc.2chromaticity.madx}{Global
          chromaticity for both beams of the LHC} 
	
	\item IR8 insertion matching for beam1 of the LHC\\
	Match the insertion IR8 with initial conditions to given values
        of the optics  functions at the IP and the end of the insertion:  
        \href{http://cern.ch/madx/madX/examples/match/lhc.insertion/job.lhc.insertion.madx}{IR8
          insertion matching} for beam1 of the LHC 
	
	\item IR8 insertion matching for beam1 of the LHC with upper
          limits on the optics functions\\ 
	Match the insertion IR8 with initial conditions to given values
        of the optics  functions at the IP and the end of the insertion
        while limiting the maximum acceptable beta functions over the
        whole insertion:  
        \href{http://cern.ch/madx/madX/examples/match/lhc.insertion-upper/job.lhc.insertion-upper.madx}{IR8
          insertion matching} for beam1 of the LHC with upper limits for
        all beta functions inside the insertion 
	
	\item Simultaneous orbit matching at IP8 for beam1 and beam2 of
          the LHC\\ 
	Match simultaneously the orbit of beam1 and beam of the LHC at
        IP8  with initial conditions to the same given values at the IP:  
        \href{http://cern.ch/madx/madX/examples/match/lhc.iporbit/job.lhc.iporbit.madx}{Orbit
          matching at IP8} for beam1 and beam2 of the LHC 
	
	\item IR8 beta squeeze for beam1 using JACOBIAN matching routine\\
	Try to find a beta squeeze for IR8 starting from 10 meters. 
        \href{http://cern.ch/madx/madX/examples/match/lhcV65.ir8squeeze/job.lhcV65.ir8squeeze.madx}{Beta
          squeeze for IR8} 
	
	\item Matching first and second order chromaticity of the LHC
          using USE\_MACRO option. \\
	Match simultaneously the first and second order chromaticity by
        defining macros which compute them using the TWISS command or
        PTC.  
        \href{http://cern.ch/madx/madX/examples/match/lhc.qpp/job.lhc.qpp.madx}{Second
          order chromaticity} 
	
	\item Matching s position using VLENGTH flag.\\
	match the positions of elements and the total sequence length
        for a simple sample sequence.  
        \href{http://cern.ch/madx/madX/examples/match/s-match/job.s-match.madx}{s position matching}
	
	\item Matching s position using USE\_MACRO.\\
	match the positions of elements and the total sequence length for a simple sample sequence using USE\_MACRO. 
        \href{http://cern.ch/madx/madX/examples/match/s-match-usemacro/job.s-match-usemacro.madx}{s position matching}

\end{itemize}

%\href{http://bruening.home.cern.ch/bruening/}{Oliver Br\"uning}, June, 2002; 
%\href{http://rdemaria.home.cern.ch/rdemaria/}{Riccardo de Maria}, August, 2007. 

%% EOF
