%%\title{Error Definitions}

\chapter{Error Definitions}  
This chapter describes the commands which provide error assignment and
output of errors assigned to elements. It is possible to assign
alignment errors and field errors to single beam elements or to ranges
or classes of beam elements.  

Elements, classes or ranges of elements are selected by the
\href{../Introduction/select.html}{SELECT} command.  

ATTENTION: since errors can only be assigned to machine elements, it is
necessary to  \href{../control/seqedit.html#flatten}{FLATTEN} a sequence
if it includes other sequences.  

Errors can be specified both with a constant or random values. Error
definitions consist of four types of statements listed below. They may
be entered after having selected a beam line by means of a
\href{../control/general.html#use}{USE} command.  

WARNING: any further \href{../control/general.html#use}{USE} command
will destroy the assigned errors. Use the
\href{../error/error_save.html}{ESAVE} option to save and reload errors.  

%\begin{itemize}
%	\item \href{error_align.html}{EALIGN: Define Misalignments}
%	\item \href{error_field.html}{Field Errors}
%\begin{itemize}
%	\item \href{error_field.html#efcomp}{EFCOMP: Components}
%\end{itemize}
%	\item \href{error_option.html}{EOPTION: Set Error Options}
%	\item \href{error_print.html}{EPRINT: List Machine Imperfections}
%	\item \href{error_save.html}{ESAVE: Save Machine Imperfections and read back from file}
%\end{itemize}

%\href{http://consult.cern.ch/xwho/people/1808}{Werner Herr} 18.6.2002 

% add other files to the end of this file

\input{error/error_align}
\input{error/error_field}
\input{error/error_option}
\input{error/error_print}
\input{error/error_save}
