%%\title{MAD-X PTC\_TWISS Module}
%  Created by: Valery KAPIN, 21-Mar-2006 
%  Changed by: Frank Schmidt 04-Apr-2006 

\chapter{PTC\_TWISS Module: Ripken Optics Parameters}


The \textbf{PTC\_TWISS module}of MAD-X [\hyperlink{F._Schmidt}{a}] is
based on the PTC code. It is a supplementary to the TWISS module. The
Twiss parameters are calculated in Ripken's style (invented by G. Ripken
in 1970 [31] and most accessible in
Ref. [\hyperlink{G.&nbsp;Ripken}{b}]). 

These parameters were available
in MAD8 using the TWISS3 command. This module is a typical example of
the advantages when using PTC and its Normal Form technique (and of
course the object-oriented Fortran90 coding): once the rather modest
programming has been performed the Twiss calculation will always be
automatically correct for all machine conditions like closed orbit,
coupling or after a new element has been introduced into the code. In a
traditional coding like in MAD8 this depends on reprogramming and
modifying the code at various places which is inherently error-prone. 

PTC\_TWISS tracks a special representation of the beam in three
degrees of freedom. It works on the coupled lattice functions defined in
Ref.[\hyperlink{G.&nbsp;Ripken}{b}], which are essentially the
projections of the lattice functions for the eigen-modes on the three
planes. 

PTC\_TWISS lists the projections of the ellipses of motion
onto the three planes (\textit{x}, \textit{p$_x$}), (\textit{y},
\textit{p$_y$}), (\textit{t}, \textit{p$_t$}) expressed them via the
Ripken's parameters \textit{b$_k$}$_,\textit{j}$,
\textit{a$_k$}$_,\textit{j, }$\textit{g$_k$}$_,\textit{j}$ along with
the phase advances \textit{m}$_\textit{j}$ in selected positions, where
index \textit{k}=1...3 refers to the plane (\textit{x}, \textit{y},...),
and the index \textit{j}=1...3 denotes the eigen-mode. The PTC\_TWISS
also calculates the dispersion values \textit{D}$_1$,...,\textit{D}$_4$. 
In MAD-X commands and tables, these parameters are denoted as  
\texttt{beta11,...,beta33, alfa11,...,alfa33, gama11,...,gama33,
  mu1,...,mu3, disp1,...,disp4}, respectively. 

The Ripken parametrization can be transformed into the Edwards-Teng
parametrization (used in twiss proper) using the formulae of
Ref. [\hyperlink{V.A.&nbsp;Lebedev}{d}]. The parameters are noted as  
\texttt{betx, bety, alfx, alfy} and the coupling matrix: R11, R12, R21
and R22. In absence of coupling it holds:  
\texttt{betx=beta11}, 
\texttt{bety=beta22}, 
\texttt{alfx=alfa11} and 
\texttt{alfy=alfa22}.

PTC\_TWISS can also compute the deltap/p-dependency of the Twiss
parameters. The column names  
\texttt{beta11p,...,beta33p, alfa11p...,alfa33p, gama11p,...,gama33p}
denote the derivatives of the optics parameters w.r.t. deltap/p. If one
is interested in evaluating deltap/p-dependency of the Twiss parameters,
one must ensure that the order (no) of the map is set to 2 at
least.. The derivatives of the dispersion w.r.t. deltap/p have column
names:  
\texttt{disp1p,...,disp4p}. Second and third order derivatives have
respective column names:  
\texttt{disp1p2,...,disp4p2} for the second order, and 
\texttt{disp1p3,...,disp4p3} for the third order.

In addition, we compute momentum compaction factor  $\alpha_c$ up to 1st
order (for icase=5) or 3rd order (for icase=56). The values appear in
the header of the ptc\_twiss output file (zero means the value has not
been computed). This feature is currently only available in the
development version. 

For clarification: in the 4-D case, there is the following
correspondence between MAD-X and the Ripken's notations:  
\texttt{beta11}
\texttt{®}\textit{b}$_\textit{x}I$, 
\texttt{beta12}
\texttt{®}\textit{b}$_\textit{x}II$, 
\texttt{beta21}
\texttt{®}\textit{b}$_\textit{y}I$, 
\texttt{beta22}
\texttt{®}\textit{b}$_\textit{y}II$,  while in the uncoupled 4-D case 
\texttt{beta11} is the same as the classical \textit{b$_x $}(
\texttt{betx}) and 
\texttt{beta22 }is \textit{b$_y $} (
\texttt{bety}), while 
\texttt{beta12 }and 
\texttt{beta21 } are zero. When there is coupling all 
\texttt{betaNN }are non-zero and 
\texttt{ beta11} , 
\texttt{beta22 }are distinctively different from \textit{b$_x $} ,
\textit{b$_y$}, respectively. 

PTC\_TWISS also tracks the eigenvectors and prints them to Twiss table
according to the SELECT command (flag=ptc\_twiss). Either all 36
components or particular components of the eigenvectors can be selected
with eign or eign\textit{ij}, respectively (\textit{j} = number of
eigenvector, \textit{i} = number of coordinate \{\textit{x}, \textit{
  p$_x$}, \textit{y}, \textit{p$_y$}, \textit{t},\textit{
  p$_t$}\}). 

For ring lattices, PTC\_TWISS computes momentum compaction, transition
energy, as well as other one-turn characteristics such as the tunes
(Q1,Q2 and if icase=6 with cavity Qs) and chromaticities (for no>=2).  

\section{Synopsis}

%% \begin{description}
%% 	\item[Synopsis] \textit{
%% \texttt{PTC\_CREATE\_UNIVERSE;}}\textit{
%% \texttt{PTC\_CREATE\_LAYOUT, model=integer,method=integer,   nst=integer, [exact];}}
%% \texttt{..............................
%% \\   SELECT, flag=ptc\_twiss, clear;}
%% \texttt{SELECT, flag=ptc\_twiss, column=name, s,   
%% \\   beta11,...,beta33,alfa11,..., alfa33,gama11,...,gama33,                 
%% \\                 beta11p,...,beta33p,alfa11p,...,alfa33p,gama11p,...,gama33p,                 
%% \\                 mu1,...,mu3,                 
%% \\                 disp1,...,disp4,                 
%% \\                 disp1p,...,disp4p,                 
%% \\                 disp1p2,...,disp4p2,                 
%% \\                 disp1p3,...,disp4p3,                 }
%% \texttt{[eign], eign11, ...,eign16,...,eign61,...,eign66;}
%% \texttt{..............................}
%% \texttt{PTC\_TWISS;}
%% \texttt{..............................}\textit{
%% \texttt{PTC\_END; }}
%% \end{description}

\begin{verbatim}
PTC_CREATE_UNIVERSE;
PTC_CREATE_LAYOUT, model=integer, method=integer, nst=integer, [exact];
..............................

SELECT, flag=ptc_twiss, clear;
SELECT, flag=ptc_twiss, column=name, s,   
        beta11,...,beta33,alfa11,..., alfa33,gama11,...,gama33,                 
        beta11p,...,beta33p,alfa11p,...,alfa33p,gama11p,...,gama33p,                 
        mu1,...,mu3,                 
        disp1,...,disp4,                 
        disp1p,...,disp4p,                 
        disp1p2,...,disp4p2,                 
        disp1p3,...,disp4p3,                 
        [eign], eign11, ...,eign16,...,eign61,...,eign66;
..............................
PTC_TWISS;
..............................
PTC_END;
\end{verbatim}

\section{Commands}
 
\begin{verbatim}
  PTC_TWISS, icase=integer, deltap=double, closed_orbit, slice_magnets, 
             range=string, file[=string], table[=string],
             initial_matrix_table, initial_matrix_manual, initial_map_manual, beta0=string,   
             betx=double, alfx=double, mux=double,   
             bety=double, alfy=double, muy=double,     
             dx=double,  dpx=double, dy=double, dpy=double,
             x=double, px=double, y=double, py=double, t=double, pt=double,
             re11=double, re12=double, ... , re16=double,
             ................................................................
             re61=double, re62=double, ... , re66=double; 
\end{verbatim}

%% \begin{description}
%% 	\item[Commands] 
%% 	\item{\texttt{\textbf{\textbf{PTC\_TWISS, \\  icase=}integer, }}
%%           \textbf{\texttt{deltap}=double, \textbf{closed\_orbit}, \textbf{slice\_magnets}, \\ 
%%             \textbf{range=}string\textbf{, file[=}string\textbf{], table[=}string\textbf{],
%% \\  initial\_matrix\_table, initial\_matrix\_manual, initial\_map\_manual, beta0=}string\textbf{,   
%% \\  betx=}double,\textbf{ alfx=}double\textbf{, mux=}double\textbf{,   
%% \\  bety=}double,\textbf{ alfy=}double\textbf{, muy=}double\textbf{,   
%% \\  dx=}double,\textbf{ dpx=}double,\textbf{ dy=}double,\textbf{ dpy=}double,\textbf{
%% \\  }\texttt{x}=double, \texttt{px}=double,   \texttt{y}=double, \texttt{py}=double,   \texttt{t}=double, \texttt{pt}=double,
%% \\  \textbf{re11}=double, \textbf{re12}=double, ... ,\textbf{re16}=double,
%% \\    ................................................................
%% \\  \textbf{re61}=double, \textbf{re62}=double, ... ,\textbf{re66}=double;   }\texttt{\textbf{ }}}

\subsection{Description} 
%  <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
% <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE,
% FILE= string;</strong> (MADX version 2)<br /> 
The \texttt{PTC\_TWISS }command causes     computation of the Twiss
parameters in the Ripken's style. It operates on the working beam line
defined in the latest \href{../control/general.html#use}{USE}
command. Several options can be specified, the most important
being \texttt{icase, deltap}, \texttt{closed\_orbit,
  slice\_magnets,}\texttt{no,} and \texttt{file, table}. (see the table
below). 
Other options should be specified  for particular tasks. 
Applications for the \texttt{PTC\_TWISS} command are similar to the
\texttt{\href{../twiss/twiss.html}{TWISS}}-command. 
The PTC\_TWISS can be applied to two basic tasks. It can calculate     either a
\hyperlink{Periodical_Solution}{periodical solution} or a
\hyperlink{Initial_Conditions}{solution with initial conditions}. 

\subsection{Options} 
	
\begin{longtable}{p{3cm} p{5cm} c p{1cm}}
  \hline 
  \textbf{Option} & \textbf{Meaning} & \textbf{Default Value} & \textbf{Value Type} \\ 
  \hline
ICASE & the user-defined dimensionality of the phase-space (4, 5 or
6). Note that the value is internally set to 56 when attempting to set
icase=6 with no cavity. & 4 & integer \\  
\hline
NO & the order of the map, which is often supplied as 1 but must be at
least set to 2 to retrieve chromaticities for instance. For evaluating
the Twiss parameters derivatives w.r.t. deltap/p, one must pass order 2
at least. & 1 & integer \\  
\hline
DELTAP & relative momentum offset for reference closed orbit. & 0.0 & double \\ 
\hline
CLOSED\_ORBIT & the switch to turn on the closed orbit calculation
(\hyperlink{Periodical_Solution}{periodical solution} ONLY).  &
.FALSE. & logical \\  
\hline
DELTAP\_DEPENDENCY & the switch to turn on computation of the Twiss and
dispersion derivatives w.r.t. deltap/p. Derivation formula assume that
icase=5, so that deltap/p is supplied as a parameter. & .FALSE. &
logical \\  
\hline
SLICE\_MAGNETS & the switch to turn on the evaluation of Twiss
parameters at each thin slice inside successive magnets, instead of at
the middle of each magnet. Slices are located at the so-called
'integration nodes' determined by the number of steps (nst) selected
when creating the PTC layout. Note that extremities and fringes are
skipped, whereas only the inner slices are kept. & .FALSE. & logical
\\  
\hline
CENTER\_MAGNETS & the switch to turn on the evaluation of Twiss
parameters at the middle of each magnet. This relies on internal slicing
and 'integration nodes' as determined by the number of steps (nst)
selected when creating the PTC layout. This number is assumed to be even
otherwise the program issues a warning.  & .FALSE. & logical \\  
\hline
\multirow{2}{*}{FILE} & 
omitted  | no output written to a file &  &  \\ 
\hline
&present | the name of the file for printing the PTC\_TWISS output. & ptc\_twiss & string \\ 
\hline
\multirow{2}{*}{TABLE} & omitted  | no output written to a internal table &  &  \\ 
\hline
& present | the name of the internal PTC\_TWISS table  & ptc\_twiss & string \\ 
\hline
\multirow{2}{*}{SUMMARY\_FILE} & omitted  | no output written to a file &  &  \\ 
\hline
& present | the name of the file for printing the PTC\_TWISS\_SUMMARY
table output. & ptc\_twiss\_summary & string \\  
\hline
\multirow{2}{*}{SUMMARY\_TABLE} & omitted  | no output written to a internal table &  &  \\ 
\hline
& present | the name of the internal PTC\_TWISS\_SUMMARY table  & ptc\_twiss\_summary & string \\ 
\hline
\href{../Introduction/ranges.html#range}{RANGE} & specifies a segment of
beam-line for  \\  the PTC\_TWISS calculation & \#s/\#e & string \\  
\hline
INITIAL\_MATRIX\_TABLE & Reading  the transfer matrix from the map-table
in memory (preceding
\texttt{\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL}} command
needed). & .FALSE. & logical \\  
\hline
INITIAL\_MATRIX\_MANUAL & Using the input variables RE11,...,RE66 (see
next entry) as the transfer matrix.  & .FALSE. & logical \\  
\hline
INITIAL\_MAP\_MANUAL & Using the input map stored beforehand in file
fort.18 (e.g. by some initial run of
\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL}). & .FALSE. & logical
\\  
\hline
RE11,..., RE66 & Values of the 6x6 transfer matrix.  & unit matrix & double \\ 
\hline
BETA0 & The name of the \texttt{BETA0}-block, which contains the Twiss
parameters for the input. When icase=6, this information must be
complemented by supplying a value for 'betz' on the ptc\_twiss command
line. & beta0 & string \\  
\hline
betx, alfx, mux, bety, alfy, muy, dx,dpx,dy, dpy &
\href{../twiss/twiss.html}{Twiss and dispersion
  parameters}\textit{b$_x,y$}, \textit{a$_x$}$_,\textit{y,
}$\textit{g$_{x,y},$}\textit{D$_x$},  \textit{D$_px$}, \textit{D$_y$},
\textit{
  D$_py$}\href{../Introduction/bibliography.html#edwards}{(Edwards and
  Teng)}. & 0 & double \\ 
\hline
x, px, y, py, t, pt & The
\href{../Introduction/tables.html#canon}{\textbf{canonical}} coordinates
of the initial orbit. & 0 & double \\  
\hline
\end{longtable}

\subsection{Remarks}

{\bf ICASE}: It can be internally corrected by the
  code. For example, if RF cavity has the voltage set to zero and
  for \hyperlink{ICASE}{icase}=6, the code sets
  \hyperlink{ICASE}{icase}=4. \\

\section{\href{Periodical_Solution}{Periodical Solution}} 
	
\begin{verbatim}
  PTC_TWISS, icase=integer, deltap=double, closed_orbit,
             range=string, file[=string], table[=string];
\end{verbatim}

\subsection{Description} 
  %  <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
  % <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE,
  % FILE= string;</strong> (MADX version 2)<br /> 
This is the simplest form of the \texttt{PTC\_TWISS} command, which
computes the periodic solution for a specified beam line. It may
accept all basic options described in the above table. 






\section{\href{Slicing_Magnets}{Evaluation of Twiss parameters inside magnets}}

\begin{verbatim}
PTC_TWISS, icase=integer, deltap=double, closed_orbit, slice\_magnets, 
           range=string, file[=string], table[=string]; 
\end{verbatim}

%% \begin{description}
%%    \item{\texttt{\textbf{\textbf{PTC\_TWISS, \\  
%%            icase=}integer, }}\textbf{\texttt{deltap}=double,
%%        \textbf{closed\_orbit}, \textbf{ slice\_magnets}\\
%%        \textbf{range=}string\textbf{, file[=}string\textbf{],
%%          table[=}string\textbf{];}}} 
%% \end{description}


\section{Description}

     % <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
     % <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 

This computes the periodic solution for a specified    beam
line and evaluates the Twiss parameters at each thin-slice
(a.k.a "integration-node") inside magnets. The number of such
integration-nodes is given by the number of steps (nst)
selected when creating the PTC layout. All other basic
options described in the above table may be selected.

\subsection{Example}
 An example is found in the
 \href{http://isscvs.cern.ch/cgi-bin/viewcvs-all.cgi/madX-examples/REF/ptc_twiss/SliceMagnets/?root=madx-examples}{\texttt{PTC\_TWISS}
   Examples'} repository.



\section{Solution with \href{Initial_Conditions}{Initial Conditions}}

\subsection{Code Logic}

\begin{verbatim}
    IF ("initial_matrix_table" = ON .AND.
    &    {the map-table exists})   THEN
       (from a Map-Table)

    ELSEIF("initial_map_manual" = ON) THEN
       (from a Given Map File)

    ELSEIF("initial_matrix_manual" = ON) THEN
       (from a Given Matrix)

    ELSEIF(BETA0 block = ON) THEN
       (from Twiss Parameters via BETA0-block)

    ELSE
       (from Given Twiss Parameters)

    ENDIF
\end{verbatim}

\subsection{Initial Values from a Map-Table}

Initial Values from a \href{map-table}{Map-Table} obtainable by a preceding
\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL} command.

\begin{verbatim}
  PTC_TWISS, icase=integer, deltap=double, closed_orbit, 
             range=string, file[=string], table[=string],
             initial_matrix_table;
\end{verbatim}


\subsubsection{Description} 
% <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
% <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
\texttt{PTC\_TWISS} calculates a solution with initial conditions given
as a map-table of preceding ring or beam-line. It requires the input
option \texttt{initial\_matrix\_table} and an existence of the map-table
in memory, which was generated by a preceding
\texttt{\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL}} command. 

\subsubsection{Example}
 An example is found in the
 \href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS}
   Examples} in the folder
 \href{http://cern.ch/frs/mad-X_examples/ptc_twiss/Example3/}{
   "Example3"}. 


\section{Initial Values from a Map-File}


Initial Values from a \href{Given_Map_File}{Map-File} obtainable by a preceding
\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL} command.

\begin{verbatim}
PTC_TWISS, icase=integer, deltap=double, closed_orbit, 
           range=string, file[=string], table[=string],
           initial_map_manual;
\end{verbatim}


 \subsubsection{Description}
 % <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
 % <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
 \texttt{PTC\_TWISS} calculates a solution with initial
 conditions given as a map-file fort.18 obtained from a
 preceding ring or beam-line. It requires the input option
 \texttt{initial\_map\_manual} and an existence of the map-file
 named a fort.18 file, which was generated by a
 preceding
 \texttt{\href{../ptc_normal/ptc_normal.html}{PTC\_NORMAL}}
 command. 


\subsubsection{Example}
An example is found in the
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS} Examples} 
in the folder
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss/Example3/}{"Example3"}. 



\section{Initial Values from a Given Matrix}

Initial Values from a \href{Given_Matrix}{Given Matrix}

\begin{verbatim}
PTC_TWISS, icase=integer, deltap=double,     
           closed_orbit, range=string, file=string, table=string,
           initial_matrix_manual, 
           re11=double, re12=double, ... , re16=double,
           ..............................................
           re61=double, re62=double, ... , re66=double;
\end{verbatim}
         
\subsubsection{Description} 
% <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
% <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
\texttt{PTC\_TWISS} calculates a solution with initial conditions given
by the matrix, which is "manually" entered on the command-line. It
requires the option \texttt{initial\_matrix\_manual}. MAD-X expects a
symplectic 6x6 transfer matrix as input. 

      
\subsubsection{Example} 
An example is found in the
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS} Examples} 
in the folder
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss/Example4/}{"Example4"}. 



\section{Initial Values from Twiss Parameters via BETA0-block}

Initial Values from Twiss Parameters via \href{BETA0-block}{BETA0-block}

\begin{verbatim}
 PTC_TWISS, icase=integer, deltap=double, closed_orbit, 
            range=string, file[=string], table[=string],
            beta0=string ;
\end{verbatim}
      
\subsubsection{Description} 
% <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
% <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
\texttt{PTC\_TWISS} calculates a solution with initial conditions given
by Twiss parameters, which are transferred from the
\texttt{BETA0}-block.   
The data in the the \texttt{BETA0}-block have to be filled by a
combination of the \href{../control/general.html#savebeta}{SAVEBETA} and 
\href{../twiss/twiss.html}{TWISS} commands of a preceding ring or
beam-line. Note, that this case is limited to uncoupled motion of the
preceding machine.  



\subsubsection{Example} 
An example is found in the
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS} Examples} 
in the folder
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss/Example1/}{"Example1"}. 



\section{Initial Values from the Given Twiss Parameters}

Initial Values from the \href{Given_Twiss}{Given Twiss} Parameters

\begin{verbatim}
PTC_TWISS, icase=integer, deltap=double, closed_orbit, 
           range=string, file[=string], table[=string],
           betx=double, alfx=double, mux=double, 
           bety=double, alfy=double, muy=double, 
           dx=double, dpx=double, dy=double, dpy=double,
           x=double, px=double, y=double, py=double, t=double, pt=double;
\end{verbatim}

\subsubsection{Description} 
% <strong>TRACK, DELTAP= double, ONEPASS, DUMP, ONETABLE, FILE= string;</strong> (MADX version 1)<br />
% <strong>TRACK, DELTAP= double, ONEPASS, DAMP, QUANTUM, DUMP, ONETABLE, FILE= string;</strong> (MADX version 2)<br /> 
\texttt{PTC\_TWISS} calculates a solution with initial conditions
given by the Twiss parameters, which are explicitly typed on the
command line. Note, that this case is also limited to uncoupled
motion of the preceding ring or beam-line. 

\subsubsection{Example} 
An example is found in the
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS} Examples} 
in the folder
\href{http://cern.ch/frs/mad-X_examples/ptc_twiss/Example2/}{"Example2"}.  


\section{References for PTC\_TWISS}

\begin{enumerate}
   \item \href{F._Schmidt}{F. Schmidt},
     "`\href{http://cern.ch/madx/doc/MPPE012.pdf}{MAD-X PTC
     Integration}'',   Proc. of the 2005 PAC Conference in Knoxville,
     USA, pp.1272. 
   \item \href{G.&nbsp;Ripken}{G.Ripken} and F.Willeke,
     "\href{http://cern.ch/madx/doc/1988_Ripken.pdf}{Methods of   Beam
     Optics}", DESY 88–114, 1988. 
   \item \href{K._Zhang}{K. Zhang}, "PTC twiss with initial TWISS
     parameters",
     \href{http://cern.ch/frs/MAD-X_minutes/Meeting-13/minutes.html}{
       MAD-X Meeting 13 (04.07.2005)}, slides in
     \href{http://cern.ch/frs/MAD-X_minutes/Meeting-13/Input_of_PTC_TWISS_module.ppt}{
       ppt}. 
   \item \href{V.A.&nbsp;Lebedev}{V.A.Lebedev} and S.A.Bogacsz,
     "\href{http://cern.ch/madx/doc/1748-0221_5_10_P10010.pdf}{Betatron
     motion with coupling of horizontal and vertical degrees of
     motion}", Thomas Jefferson National Accelerator Facility 2010. 
\end{enumerate}

\bf{See Also} 
   \texttt{\href{../twiss/twiss.html}{TWISS}},
   \href{http://cern.ch/frs/mad-X_examples/ptc_twiss}{\texttt{PTC\_TWISS}
     Examples}. 



%\href{mailto:kapin@itep.ru}{ V.Kapin}(ITEP) and 
%\href{mailto:Frank.Schmidt@cern.ch}{ F.Schmidt}, March  2006

