%%\title{Dipedge}
%  Changed by: Frank Schmidt, 27-FEB-2005 

\section{Dipedge Element}

A thin element describing the edge focusing of a dipole has been
introduced in order to make it possible to track trajectories in the
presence of dipoles with pole face angles. Only linear terms are
considered since the higher order terms would the tracking
non-symplectic. The transformation of the machine elements into thin
lenses leaves dipedge untouched and splits correctly the SBENDS's.  

It does not make sense to use it alone. It can be specified at the
entrance and the exit of a SBEND. They are defined by the commands:  
\begin{verbatim}
label : dipedge, h = real, e1 = real, fint = real, hgap = real, tilt = real;
\end{verbatim} 

It has zero length and five attributes. 
\begin{itemize}
   \item H: Is angle/length or 1/rho (default: 0 m$^{-1}$ - for the
     default the dipedge element has no effect). (must be equal to that
     of the associated SBEND)  
   \item E1: The rotation angle for the pole face.The sign convention is
     as for a SBEND \href{bend.html}{Bending Magnet}. Note that it is
     different for an entrance and an exit. (default: 0 rad).  
   \item FINT: field integral as for SBEND
     \href{local_system.html#sbend}{sector bend}. Note that each dipedge
     has its own fint, so fintx is no longer necessary.  
   \item HGAP: half gap height of the associated SBEND
     \href{bend.html}{Bending Magnet}.   
   \item TILT: The roll angle about the longitudinal axis (default: 0
     rad, i.e. a horizontal bend). A positive angle represents a
     clockwise rotation.  
\end{itemize}

%\href{http://www.cern.ch/Frank.Schmidt/frs_sign.html}{frs}, February 27, 2005  
