%%\title{The MAD Program}
%  Changed by: Chris ISELIN, 17-Jul-1997 
%  Changed by: Hans Grote, 10-Jun-2002 
\label{module_doc}

\def\documentlabel#1{\gdef\@documentlabel{#1}}
\gdef\@documentlabel{\tt Preliminary draft}
%\documentlabel{CERN/ATS/... \\ (Rev. 1)}


\begin{titlepage}

\begin{center}\normalsize
EUROPEAN LABORATORY FOR PARTICLE PHYSICS
\end{center}
\vskip 0.7cm

\begin{flushright}
\@documentlabel \\                   % document label
\end{flushright}
\vskip 2.3cm

\begin{center}\LARGE                 % document title
{\bf The MAD-X Program} \\
(Methodical Accelerator Design) \\
Version 5.01.00 \\
{\bf User's Reference Manual}
\end{center}
\vskip 1.5em

\begin{center}                       % authors
Hans Grote \\
Frank Schmidt \\
Laurent Deniau
\vskip 2em

{\large Abstract}
\end{center}
\begin{quotation}
MAD-X is a tool for charged-particle optics in alternating-gradient
accelerators and beam lines. It can handle very large and very small
accelerators and solves various problems on such machines.  

\textbf{{ This new version supersedes MAD-8 which has been frozen.}}

The input format has been modified slightly, mainly in order to make it
safer, and to allow the introduction of new features (WHILE loop,
MACROs, and others).  

The MAD-X framework should make it easy to add new features in the form
of program modules. A \href{module_doc.html}{ Module Writer's Guide} is
available and contains guidelines and examples. The authors of MAD hope
that such modules will also be contributed and documented by others. As
required, the table of contents and the indices will be revised. The
contributions of other authors are acknowledged in the relevant
chapters.  

Last but not least it is only fair to mention that a large part of the
documentation provided here is based on the MAD-8 documentation that was
originally written and published by F. C. Iselin.  

Comments from readers are most welcome. They may be sent to the
following email address: \texttt{mad@cern.ch}
\end{quotation}
\vfill

\begin{center}
Geneva, Switzerland \\
\today
\end{center}

\end{titlepage}



