%%\title{Element Input Format}
%  Changed by: Chris ISELIN, 24-Jan-1997 
%  Changed by: Hans Grote, 22-Jan-2003 

\section{Element Input Format}
All physical elements are defined by statements of the form 
\begin{verbatim}
label: keyword {,attribute};
\end{verbatim} 
where 
\begin{itemize}
	\item \href{label.html}{label} is a name to be given to the element (in the example QF), 
	\item \href{keyword.html}{keyword} is an element type keyword (in the example QUADRUPOLE). 
	\item \href{attribute.html}{attribute} normally has the form
          "attribute-name = attribute-value" or
          "attribute-name := attribute-value" (except for multipoles).  
	\item \href{label.html}{attribute-name} selects the attribute,
          as defined for the element type keyword (in the example L and
          K1).  
	\item \href{attribute.html}{attribute-value} gives it a value
          (in the example 1.8 and 0.015832). The value may be specified
          by an expression. The "=" assigns the value on the right to
          the attribute at the time of definition, regardless of any
          further change of the right hand side; the ":=" re-evaluates
          the expression at the right every time the attribute is being
          used. For constant right hand sides, "=" and ":=" are of
          course equivalent.  
\end{itemize} 

Omitted attributes are assigned a default value, normally zero. 

Example: 
\begin{verbatim}
QF: QUADRUPOLE, L=1.8, K1=0.015832;
\end{verbatim} 

A special format is used for a \href{multipole.html}{multipole}: 
\begin{verbatim}
m:multipole, kn= {kn0, kn1, kn2, ..., knmax},
             ks= {ks0, ks1, ks2, ..., ksmax};
\end{verbatim} 
where kn and ks give the integrated normal and skew strengths,
respectively. The commas are mandatory, each strength can be an
expression; their position defines the order. example:  
\begin{verbatim}
m: multipole, kn={0,0,0,myoct*lrad}, ks={0,0,0,0,-1.e-5};
\end{verbatim} 
defines a multipole with a normal octupole, and a skew decapole component. 

To know the current maximum order, enter the command 
\begin{verbatim}
help, multipole;
\end{verbatim} 
and count. 

%\href{http://www.cern.ch/hansg/hansg_sign.html}{hansg}, January 24, 1997 
